---
title: "final result"
output:
  word_document: default
  html_document: default
---

# data preparation
```{r}
setwd("C:/Users/YUHONG/Documents/R/DataScienceProject")
# load data
data = read.csv("datasets/all_plots.csv", stringsAsFactors = FALSE)
```

only extract DBH-related data for DBH modelling
```{r}
DBHdata = data[1:9]
```

check NAs and blank
```{r}
# blank "Large", assume "No" where it's blank
DBHdata[DBHdata$Large == "", "Tree.Number"]
DBHdata[DBHdata$Large == "", "Large"] = "No"
```
```{r}
# remove rows with NA in "DBH.year.6"
DBHdata[is.na(DBHdata$DBH.year.6), 1:7] # 9 NAs for "DBH.year.6"
DBHdata = DBHdata[!is.na(DBHdata$DBH.year.6), ]
```
#------------------------------------------------------------------------------------------
Referring to lower priority hypotheses: 
- The rate of natural mortality will be greater in control plots as compared to T1 and T2.
NOT satisfy: all blank Alive.at.year.6 are in T2:Radial.
#-------------------------------------------------------------------------------------------

```{r}
# remove rows with NA in "DBH.year.0"
DBHdata[is.na(DBHdata$DBH.year.0), 1:7] # 2 NAs for "DBH.year.0"
DBHdata = DBHdata[!is.na(DBHdata$DBH.year.0), ]
```
```{r}
# NA in "Class.year.0" and NA in "Class.year.6", no modification
DBHdata[is.na(DBHdata$Class.year.0) | is.na(DBHdata$Class.year.6), ]
# both in Plot5(T2: Radial), Tree.Number 148, 149, they have extremely large DBH.
```

make Treatment and Large factor
```{r}
DBHdata$Treatment = factor(DBHdata$Treatment)
DBHdata$Large = factor(DBHdata$Large)
```

rows where DBH.year.6 <= DBH.year.0 (tree growth <= 0)
```{r}
DBHdata[DBHdata$DBH.year.6 <= DBHdata$DBH.year.0, ]
```






# data analysis
```{r}
# points outside the main group
par(mfrow=c(1,2))
plot(DBHdata$DBH.year.0, DBHdata$DBH.year.6, type = "p", col = factor(DBHdata$DBH.year.6/DBHdata$DBH.year.0 > 1.25 | DBHdata$DBH.year.6/DBHdata$DBH.year.0 < 0.9 | DBHdata$DBH.year.0 > 80))
plot(DBHdata$DBH.year.0, DBHdata$DBH.year.6/DBHdata$DBH.year.0, type = "p", col = factor(DBHdata$DBH.year.6/DBHdata$DBH.year.0 > 1.25 | DBHdata$DBH.year.6/DBHdata$DBH.year.0 < 0.9 | DBHdata$DBH.year.0 > 80), ylab = "gradient")
DBHdata[DBHdata$DBH.year.6/DBHdata$DBH.year.0 > 1.25 | DBHdata$DBH.year.6/DBHdata$DBH.year.0 < 0.9 | DBHdata$DBH.year.0 > 80, ]

# remove these outliers
DBHdatarm = DBHdata[DBHdata$DBH.year.6/DBHdata$DBH.year.0 < 1.25 & DBHdata$DBH.year.6/DBHdata$DBH.year.0 > 0.9 & DBHdata$DBH.year.0 < 80, ]
```

```{r}
cor(DBHdata$DBH.year.0, DBHdata$DBH.year.6)
cor(log(DBHdata$DBH.year.0), log(DBHdata$DBH.year.6))

cor(DBHdatarm$DBH.year.0, DBHdatarm$DBH.year.6)
cor(log(DBHdatarm$DBH.year.0), log(DBHdatarm$DBH.year.6))
```






# random effect
```{r}
par(mfrow=c(3,4))
for (i in 1:12) {
  plot(DBHdata[DBHdata$Plot == i, "DBH.year.0"], DBHdata[DBHdata$Plot == i, "DBH.year.6"], xlab = paste("Plot", i, "DBH.year.0"), ylab = paste("Plot", i, "DBH.year.6"), xlim = c(20,70), ylim = c(20,70), col = DBHdata[DBHdata$Plot == i, "Treatment"])
  abline(0,1)
}

for (i in 1:12) {
  plot(log(DBHdata[DBHdata$Plot == i, "DBH.year.0"]), log(DBHdata[DBHdata$Plot == i, "DBH.year.6"]), xlab = paste("Plot", i, "log(DBH.year.0)"), ylab = paste("Plot", i, "log(DBH.year.6)"), xlim = c(3.0,4.3), ylim = c(3.0, 4.3), col = DBHdata[DBHdata$Plot == i, "Treatment"])
  abline(0,1)
}
```






# "In Gap" term
add "InGap" term
```{r}
DBHdatarm$InGap = (DBHdatarm$Treatment == "T2: Radial" & DBHdatarm$Large == "Yes")
```
make InGap factor
```{r}
DBHdatarm$InGap = factor(DBHdatarm$InGap)
```






# current final model
random effect using InGap for log(DBHdatarm$DBH.year.6)
```{r}
library(lme4)
mixed2.0 = lmer(log(DBH.year.6)~log(DBH.year.0)+Treatment+InGap+log(DBH.year.0):InGap+(1|Plot), data = DBHdatarm)
summary(mixed2.0)
VarCorr(mixed2.0)
y.hat <- fitted(mixed2.0) # Fitted values
int.hat <- ranef(mixed2.0)[[1]][[1]] # Predicted intercepts
res.hat <- residuals(mixed2.0) # Estimated residuals
tss = sum((log(DBHdatarm$DBH.year.6)-mean(log(DBHdatarm$DBH.year.6)))^2) #TSS
rss = sum((log(DBHdatarm$DBH.year.6)-y.hat)^2) #RSS
(tss-rss)/tss #R^2 = 0.9801595
```
```{r}
qqnorm(int.hat, main="Random Intercepts"); qqline(int.hat)
qqnorm(res.hat, main="Residuals"); qqline(res.hat)
plot(y.hat, res.hat, xlab="Fitted Values", ylab="Residuals")
abline(h=0, lty=2)
plot(log(DBHdatarm$DBH.year.6), y.hat, col = DBHdatarm$Treatment)
abline(0, 1)
plot(log(DBHdatarm$DBH.year.6), y.hat, col = DBHdatarm$Large)
abline(0, 1)
plot(log(DBHdatarm$DBH.year.6), y.hat, col = DBHdatarm$InGap)
abline(0, 1)
```






# other plots/ interpretation
```{r}
# add tree growth attribute
DBHdata$Tree.growth = DBHdata$DBH.year.6-DBHdata$DBH.year.0

plot(DBHdata$DBH.year.0, DBHdata$Tree.growth, col = DBHdata$Large, ylim = c(-10,20))
legend("topright", legend = c("large", "not large"), fill = c("red", "black"), cex = 0.8)

plot(DBHdata[DBHdata$Treatment == "Control", "DBH.year.0"], DBHdata[DBHdata$Treatment == "Control", "Tree.growth"], col = DBHdata[DBHdata$Treatment == "Control", "Large"], xlab = "Control DBH.year.0", ylab = "Control tree growth", ylim = c(-10,20))
legend("topright", legend = c("large", "not large"), fill = c("red", "black"), cex = 0.8)

plot(DBHdata[DBHdata$Treatment == "T1: Gap", "DBH.year.0"], DBHdata[DBHdata$Treatment == "T1: Gap", "Tree.growth"], col = DBHdata[DBHdata$Treatment == "T1: Gap", "Large"], xlab = "T1 DBH.year.0", ylab = "T1 tree growth", ylim = c(-10,20))
legend("topright", legend = c("large", "not large"), fill = c("red", "black"), cex = 0.8)

plot(DBHdata[DBHdata$Treatment == "T2: Radial", "DBH.year.0"], DBHdata[DBHdata$Treatment == "T2: Radial", "Tree.growth"], col = DBHdata[DBHdata$Treatment == "T2: Radial", "Large"], xlab = "T2 DBH.year.0", ylab = "T2 tree growth", ylim = c(-10,20))
legend("topright", legend = c("large", "not large"), fill = c("red", "black"), cex = 0.8)
```

```{r}
# plot constrain according to mixed2.0
plot(DBHdatarm$DBH.year.0, DBHdatarm$DBH.year.6, col = DBHdatarm$InGap, xlim = c(20,80))
abline(v = 52.76, col = "blue") # constrain for T2InGap > T2 not InGap
abline(v = 57.87, col = "green") # constrain for T2InGap > T1
abline(v = 74.42, col = "orange") # constrain for T2InGap > Control
legend("topleft", c("InGap points", "not InGap points", "InGap>T2notInGap constrain", "InGap>T1 constrain", "InGap>Control constrain"), fill = c("black", "red", "blue", "green", "orange"), cex = 0.9)
```

```{r}
plot(DBHdatarm[DBHdatarm$Treatment == "Control", "DBH.year.0"], DBHdatarm[DBHdatarm$Treatment == "Control", "DBH.year.6"], xlim = c(20,70), ylim = c(20,70), xlab = "DBHdatarm$DBH.year.0", ylab = "DBHdatarm$DBH.year.6")
points(DBHdatarm[DBHdatarm$InGap == TRUE, "DBH.year.0"], DBHdatarm[DBHdatarm$InGap == TRUE, "DBH.year.6"], col = "red")
legend("bottomright", legend = c("Control", "InGap"), fill = c("black", "red"))

plot(DBHdatarm[DBHdatarm$Treatment == "T1: Gap", "DBH.year.0"], DBHdatarm[DBHdatarm$Treatment == "T1: Gap", "DBH.year.6"], xlim = c(20,70), ylim = c(20,70), xlab = "DBHdatarm$DBH.year.0", ylab = "DBHdatarm$DBH.year.6")
points(DBHdatarm[DBHdatarm$InGap == TRUE, "DBH.year.0"], DBHdatarm[DBHdatarm$InGap == TRUE, "DBH.year.6"], col = "red")
legend("bottomright", legend = c("T1: Gap", "InGap"), fill = c("black", "red"))

plot(DBHdatarm[DBHdatarm$Treatment == "T2: Radial", "DBH.year.0"], DBHdatarm[DBHdatarm$Treatment == "T2: Radial", "DBH.year.6"], xlim = c(20,70), ylim = c(20,70), xlab = "DBHdatarm$DBH.year.0", ylab = "DBHdatarm$DBH.year.6")
points(DBHdatarm[DBHdatarm$InGap == TRUE, "DBH.year.0"], DBHdatarm[DBHdatarm$InGap == TRUE, "DBH.year.6"], col = "red")
legend("bottomright", legend = c("T2: Radial", "InGap"), fill = c("black", "red"))
```
```{r}
plot(DBHdatarm[DBHdatarm$Treatment != "T2: Radial", "DBH.year.0"], DBHdatarm[DBHdatarm$Treatment != "T2: Radial", "DBH.year.6"], col = DBHdatarm[DBHdatarm$Treatment != "T2: Radial", "Treatment"])

plot(DBHdatarm[DBHdatarm$Treatment != "T1: Gap", "DBH.year.0"], DBHdatarm[DBHdatarm$Treatment != "T1: Gap", "DBH.year.6"], col = DBHdatarm[DBHdatarm$Treatment != "T1: Gap", "Treatment"])
```

```{r}
#library(merTools)
X_ingap <- data.frame('DBH.year.0'=1:100, 'InGap'=TRUE, 'Plot'=6, 'Treatment'='T2: Radial')
PI_2 <- predictInterval(merMod = mixed2.0, newdata = X_ingap, level = 0.95, n.sims = 1000, stat = "median", type="linear.prediction", include.resid.var = TRUE)


for (t in c("Control", "T1: Gap", "T2: Radial")) {
  X <- data.frame('DBH.year.0'=1:100, 'InGap'=FALSE, 'Plot'=6, 'Treatment'=t)
  PI <- predictInterval(merMod = mixed2.0, newdata = X, level = 0.95, n.sims = 1000, stat = "median", type="linear.prediction", include.resid.var = TRUE)
  
  plot(1:100, exp(PI$fit), type='n', xlim=c(20,60), ylim = c(10,80), xlab = "DBH.year.0", ylab = "predicted DBH.year.6")
  polygon(c(1:100, 100:1), c(exp(PI$upr), rev(exp(PI$lwr))), col = "light grey", border = "light grey")
  lines(1:100, exp(PI$fit), type = "l")
  polygon(c(1:100, 100:1), c(exp(PI_2$upr), rev(exp(PI_2$lwr))), col = rgb(255,0,0, 50, maxColorValue = 255), border = rgb(255,0,0, 50, maxColorValue = 255))
  lines(1:100, exp(PI_2$fit), type = "l", col = "red")
  legend("topleft", legend = c("InGap", t), fill = c("red", "black"))
}
```





