---
title: "Model development process"
output: html_document
---

This file contains model development process. The selected models are the ones that are representative. The content is identical to Tree model development in the report.

# load packages
```{r}
setwd("C:/Users/YUHONG/Documents/R/DataScienceProject")
library(car)
library(fitdistrplus)
library(lme4)
library(MASS)
```

# data preparation
```{r}
# load data
data = read.csv("datasets/all_plots.csv", stringsAsFactors = FALSE)

# only extract DBH-related data for DBH modelling
DBHdata = data[1:9]
```

check NAs and blank
```{r}
# blank "Large", assume "No" where it's blank
DBHdata[DBHdata$Large == "", "Tree.Number"]
DBHdata[DBHdata$Large == "", "Large"] = "No"
```
```{r}
# remove rows with NA in "DBH.year.6"
DBHdata[is.na(DBHdata$DBH.year.6), 1:7] # 9 NAs for "DBH.year.6"
DBHdata = DBHdata[!is.na(DBHdata$DBH.year.6), ]
```
#------------------------------------------------------------------------------------------
Referring to lower priority hypotheses: 
- The rate of natural mortality will be greater in control plots as compared to T1 and T2.
NOT satisfy: all blank Alive.at.year.6 are in T2:Radial.
#-------------------------------------------------------------------------------------------
```{r}
# remove rows with NA in "DBH.year.0"
DBHdata[is.na(DBHdata$DBH.year.0), 1:7] # 2 NAs for "DBH.year.0"
DBHdata = DBHdata[!is.na(DBHdata$DBH.year.0), ]
```
```{r}
# NA in "Class.year.0" and NA in "Class.year.6", no modification
DBHdata[is.na(DBHdata$Class.year.0) | is.na(DBHdata$Class.year.6), ]
# both in Plot5(T2: Radial), Tree.Number 148, 149, they have extremely large DBH.
```

make Treatment and Large factor
```{r}
DBHdata$Treatment = factor(DBHdata$Treatment)
DBHdata$Large = factor(DBHdata$Large)
```

rows where DBH.year.6 <= DBH.year.0 (tree growth <= 0)
```{r}
DBHdata[DBHdata$DBH.year.6 <= DBHdata$DBH.year.0, ]
DBHpositive = DBHdata[DBHdata$DBH.year.6 > DBHdata$DBH.year.0, ] # remove rows with tree growth<=0
```






# model development process
#-------------------------------------------------------------------------------------------
add Tree.growth
```{r}
DBHdata$Tree.growth = DBHdata$DBH.year.6 - DBHdata$DBH.year.0
```
```{r}
plot(DBHdata$DBH.year.0, DBHdata$Tree.growth, type = "p", col = DBHdata$Treatment, xlab = "DBH at Year 0", ylab = "Tree growth", main = "Comparing tree growth with treatments")
legend("topright", legend = c("Control", "T1: Gap", "T2: Radial"), fill = c("black", "red", "green"), bty='n')

plot(DBHdata$DBH.year.0, DBHdata$Tree.growth, type = "p", col = DBHdata$Large, xlab = "DBH.year.0", ylab = "Tree.growth")
legend("topright", legend = c("Large trees", "not Large trees"), fill = c("red", "black"))
```

# model0
data: raw data (DBHdata)
model: gaussian glm
response variable: Tree.growth = DBH.year.6 - DBH.year.0
predictors: DBH.year.0, Treatment, Large, DBH.year.0:Large, Treatment:Large
```{r}
# best model derived from step(Tree.growth~DBH.year.0*Large*Treatment)
model0 = glm(Tree.growth~DBH.year.0+Treatment+Large+DBH.year.0:Large+Treatment:Large, data = DBHdata)
summary(model0)

tss = sum((DBHdata$Tree.growth-mean(DBHdata$Tree.growth))^2) # TSS
rss = sum((DBHdata$Tree.growth-model0$fitted.values)^2) # RSS
(tss-rss)/tss # R^2 = 0.2433805

plot(model0)
plot(DBHdata$Tree.growth, model0$fitted.values, type = "p", col = DBHdata$Treatment)
abline(0, 1)
```
Problems:
1. very small R^2 value
2. large residuals compared to response variable range
3. not flat residuals
4. very heavy tails in the Normal Q-Q plot -> distribution is not normal

# distribution of Tree.growth
```{r}
shapiro.test(DBHdata$Tree.growth) # not normal, p-value < 2.2e-16
leveneTest(Tree.growth~Treatment, data = DBHdata) # different variance
leveneTest(Tree.growth~Large, data = DBHdata) # different variance
descdist(DBHdata$Tree.growth) # closer to lognormal distribution
```
log(Tree.growth) might be close to normal distribution. In order to take log on Tree.growth, we use the data which removes Tree.growth<=0 (DBHpositive).
```{r}
shapiro.test(log(DBHpositive$Tree.growth)) # not normal, but closer to normal than DBHdata$Tree.growth, p-value = 1.237e-05
leveneTest(log(Tree.growth)~Treatment, data = DBHpositive) # consistent variance
leveneTest(log(Tree.growth)~Large, data = DBHpositive) # consistent variance
descdist(log(DBHpositive$Tree.growth))
```
Although taking log on Tree.growth makes variance consistent across groups, variance of both Tree.growth and log(Tree.growth) is still very large (variance is about twice mean)
```{r}
print(paste("DBHdata$Tree.growth: mean=", mean(DBHdata$Tree.growth), "variance=", var(DBHdata$Tree.growth)))
print(paste("DBHpositive$Tree.growth: mean=", mean(log(DBHpositive$Tree.growth)), "variance=", var(log(DBHpositive$Tree.growth))))
```

extra plots for distribution of Tree.growth
```{r}
fitdis1 = fitdist(DBHdata$Tree.growth, "norm")
summary(fitdis1)
par(mfrow = c(2, 2))
denscomp(fitdis1, legendtext = "Normal")
qqcomp(fitdis1, legendtext = "Normal")
cdfcomp(fitdis1, legendtext = "Normal")
ppcomp(fitdis1, legendtext = "Normal")

fitdis2 = fitdist(log(DBHpositive$Tree.growth), "norm")
summary(fitdis2)
par(mfrow = c(2, 2))
denscomp(fitdis2, legendtext = "Normal")
qqcomp(fitdis2, legendtext = "Normal")
cdfcomp(fitdis2, legendtext = "Normal")
ppcomp(fitdis2, legendtext = "Normal")
```



#-------------------------------------------------------------------------------------------
# model1
data: remove Tree.growth<=0 in order to take log on Tree.growth (DBHpositive)
model: gaussian glm
response variable: log(Tree.growth)
predictors: log(DBH.year.0), Treatment, Large, log(DBH.year.0):Large, Treatment:Large
```{r}
model1 = glm(log(Tree.growth)~log(DBH.year.0)+Treatment+Large+log(DBH.year.0):Large+Treatment:Large, data = DBHpositive)
summary(model1)

tss = sum((log(DBHpositive$Tree.growth)-mean(log(DBHpositive$Tree.growth)))^2) # TSS
rss = sum((log(DBHpositive$Tree.growth)-model1$fitted.values)^2) # RSS
(tss-rss)/tss # R^2 = 0.3530835

plot(model1)
plot(log(DBHpositive$Tree.growth), model1$fitted.values, type = "p", col = DBHpositive$Treatment)
abline(0, 1)
```
Problems:
1. still quite small R^2 value
2. large residuals compared to response variable range
3. heavy tails in the Normal Q-Q plot



#-------------------------------------------------------------------------------------------
# other approaches see the end of this document



#-------------------------------------------------------------------------------------------
# model2
data: raw data (DBHdata)
model: gaussian glm
response variable: DBH.year.6
predictors: DBH.year.0, Treatment, Large, DBH.year.0:Large, Treatment:Large
```{r}
# best model derived from DBH.year.6~DBH.year.0*Treatment*Large
model2 = glm(DBH.year.6~DBH.year.0+Treatment+Large+DBH.year.0:Large+Treatment:Large, data = DBHdata)
summary(model2)

tss = sum((DBHdata$DBH.year.6-mean(DBHdata$DBH.year.6))^2) # TSS
rss = sum((DBHdata$DBH.year.6-model2$fitted.values)^2) # RSS
(tss-rss)/tss # R^2 = 0.9467556

plot(model2)
plot(DBHdata$DBH.year.6, model2$fitted.values, type = "p", col = DBHdata$Treatment)
legend("bottomright", legend = c("Control", "T1", "T2"), fill = c("black", "red", "green"))
abline(0,1)
```



#-------------------------------------------------------------------------------------------
remove outliers
```{r}
# points outside the main group
plot(DBHdata$DBH.year.0, DBHdata$DBH.year.6/DBHdata$DBH.year.0, type = "p", col = factor(DBHdata$DBH.year.6/DBHdata$DBH.year.0 > 1.25 | DBHdata$DBH.year.6/DBHdata$DBH.year.0 < 0.9 | DBHdata$DBH.year.0 > 80), xlab = "DBH.year.0", ylab = "gradient")
legend("bottomright", legend = "outliers in raw data", fill = "red")

plot(DBHdata$DBH.year.0, DBHdata$DBH.year.6, type = "p", col = factor(DBHdata$DBH.year.6/DBHdata$DBH.year.0 > 1.25 | DBHdata$DBH.year.6/DBHdata$DBH.year.0 < 0.9 | DBHdata$DBH.year.0 > 80), xlab = "DBH at Year 0", ylab = "DBH at Year 6", main = "Identifying outliers")
legend("bottomright", legend = "outliers in raw data", fill = "red", bty='n')

DBHdata[DBHdata$DBH.year.6/DBHdata$DBH.year.0 > 1.25 | DBHdata$DBH.year.6/DBHdata$DBH.year.0 < 0.9 | DBHdata$DBH.year.0 > 80, ]

# remove these outliers
DBHdatarm = DBHdata[DBHdata$DBH.year.6/DBHdata$DBH.year.0 < 1.25 & DBHdata$DBH.year.6/DBHdata$DBH.year.0 > 0.9 & DBHdata$DBH.year.0 < 80, ]
```

# model3
data: remove outliers (DBHdatarm)
model: gaussian glm
response variable: log(DBH.year.6)
predictors: log(DBH.year.0), Treatment, Large, log(DBH.year.0):Large, Treatment:Large
```{r}
model3 = glm(log(DBH.year.6)~log(DBH.year.0)+Treatment+Large+log(DBH.year.0):Large+Treatment:Large, data = DBHdatarm)
summary(model3)

tss = sum((log(DBHdatarm$DBH.year.6)-mean(log(DBHdatarm$DBH.year.6)))^2) # TSS
rss = sum((log(DBHdatarm$DBH.year.6)-model3$fitted.values)^2) # RSS
(tss-rss)/tss # R^2 = 0.9783684

plot(model3)
plot(log(DBHdatarm$DBH.year.6), model3$fitted.values, type = "p", col = DBHdatarm$Treatment)
legend("bottomright", legend = c("Control", "T1", "T2"), fill = c("black", "red", "green"))
abline(0,1)
```



#-------------------------------------------------------------------------------------------
add "InGap" term
```{r}
DBHdatarm$InGap = (DBHdatarm$Treatment == "T2: Radial" & DBHdatarm$Large == "Yes")
DBHdatarm$InGap = factor(DBHdatarm$InGap) # make InGap factor
```
```{r}
plot(DBHdatarm$DBH.year.0, DBHdatarm$DBH.year.6, type = "p", col = DBHdatarm$InGap, xlab = "DBH at Year 0", ylab = "DBH at Year at 6", main = "Comparing DBH at Year 6 with InGap")
legend("bottomright", legend = c("InGap trees", "not InGap trees"), fill = c("red", "black"), bty='n')
```

# model3 using InGap - model3.2
data: remove outliers (DBHdatarm)
model: gaussian glm
response variable: log(DBH.year.6)
predictors: log(DBH.year.0), Treatment, InGap, log(DBH.year.0):InGap
```{r}
model3.2 = glm(log(DBH.year.6)~log(DBH.year.0)+Treatment+InGap+log(DBH.year.0):InGap, data = DBHdatarm)
summary(model3.2)

tss = sum((log(DBHdatarm$DBH.year.6)-mean(log(DBHdatarm$DBH.year.6)))^2) # TSS
rss = sum((log(DBHdatarm$DBH.year.6)-model3.2$fitted.values)^2) # RSS
(tss-rss)/tss # R^2 = 0.9786269

plot(model3.2)
plot(log(DBHdatarm$DBH.year.6), model3.2$fitted.values, type = "p", col = DBHdatarm$Treatment)
legend("bottomright", legend = c("Control", "T1", "T2"), fill = c("black", "red", "green"))
abline(0,1)
```



#-------------------------------------------------------------------------------------------
# model4
data: remove outliers (DBHdatarm)
model: mixed effect model with random intercept for each Plot
response variable: log(DBH.year.6)
predictors:
fixed effect: log(DBH.year.0), Treatment, InGap, log(DBH.year.0):InGap;
random effect: random intercept for Plot (1|Plot)
```{r}
model4 = lmer(log(DBH.year.6)~log(DBH.year.0)+Treatment+InGap+log(DBH.year.0):InGap+(1|Plot), data = DBHdatarm)
summary(model4)
y.hat <- fitted(model4) # Fitted values
int.hat <- ranef(model4)[[1]][[1]] # Predicted intercepts
res.hat <- residuals(model4) # Estimated residuals
tss = sum((log(DBHdatarm$DBH.year.6)-mean(log(DBHdatarm$DBH.year.6)))^2) # TSS
rss = sum((log(DBHdatarm$DBH.year.6)-y.hat)^2) # RSS
(tss-rss)/tss # R^2 = 0.9801595
```
```{r}
qqnorm(int.hat, main="Random Intercepts"); qqline(int.hat)
qqnorm(res.hat, main="Normal Q-Q"); qqline(res.hat)
plot(y.hat, res.hat, xlab="Fitted Values", ylab="Residuals", main = "Residuals vs Fitted")
abline(h=0, lty=2)
plot(log(DBHdatarm$DBH.year.6), y.hat, col = DBHdatarm$Treatment)
abline(0, 1)
plot(log(DBHdatarm$DBH.year.6), y.hat, col = DBHdatarm$InGap)
abline(0, 1)
```
#-------------------------------------------------------------------------------------------



# other approaches
further remove marked points which gives large residuals in model1
```{r}
# marked points
DBHpositive[c("14", "198", "217", "96"), ]
DBHpositiverm = subset(DBHpositive, !(Tree.Number == 46 | Tree.Number == 225 | Tree.Number == 230 | Tree.Number == 71))

# visualise marked points
plot(log(DBHpositiverm$DBH.year.0), log(DBHpositiverm$Tree.growth), type = "p", xlim = c(3.0,4.6), ylim = c(-3,3))
points(log(DBHpositive[c("14", "198", "217", "96"), "DBH.year.0"]), log(DBHpositive[c("14", "198", "217", "96"), "Tree.growth"]), col = "red")
```
```{r}
model1.1 = glm(log(Tree.growth)~log(DBH.year.0)+Treatment+Large+log(DBH.year.0):Large+Treatment:Large, data = DBHpositiverm)
summary(model1.1)

tss = sum((log(DBHpositiverm$Tree.growth)-mean(log(DBHpositiverm$Tree.growth)))^2) # TSS
rss = sum((log(DBHpositiverm$Tree.growth)-model1.1$fitted.values)^2) # RSS
(tss-rss)/tss # R^2 = 0.3805371

plot(model1.1)
plot(log(DBHpositiverm$Tree.growth), model1.1$fitted.values, type = "p", col = DBHpositiverm$Treatment)
abline(0, 1)
```

boxcox transformation on DBHpositive$Tree.growth
```{r}
b = boxcox(DBHpositive$Tree.growth~1, lambda = seq(0, 0.5, 0.05))
lambda = b$x[which.max(b$y)]
DBHpositive$Tree.growth.boxcox = (DBHpositive$Tree.growth^lambda-1)/lambda
```
```{r}
model1.2 = glm(Tree.growth.boxcox~DBH.year.0+Treatment+Large+DBH.year.0:Large+Treatment:Large, data = DBHpositive)
summary(model1.2)

tss = sum((DBHpositive$Tree.growth.boxcox-mean(DBHpositive$Tree.growth.boxcox))^2) # TSS
rss = sum((DBHpositive$Tree.growth.boxcox-model1.2$fitted.values)^2) # RSS
(tss-rss)/tss # R^2 = 0.3642381

plot(model1.2)
plot(DBHpositive$Tree.growth.boxcox, model1.2$fitted.values, type = "p", col = DBHpositive$Treatment, main = "Fitted vs Actual (color Treatment)")
legend("topleft", legend = c("Control", "T1", "T2"), fill = c("black", "red", "green"))
abline(0, 1)
plot(DBHpositive$Tree.growth.boxcox, model1.2$fitted.values, type = "p", col = DBHpositive$Large, main = "Fitted vs Actual (color Large)")
legend("topleft", legend = c("not-large", "large"), fill = c("black", "red"))
abline(0, 1)
```

glm with gamma link for DBHpositiverm$Tree.growth
```{r}
# DBHpositive$Tree.growth is close to gamma distribution with a diverging tail
descdist(DBHpositive$Tree.growth)

fitdis3 = fitdist(DBHpositive$Tree.growth, "gamma")
summary(fitdis3)
par(mfrow = c(2, 2))
denscomp(fitdis3, legendtext = "Gamma")
qqcomp(fitdis3, legendtext = "Gamma")
cdfcomp(fitdis3, legendtext = "Gamma")
ppcomp(fitdis3, legendtext = "Gamma")
```
```{r}
# remove large outliers to exclude the diverging tail shown in the section above
DBHpositiverm = DBHpositive[DBHpositive$Tree.growth <= 10 & DBHpositive$DBH.year.0 < 80, ]

# best model derived from step(Tree.growth~DBH.year.0*Treatment*Large)
model1.3 = glm(Tree.growth~DBH.year.0+Treatment+Large+DBH.year.0:Treatment+DBH.year.0:Large, family = "Gamma", data = DBHpositiverm)
summary(model1.3)

tss = sum((DBHpositiverm$Tree.growth-mean(DBHpositiverm$Tree.growth))^2) # TSS
rss = sum((DBHpositiverm$Tree.growth-model1.3$fitted.values)^2) # RSS
(tss-rss)/tss # R^2 = 0.4002613

plot(model1.3)
plot(DBHpositiverm$Tree.growth, model1.3$fitted.values, type = "p", col = DBHpositiverm$Treatment)
abline(0, 1)
```

mixed effect model
```{r}
DBHpositiverm = DBHpositive[DBHpositive$Tree.growth <= 10 & DBHpositive$DBH.year.0 < 80, ]

model1.4 = lmer(Tree.growth~DBH.year.0+Treatment+Large+DBH.year.0:Large+Treatment:Large+(DBH.year.0+Treatment+Large+DBH.year.0:Large+Treatment:Large|Plot), data = DBHpositiverm)
summary(model1.4)
y.hat <- fitted(model1.4) # Fitted values
int.hat <- ranef(model1.4)[[1]][[1]] # Predicted intercepts
res.hat <- residuals(model1.4) # Estimated residuals
tss = sum((DBHpositiverm$Tree.growth-mean(DBHpositiverm$Tree.growth))^2) # TSS
rss = sum((DBHpositiverm$Tree.growth-y.hat)^2) # RSS
(tss-rss)/tss # R^2 = 0.5126373
```
```{r}
qqnorm(int.hat, main="Random Intercepts"); qqline(int.hat)
qqnorm(res.hat, main="Residuals"); qqline(res.hat)
plot(y.hat, res.hat, xlab="Fitted Values", ylab="Residuals")
abline(h=0, lty=2)
plot(DBHpositiverm$Tree.growth, y.hat, col = DBHpositiverm$Treatment)
abline(0, 1)
plot(DBHpositiverm$Tree.growth, y.hat, col = DBHpositiverm$Large)
abline(0, 1)
```





