---
title: "Browsing exclusion_modelling"
author: "Chayanit Jaroonsophonsak"
date: "27/07/2022"
output: word_document
---

Note: Only plots 7(T2), 9(T2), 10(T1), and 12(T1) have fenced quadrats. 
So, we are only able to test the effect of browsing exclusion between these 2 treatments, but not the control

Note2: even quadrats are fenced, odd quadrats are unfenced 

Current issues:
1. Diversity Modelling for (Y3-6) & (Y0-6)
Linear model fit the data moderately good, but not great. Data seems to have lighter tails than Normal distribution. Want to try using Beta GLM but don't know how to transform the data into between [0,1] 
Tried using percentage change instead, but also had to solve the negative values issue. After adding a constant throughout, some points become greater than 1 -> still can't fit beta

# Import data
```{r}
data=read.csv("try_1.csv")
head(data)
focused_data=data[data$Plot_number %in% c(7,9,10,12),] #only retain data from plots 7,9,10,12

Y0_data=focused_data[focused_data$Year_monitoring==0,]
Y3_data=focused_data[focused_data$Year_monitoring==3,]
Y6_data=focused_data[focused_data$Year_monitoring==6,]
```

# Species richness
```{r}
#calculate species richness in each quadrat each year 
library(dplyr)
Y0_rich=c()
for (i in c(7,9,10,12))
  {for (j in seq(1,32,1))
    {dataset=Y0_data[Y0_data$Plot_number==i & Y0_data$Quadrat_number==j,]
    count_scientifname <- dataset %>% count(Scientific.Name)
    a=nrow(count_scientifname)
    Y0_rich=append(Y0_rich,a)
  }
}
Y0_rich

Y3_rich=c()
for (i in c(7,9,10,12))
  {for (j in seq(1,32,1))
    {dataset=Y3_data[Y3_data$Plot_number==i & Y3_data$Quadrat_number==j,]
    count_scientifname <- dataset %>% count(Scientific.Name)
    a=nrow(count_scientifname)
    Y3_rich=append(Y3_rich,a)
  }
}
Y3_rich

Y6_rich=c()
for (i in c(7,9,10,12))
  {for (j in seq(1,32,1))
    {dataset=Y6_data[Y6_data$Plot_number==i & Y6_data$Quadrat_number==j,]
    count_scientifname <- dataset %>% count(Scientific.Name)
    a=nrow(count_scientifname) 
    Y6_rich=append(Y6_rich,a)
  }
}
Y6_rich

#Calculate changes in species richness
Y03_rich=Y3_rich-Y0_rich 
Y03_rich 

Y36_rich=Y6_rich-Y3_rich
Y36_rich 

Y06_rich=Y6_rich-Y0_rich
Y06_rich

#Generate a dataframe for the changes in species richness 
Time=rep(c('0-3','3-6','0-6'),c(128,128,128))
Plot_number=rep(rep(c(7,9,10,12),c(32,32,32,32)),3)
Treatment=rep(rep(c("Radial","Gap"),c(64,64)),3)
Quadrat_number=rep(seq(1,32,1),12)
Quadrat_fenced=rep(c('False','True'),192)
Richness_change=c(Y03_rich,Y36_rich, Y06_rich)
rich_df=data.frame(Time, Plot_number, Treatment, Quadrat_number, Quadrat_fenced, Richness_change)
```

# H: Removal of browsing will cause a greater increase in understorey species richness in T1 compared to T2 
## Visualisation
```{r}
#Year 0 to 3
rich_fenced_03=rich_df[rich_df$Time=="0-3" & Quadrat_fenced=="True",]
plot(rich_fenced_03$Richness_change, xlab="16 fenced quadrats for each plot: 7, 9, 10, 12", ylab="change in species richness", main="Change in species richness from Y0 to Y3 after browsing exclusion", col=factor(rich_fenced_03$Treatment))
abline(h=0, lty=2)
legend("bottomright", legend = paste(c("T1","T2")), col=c("black","red"), pch=1)

boxplot(rich_fenced_03$Richness_change~rich_fenced_03$Treatment, xlab="Treatment", ylab="Change in species richness", main="Change in species richness from Y0 to Y3 after browsing exclusion")

hist(rich_fenced_03$Richness_change)

#Year 3 to 6
rich_fenced_36=rich_df[rich_df$Time=="3-6" & Quadrat_fenced=="True",]
plot(rich_fenced_36$Richness_change, xlab="16 fenced quadrats for each plot: 7, 9, 10, 12", ylab="change in species richness", main="Change in species richness from Y3 to Y6 after browsing exclusion", col=factor(rich_fenced_36$Treatment))
abline(h=0, lty=2)
legend("bottomright", legend = paste(c("T1","T2")), col=c("black","red"), pch=1)

boxplot(rich_fenced_36$Richness_change~rich_fenced_36$Treatment, xlab="Treatment", ylab="Change in species richness", main="Change in species richness from Y3 to Y6 after browsing exclusion")

hist(rich_fenced_36$Richness_change)

#Year 0 to 6
rich_fenced_06=rich_df[rich_df$Time=="0-6" & Quadrat_fenced=="True",]
plot(rich_fenced_06$Richness_change, xlab="16 fenced quadrats for each plot: 7, 9, 10, 12", ylab="change in species richness", main="Change in species richness from Y0 to Y6 after browsing exclusion", col=factor(rich_fenced_06$Treatment))
abline(h=0, lty=2)
legend("bottomright", legend = paste(c("T1","T2")), col=c("black","red"), pch=1)

boxplot(rich_fenced_06$Richness_change~rich_fenced_06$Treatment, xlab="Treatment", ylab="Change in species richness", main="Change in species richness from Y0 to Y6 after browsing exclusion")

hist(rich_fenced_06$Richness_change)
```

## Modelling
### Year 0 to 3
```{r}
#Linear model
rich_03_lm=lm(Richness_change~Treatment, rich_fenced_03)
summary(rich_03_lm)
plot(rich_03_lm)
rich_03_lm2=lm(Richness_change~1, rich_fenced_03)
anova(rich_03_lm2,rich_03_lm)
```

```{r}
#Poisson GLM
#manipulate data by adding a constant to every y, so that they are all non-negative
a=min(rich_fenced_03$Richness_change)
rich_fenced_03$Richness_change_pos=rich_fenced_03$Richness_change + abs(a)
hist(rich_fenced_03$Richness_change_pos)

#visualisation and test if the data is poisson
hist(rich_fenced_03$Richness_change_pos)
library(fitdistrplus)
poisson.model1=fitdist(rich_fenced_03$Richness_change_pos,"pois",method=c("mle"))
poisson.model1
poisson.lambda=9.984375

plot(poisson.model1)

n=length(rich_fenced_03$Richness_change_pos)
samp.pct <- (1:n-0.5)/n
qqplot(x=qpois(samp.pct,poisson.lambda), y=sort(rich_fenced_03$Richness_change_pos), xlab = 'Theoretical quantiles', ylab = 'Empirical quantiles', main='Q-Q plot Poisson')
abline(0,1,col="red")


#fit the model
rich_03_pois=glm(Richness_change_pos~Treatment, rich_fenced_03, family = poisson)
summary(rich_03_pois)
#goodness of fit
qchisq(0.95,62)
library(msme)
P__disp(rich_03_pois)
#not adequate
```

```{r}
#Negative Binomial GLM
#Test if the data is Negative Binomial 
nbinom.model1=fitdist(rich_fenced_03$Richness_change_pos,"nbinom",method=c("mle"))
nbinom.model1
size.nbinom=15.299430		#first parameter	
mu.nbinom=9.984064
prob.nbinom=size.nbinom/(size.nbinom+mu.nbinom) 
prob.nbinom 

plot(nbinom.model1)
qqplot(x=qnbinom(samp.pct,size.nbinom,prob.nbinom), y=sort(rich_fenced_03$Richness_change_pos), xlab = 'Theoretical quantiles', ylab = 'Empirical quantiles', main='Q-Q plot Negative Binomial')
abline(0,1,col="red")

#fit the model
rich_03_nb=glm.nb(Richness_change_pos~Treatment, rich_fenced_03)
summary(rich_03_nb)
#goodness of fit
qchisq(0.95,62)
#Adequate
```

#Year 3 to 6
```{r}
#Linear model
rich_36_lm=lm(Richness_change~Treatment, rich_fenced_36)
summary(rich_36_lm)
plot(rich_36_lm)

rich_36_lm2=lm(Richness_change~1, rich_fenced_36)
anova(rich_36_lm2,rich_36_lm)
```

```{r}
#Poisson GLM
#manipulate data by adding a constant to every y, so that they are all non-negative
b=min(rich_fenced_36$Richness_change)
rich_fenced_36$Richness_change_pos=rich_fenced_36$Richness_change + abs(b)

#visualisation and test if the data is poisson
hist(rich_fenced_36$Richness_change_pos)
poisson.model2=fitdist(rich_fenced_36$Richness_change_pos,"pois",method=c("mle"))
poisson.model2
poisson.lambda2=9.125

plot(poisson.model2)

n2=length(rich_fenced_36$Richness_change_pos)
samp.pct2 <- (1:n2-0.5)/n2
qqplot(x=qpois(samp.pct2,poisson.lambda2), y=sort(rich_fenced_36$Richness_change_pos), xlab = 'Theoretical quantiles', ylab = 'Empirical quantiles', main='Q-Q plot Poisson')
abline(0,1,col="red")


#fit the model
rich_36_pois=glm(Richness_change_pos~Treatment, rich_fenced_36, family = poisson)
summary(rich_36_pois)
#goodness of fit
qchisq(0.95,62)
#not adequate
P__disp(rich_36_pois)
```
```{r}
#Negative Binomial GLM
#Test if the data is Negative Binomial 
nbinom.model2=fitdist(rich_fenced_36$Richness_change_pos,"nbinom",method=c("mle"))
nbinom.model2
size.nbinom2=6.357373		#first parameter	
mu.nbinom2=9.124743
prob.nbinom2=size.nbinom2/(size.nbinom2+mu.nbinom2) 
prob.nbinom2 

plot(nbinom.model2)
qqplot(x=qnbinom(samp.pct2,size.nbinom2,prob.nbinom2), y=sort(rich_fenced_36$Richness_change_pos), xlab = 'Theoretical quantiles', ylab = 'Empirical quantiles', main='Q-Q plot Negative Binomial')
abline(0,1,col="red")
#better than Poisson, but is it good enough (?)

#fit the model
rich_36_nb=glm.nb(Richness_change_pos~Treatment, rich_fenced_36)
summary(rich_36_nb)
#goodness of fit
qchisq(0.95,62)
#Adequate
```

#Year 0 to 6
```{r}
#Linear model
rich_06_lm=lm(Richness_change~Treatment, rich_fenced_06)
summary(rich_06_lm)
plot(rich_06_lm)

rich_06_lm2=lm(Richness_change~1, rich_fenced_06)
anova(rich_06_lm2,rich_06_lm)
```

```{r}
#Poisson GLM
#manipulate data by adding a constant to every y, so that they are all non-negative
c=min(rich_fenced_06$Richness_change)
rich_fenced_06$Richness_change_pos=rich_fenced_06$Richness_change + abs(c)

#visualisation and test if the data is poisson
hist(rich_fenced_06$Richness_change_pos)
poisson.model3=fitdist(rich_fenced_06$Richness_change_pos,"pois",method=c("mle"))
poisson.model3
poisson.lambda3=8.109375	

plot(poisson.model3)

n3=length(rich_fenced_06$Richness_change_pos)
samp.pct3 <- (1:n3-0.5)/n3
qqplot(x=qpois(samp.pct3,poisson.lambda3), y=sort(rich_fenced_06$Richness_change_pos), xlab = 'Theoretical quantiles', ylab = 'Empirical quantiles', main='Q-Q plot Poisson')
abline(0,1,col="red")

#fit the model
rich_06_pois=glm(Richness_change_pos~Treatment, rich_fenced_06, family = poisson)
summary(rich_06_pois)
#goodness of fit
qchisq(0.95,62)
#not adequate
P__disp(rich_06_pois)
```

```{r}
#Negative Binomial GLM
#Test if the data is Negative Binomial 
nbinom.model3=fitdist(rich_fenced_06$Richness_change_pos,"nbinom",method=c("mle"))
nbinom.model3
size.nbinom3=6.794577		#first parameter	
mu.nbinom3=8.110003	
prob.nbinom3=size.nbinom3/(size.nbinom3+mu.nbinom3) 
prob.nbinom3 

plot(nbinom.model3)
qqplot(x=qnbinom(samp.pct3,size.nbinom3,prob.nbinom3), y=sort(rich_fenced_06$Richness_change_pos), xlab = 'Theoretical quantiles', ylab = 'Empirical quantiles', main='Q-Q plot Negative Binomial')
abline(0,1,col="red")
#good

#fit the model
rich_06_nb=glm.nb(Richness_change_pos~Treatment, rich_fenced_06)
summary(rich_06_nb)
#goodness of fit
qchisq(0.95,62)
#Adequate
```

# H: Removal of Browsing will cause Species richness to be higher across all treatments (T1 & T2)
## Visualisation
```{r}
#Year 0 to 3
rich_03_df=rich_df[rich_df$Time=="0-3",]
plot(rich_03_df$Richness_change, xlab="32 quadrats for each plot: 7, 9, 10, 12", ylab="change in species richness", main="Change in species richness in fenced and unfenced quadrats from Y0 to Y3 ", col=factor(rich_03_df$Quadrat_fenced))
abline(h=0, lty=2)
legend("bottomright", legend = paste(c("unfenced","fenced")), col=c("black","red"), pch=1)

boxplot(rich_03_df$Richness_change~rich_03_df$Quadrat_fenced, xlab="Fenced quadrat?", ylab="Change in species richness", main="Change in species richness from Y0 to Y3")

library(ggplot2)
ggplot(rich_03_df, aes(x=Richness_change, color=Quadrat_fenced)) + geom_histogram(fill="white", position="dodge") + theme(legend.position="top")

#Year 3 to 6
rich_36_df=rich_df[rich_df$Time=="3-6",]
plot(rich_36_df$Richness_change, xlab="32 quadrats for each plot: 7, 9, 10, 12", ylab="change in species richness", main="Change in species richness in fenced and unfenced quadrats from Y3 to Y6 ", col=factor(rich_36_df$Quadrat_fenced))
abline(h=0, lty=2)
legend("bottomright", legend = paste(c("unfenced","fenced")), col=c("black","red"), pch=1)

boxplot(rich_36_df$Richness_change~rich_36_df$Quadrat_fenced, xlab="Fenced quadrat?", ylab="Change in species richness", main="Change in species richness from Y3 to Y6")

ggplot(rich_36_df, aes(x=Richness_change, color=Quadrat_fenced)) + geom_histogram(fill="white", position="dodge") + theme(legend.position="top")

#Year 0 to 6
rich_06_df=rich_df[rich_df$Time=="0-6",]
plot(rich_06_df$Richness_change, xlab="32 quadrats for each plot: 7, 9, 10, 12", ylab="change in species richness", main="Change in species richness in fenced and unfenced quadrats from Y0 to Y6 ", col=factor(rich_06_df$Quadrat_fenced))
abline(h=0, lty=2)
legend("bottomright", legend = paste(c("unfenced","fenced")), col=c("black","red"), pch=1)

boxplot(rich_06_df$Richness_change~rich_06_df$Quadrat_fenced, xlab="Fenced quadrat?", ylab="Change in species richness", main="Change in species richness from Y0 to Y6")

ggplot(rich_06_df, aes(x=Richness_change, color=Quadrat_fenced)) + geom_histogram(fill="white", position="dodge") + theme(legend.position="top")
```

## Modelling (tbc)
### Y0 to 3
```{r}
# Linear model
```


# Species diversity (Shannon's diversity index)
*First idea was to use (num of specific species)/(num of total unique species) but...
* It seems to make more sense(?) if p is calculated by score(species i)/(total score in the quadrat)
reference: http://www.bioline.org.br/pdf?st09043

```{r}
#Calculate Shannon's diversity index

Y0_div=c()
for (i in c(7,9,10,12))
  {for (j in seq(1,32,1))
    {dataset=Y0_data[Y0_data$Plot_number==i & Y0_data$Quadrat_number==j,]
    total_score=sum(dataset$Score)
    p_scientifname <- dataset %>%
    group_by(Scientific.Name) %>%
    summarise(s = sum(Score)) 
    p_scientifname$p=p_scientifname$s/total_score
    a=-sum((p_scientifname$p)*log(p_scientifname$p))
    Y0_div=append(Y0_div,a)
  }
}
Y0_div

Y3_div=c()
for (i in c(7,9,10,12))
  {for (j in seq(1,32,1))
    {dataset=Y3_data[Y3_data$Plot_number==i & Y3_data$Quadrat_number==j,]
    total_score=sum(dataset$Score)
    p_scientifname <- dataset %>%
    group_by(Scientific.Name) %>%
    summarise(s = sum(Score)) 
    p_scientifname$p=p_scientifname$s/total_score
    a=-sum((p_scientifname$p)*log(p_scientifname$p))
    Y3_div=append(Y3_div,a)
  }
}
Y3_div

Y6_div=c()
for (i in c(7,9,10,12))
  {for (j in seq(1,32,1))
    {dataset=Y6_data[Y6_data$Plot_number==i & Y6_data$Quadrat_number==j,]
    total_score=sum(dataset$Score)
    p_scientifname <- dataset %>%
    group_by(Scientific.Name) %>%
    summarise(s = sum(Score)) 
    p_scientifname$p=p_scientifname$s/total_score
    a=-sum((p_scientifname$p)*log(p_scientifname$p))
    Y6_div=append(Y6_div,a)
  }
}
Y6_div

# Calculate changes in species diversity
Y03_div=Y3_div-Y0_div
Y03_div

Y36_div=Y6_div-Y3_div
Y36_div

Y06_div=Y6_div-Y0_div
Y06_div

#Generate a dataframe for the changes in species diversity
Diversity_change=c(Y03_div,Y36_div,Y06_div)
div_df=data.frame(Time, Plot_number, Treatment, Quadrat_number, Quadrat_fenced, Diversity_change)

# Calculate percentage change in species diversity (to be used in Beta GLM later)
Y03_divpc=Y03_div/Y0_div
Y03_divpc

Y36_divpc=Y36_div/Y3_div
Y36_divpc

Y06_divpc=Y06_div/Y0_div
Y06_divpc

div_df$Diversity_pcchange=c(Y03_divpc,Y36_divpc,Y06_divpc)
```

# H: Removal of browsing will cause a greater increase in understorey species diversity in T1 as compared to T2 
## Visualisation
```{r}
#Year 0 to 3
div_fenced_03=div_df[div_df$Time=="0-3" & Quadrat_fenced=="True",]
plot(div_fenced_03$Diversity_change, xlab="16 fenced quadrats for each plot: 7, 9, 10, 12", ylab="change in species diversity", main="Change in species diversity from Y0 to Y3 after browsing exclusion", col=factor(div_fenced_03$Treatment))
abline(h=0, lty=2)
legend("bottomright", legend = paste(c("T1","T2")), col=c("black","red"), pch=1)

boxplot(div_fenced_03$Diversity_change~div_fenced_03$Treatment, xlab="Treatment", ylab="Change in species diversity", main="Change in species diversity from Y0 to Y3 after browsing exclusion")

plot(density(div_fenced_03$Diversity_change), main = "Probability density of the change in species diversity from Y0 to Y3", xlab="Change in species diversity")

#Year 3 to 6
div_fenced_36=div_df[div_df$Time=="3-6" & Quadrat_fenced=="True",]
plot(div_fenced_36$Diversity_change, xlab="16 fenced quadrats for each plot: 7, 9, 10, 12", ylab="change in species diversity", main="Change in species diversity from Y3 to Y6 after browsing exclusion", col=factor(div_fenced_36$Treatment))
abline(h=0, lty=2)
legend("bottomright", legend = paste(c("T1","T2")), col=c("black","red"), pch=1)

boxplot(div_fenced_36$Diversity_change~div_fenced_36$Treatment, xlab="Treatment", ylab="Change in species diversity", main="Change in species diversity from Y3 to Y6 after browsing exclusion")

plot(density(div_fenced_36$Diversity_change), main = "Probability density of the change in species diversity from Y3 to Y6", xlab="Change in species diversity")

#Year 0 to 6
div_fenced_06=div_df[div_df$Time=="0-6" & Quadrat_fenced=="True",]
plot(div_fenced_06$Diversity_change, xlab="16 fenced quadrats for each plot: 7, 9, 10, 12", ylab="change in species diversity", main="Change in species diversity from Y0 to Y6 after browsing exclusion", col=factor(div_fenced_06$Treatment))
abline(h=0, lty=2)
legend("bottomright", legend = paste(c("T1","T2")), col=c("black","red"), pch=1)

boxplot(div_fenced_06$Diversity_change~div_fenced_06$Treatment, xlab="Treatment", ylab="Change in species diversity", main="Change in species diversity from Y0 to Y6 after browsing exclusion")

plot(density(div_fenced_06$Diversity_change), main = "Probability density of the change in species diversity from Y0 to Y6", xlab="Change in species diversity")
```

## Modelling
### Year 0 to 3
```{r}
# Linear model
div_03_lm=lm(Diversity_change~Treatment, div_fenced_03)
summary(div_03_lm)
plot(div_03_lm)
div_03_lm2=lm(Diversity_change~1, div_fenced_03)
anova(div_03_lm2,div_03_lm)
#intercept only model is better
```

```{r}
# Gamma and Inverse Gaussian GLM
#manipulate data by adding a constant to every y, so that they are all positive
d=min(div_fenced_03$Diversity_change)-0.0001 #-0.0001 so that y is not 0 
div_fenced_03$Diversity_change_pos=div_fenced_03$Diversity_change + abs(d)

#visualisation and test if the data is gamma or inverse Gaussian
plot(density(div_fenced_03$Diversity_change_pos))
#Unlikely to be gamma cause it's not positively skewed
gamma.model=fitdist(div_fenced_03$Diversity_change_pos,"gamma",method=c("mle"))
gamma.model
plot(gamma.model)
#looking terrible

library(GeneralizedHyperbolic)
invgstrt <- nigFitStart(div_fenced_03$Diversity_change_pos, startValues = "FN")
invgstrt$paramStart
invg.model=fitdist(div_fenced_03$Diversity_change_pos,"invgauss",start = list(mean=1.3226404, dispersion=0.3821182, shape=1))
invg.model
plot(invg.model)
#looking terrible


#fit the model (Gamma)
div_03_gamma=glm(Diversity_change_pos~Treatment, div_fenced_03, family = Gamma)
summary(div_03_gamma)
#goodness of fit
qchisq(0.95,62)
#adequate but bad fit (?)

#So, choose linear model
```

```{r}
#Beta GLM (might try, but needs y to be b/w 0 and 1, so need percentage change) (tbc)
beta.model=fitdist(div_fenced_03$Diversity_change_pos,"beta",method=c("mle"))
beta.model
plot(beta.model)
```

### Year 3 to 6 (tbc) other distributions with lighter tails than Normal? (try Beta)
```{r}
# Linear model
div_36_lm=lm(Diversity_change~Treatment, div_fenced_36)
summary(div_36_lm)
plot(div_36_lm)
#data seems to have ligter tails than Normal
div_36_lm2=lm(Diversity_change~1, div_fenced_36)
anova(div_36_lm2,div_36_lm)
#intercept only model is better
```

```{r}
# Gamma GLM
#manipulate data by adding a constant to every y, so that they are all positive
e=min(div_fenced_36$Diversity_change)-0.0001 #-0.0001 so that y is not 0 
div_fenced_36$Diversity_change_pos=div_fenced_36$Diversity_change + abs(e)

#visualisation and test if the data is gamma
plot(density(div_fenced_36$Diversity_change_pos))

gamma.model2=fitdist(div_fenced_36$Diversity_change_pos,"gamma",method=c("mle"))
gamma.model2
plot(gamma.model2)
#looking alright except gamma might be to heavy-tailed for our data on the right 

#fit the model (Gamma)
div_36_gamma=glm(Diversity_change_pos~Treatment, div_fenced_36, family = Gamma)
plot(div_36_gamma)
summary(div_36_gamma)
#goodness of fit
qchisq(0.95,62)
#adequate 
```

```{r}
# Inverse Gaussian GLM
invgstrt <- nigFitStart(div_fenced_36$Diversity_change_pos, startValues = "FN")
invgstrt$paramStart
invg.model2=fitdist(div_fenced_36$Diversity_change_pos,"invgauss",start = list(mean=0.9684214, dispersion=0.4381066, shape=1))
invg.model2
plot(invg.model2)
#looking terrible
```

```{r}
# Beta GLM
# manipulate data by adding a constant to every y, so that they are all non-negative 
e2=min(div_fenced_36$Diversity_pcchange)-0.00001 
div_fenced_36$Diversity_pcchange_pos=div_fenced_36$Diversity_pcchange + abs(e2)
length(div_fenced_36[div_fenced_36$Diversity_pcchange_pos>1]) #1 value is greater than 1

div_36_beta=glm(Diversity_pcchange_pos~Treatment, div_fenced_36, family = Beta)
#error, can't fit the model
```

### Year 0 to 6 (tbc) (try Beta)
```{r}
# Linear model
div_06_lm=lm(Diversity_change~Treatment, div_fenced_06)
summary(div_06_lm)
plot(div_06_lm)
div_06_lm2=lm(Diversity_change~1, div_fenced_06)
anova(div_06_lm2,div_06_lm)
#intercept only model is better
```

```{r}
# Gamma GLM
#manipulate data by adding a constant to every y, so that they are all positive
f=min(div_fenced_06$Diversity_change)-0.000001 #-0.0001 so that y is not 0 
div_fenced_06$Diversity_change_pos=div_fenced_06$Diversity_change + abs(f)

#visualisation and test if the data is gamma
plot(density(div_fenced_06$Diversity_change_pos))

gamma.model3=fitdist(div_fenced_06$Diversity_change_pos,"gamma",method=c("mle"))
gamma.model3
plot(gamma.model3)
#looking alright except gamma might be to heavy-tailed for our data on the right 

#fit the model (Gamma)
div_06_gamma=glm(Diversity_change_pos~Treatment, div_fenced_06, family = Gamma)
summary(div_06_gamma)
#goodness of fit
qchisq(0.95,62)
#adequate 
```

```{r}
# Inverse Gaussian GLM
invgstrt <- nigFitStart(div_fenced_06$Diversity_change_pos, startValues = "FN")
invgstrt$paramStart
invg.model3=fitdist(div_fenced_36$Diversity_change_pos,"invgauss",start = list(mean=0.8463209, dispersion=0.4788162, shape=1))
invg.model3
plot(invg.model3)
#looking terrible
```


# Relative abundance of species (Option1: Shannon's Equitability index = Evenness)
```{r}
# Calculate Shannon's Equitability index 
Y0_even=Y0_div/log(Y0_rich)
Y0_even

Y3_even=Y3_div/log(Y3_rich)
Y3_even

Y6_even=Y6_div/log(Y6_rich)
Y6_even

# Calculate the change in evenness
Y03_even=Y3_even-Y0_even
Y03_even

Y36_even=Y6_even-Y3_even
Y36_even 

#Generate a dataframe for the changes in evenness
Evenness_change=c(Y03_even,Y36_even)
even_df=data.frame(Time, Plot_number, Treatment, Quadrat_number, Quadrat_fenced, Evenness_change)
```

# Relative abundance of species (Option2: Braun-Blanquet Cover Abundance Scale = Abundance then compare between treatments)

???????????????????
H: Removal of browsing will cause a greater change in the relative abundances of species in the understorey community in T1 as compared to T2 
= test whether Relative Abundance [ A(T1)/A(T2) ] increases over time = the change in Relative Abundance is greater than 0 

H: Removal of browsing will cause a change in the relative abundances of species in the understorey community across all treatments (T1, T2)
= test whether Relative Abundance [ A(T1)/A(T2) ] changes over time = the change in Relative Abundance is not equal to 0

How do we decide WHICH quadrat relative to WHICH quadrat 
????????????????????

```{r} 
# Calculate abundance for each quadrat (sum of scores)
Y0_abun=c()
for (i in c(7,9,10,12))
  {for (j in seq(1,32,1))
    {dataset=Y0_data[Y0_data$Plot_number==i & Y0_data$Quadrat_number==j,]
    a=sum(dataset$Score)
    Y0_abun=append(Y0_abun,a)
  }
}
Y0_abun

Y3_abun=c()
for (i in c(7,9,10,12))
  {for (j in seq(1,32,1))
    {dataset=Y3_data[Y3_data$Plot_number==i & Y3_data$Quadrat_number==j,]
    a=sum(dataset$Score)
    Y3_abun=append(Y3_abun,a)
  }
}
Y3_abun

Y6_abun=c()
for (i in c(7,9,10,12))
  {for (j in seq(1,32,1))
    {dataset=Y6_data[Y6_data$Plot_number==i & Y6_data$Quadrat_number==j,]
    a=sum(dataset$Score)
    Y6_abun=append(Y6_abun,a)
  }
}
Y6_abun

# Calculate relative abundance [ A(T1)/A(T2) ]
# 1. Plot 10(T1) relative to Plot 9(T2) & Plot 12(T1) relative to Plot 7(T2)
Y0_rabun1=Y0_abun[65:96]/Y0_abun[33:64]
Y0_rabun2=Y0_abun[97:128]/Y0_abun[1:32]

Y3_rabun1=Y3_abun[65:96]/Y3_abun[33:64]
Y3_rabun2=Y3_abun[97:128]/Y3_abun[1:32]

Y6_rabun1=Y6_abun[65:96]/Y6_abun[33:64]
Y6_rabun2=Y6_abun[97:128]/Y6_abun[1:32]

# 2. Plot 10(T1) relative to Plot 7(T2) & Plot 12(T1) relative to Plot 9(T2)
Y0_rabun3=Y0_abun[65:96]/Y0_abun[1:32]
Y0_rabun4=Y0_abun[97:128]/Y0_abun[33:64]

Y3_rabun3=Y3_abun[65:96]/Y3_abun[1:32]
Y3_rabun4=Y3_abun[97:128]/Y3_abun[33:64]

Y6_rabun3=Y6_abun[65:96]/Y6_abun[1:32]
Y6_rabun4=Y6_abun[97:128]/Y6_abun[33:64]

# 3. Combined
Y0_rabun=c(Y0_rabun1, Y0_rabun2, Y0_rabun3, Y0_rabun4)
Y3_rabun=c(Y3_rabun1, Y3_rabun2, Y3_rabun3, Y3_rabun4)
Y6_rabun=c(Y6_rabun1, Y6_rabun2, Y6_rabun3, Y6_rabun4)

Y0_rabun
Y3_rabun
Y6_rabun

# Calculate changes in relative abundance
Y03_rabun=Y3_rabun-Y0_rabun
Y36_rabun=Y6_rabun-Y3_rabun

Y03_rabun
Y36_rabun
```
