---
title: "Browsing_richness_modelling"
author: "Chayanit Jaroonsophonsak"
output: word_document
---

# Import and clean data
```{r}
library(glmtoolbox)
library(rstudioapi)
library(pROC)

setwd(dirname(getActiveDocumentContext()$path))
rich_df = read.csv("../dataset/tables/species_richness_quadrats.csv")

# Extract data with relevant plots 
relevant_plots = c(7,9,10,12)
rich_df = rich_df[rich_df$Plot.Number %in% relevant_plots,]
rich_df

# Convert some variables into different data types 
str(rich_df)

rich_df$Plot.Number = factor(rich_df$Plot.Number)
rich_df$Quadrat.Number = factor(rich_df$Quadrat.Number)

rich_df$Gap <- rich_df$Gap=='True'
rich_df$Fenced <- rich_df$Fenced=='True'

# Remove null rows
rich_df <- na.omit(rich_df)
```

# Generate new variables, indicating whether or not species richness has increased in each quadrat at different periods (1 = increased, 0 = did not increase) 
```{r}
rich_df$Y0.Y3.Increase <- rich_df$X3 > rich_df$X0
rich_df$Y0.Y3.Increase <- as.numeric(rich_df$Y0.Y3.Increase)

rich_df$Y3.Y6.Increase <- rich_df$X6 > rich_df$X3
rich_df$Y3.Y6.Increase <- as.numeric(rich_df$Y3.Y6.Increase)

rich_df$Y0.Y6.Increase <- rich_df$X6 > rich_df$X0
rich_df$Y0.Y6.Increase <- as.numeric(rich_df$Y0.Y6.Increase)
```

# A function to generate confusion matrix and diagnose binomial response
```{r}
confusion <- function(model, y, data, threshold) {
  
  y_hat <- predict(model, data, type='response') > threshold
  conf <- matrix(0, nrow=2, ncol=2)
  rownames(conf) <- c('Predict 1', 'Predict 0')
  colnames(conf) <- c('Observe 1', 'Observe 0')
  
  conf[1,1] <- sum(y==1 & y_hat==1)
  conf[1,2] <- sum(y==0 & y_hat==1)
  conf[2,1] <- sum(y==1 & y_hat==0)
  conf[2,2] <- sum(y==0 & y_hat==0)
  
  tp_rate <- conf[1,1] / sum(conf[1,]) # Sensitivity
  tn_rate <- conf[2,2] / sum(conf[2,]) # Specificity
  
  return(list(Confusion=conf, Sensitivity=tp_rate, Specificity=tn_rate))
}
```

# #######################################################################################
# Hypothesis 1: Removal of browsing will cause a greater increase
# in understorey species richness in gap treatment (T1) compared to radial treatment (T2)
# ---------------------------------------------------------------------------------------
# Hypothesis 2: 
# #######################################################################################

# ##### Year 0 to year 3 #####
# Visualisation

# Modelling
```{r}
# Model 1: predictors with all main & interaction effects 
b_rich_Y0_Y3_01 = glm(Y0.Y3.Increase ~ (Treatment + Gap + Fenced)^2, family='binomial', data=rich_df)
summary(b_rich_Y0_Y3_01)


# Model 2: Step-wise selection from Model 1 (AIC) 
b_rich_Y0_Y3_02 = step(b_rich_Y0_Y3_01, k=2)
summary(b_rich_Y0_Y3_02)
# TreatmentRadial:FencedTRUE is not significant at 5% significance level


#compare model 2 and 1
anova(b_rich_Y0_Y3_02, b_rich_Y0_Y3_01)
pchisq(0.27015,2)
# there's no significant evidence that model 1 is better than model 2 


# Model 3: only include main effects 
b_rich_Y0_Y3_03 = glm(Y0.Y3.Increase ~ Treatment + Gap + Fenced, family='binomial', data=rich_df)
summary(b_rich_Y0_Y3_03)


# Model 4: Step-wise selection from Model 3 (AIC)  
step(b_rich_Y0_Y3_03)
# does not give new model


# compare model 3 and 2 
anova(b_rich_Y0_Y3_03, b_rich_Y0_Y3_02)
pchisq(3.5637, 1)
# there's no significant evidence that model 2 is better than model 3
# However, model 2 has lower AIC than model 3 


# Model 2 Goodness of fit 
r <- roc(rich_df$Y0.Y3.Increase, predict(b_rich_Y0_Y3_02, rich_df, type = 'response'))
auc(r)
plot.roc(r)

c <- confusion(b_rich_Y0_Y3_02, rich_df$Y0.Y3.Increase, rich_df, 0.5)
c$Confusion
c$Sensitivity # True Positive Rate
c$Specificity

# Model 3 Goodness of fit 
r <- roc(rich_df$Y0.Y3.Increase, predict(b_rich_Y0_Y3_03, rich_df, type = 'response'))
auc(r)
plot.roc(r)

c <- confusion(b_rich_Y0_Y3_03, rich_df$Y0.Y3.Increase, rich_df, 0.5)
c$Confusion
c$Sensitivity # True Positive Rate
c$Specificity

# Model 2 seems to have better goodness of fit 
# Choose model 2 
```
# Interpretation
- The estimated odds ratio of increase in species richness from Year 0 to Year 3 decreases by 39.37% for those quadrats with Radial treatment compared to those with Gap treatment.
- Cannot reject H1 
```{r}
exp(-1.9845+1.4842)
1-exp(-1.9845+1.4842)
```

# Try limiting the dataset to just 'fenced' quadrats
# Only focus on fenced quadrats
```{r}
rich_h1_df = rich_df[rich_df$Fenced == 'TRUE',]
```

```{r}
# Model 1: Treatment + Gap + Treatment:Gap  
b_rich_Y0_Y3_try_01 = glm(Y0.Y3.Increase ~ (Treatment + Gap)^2, family='binomial', data=rich_h1_df)
summary(b_rich_Y0_Y3_try_01)
# None of predictors appears significant

# Model 2: Step-wise selection from Model 1 (AIC) 
b_rich_Y0_Y3_try_02 = step(b_rich_Y0_Y3_try_01)
summary(b_rich_Y0_Y3_try_02)
# model with only 1 predictor: Gap (insignificant)
# Treatment has an insignificant effect on the response at 5% significance level 

# Model 3: Treatment as the only predictor
b_rich_Y0_Y3_try_03 = glm(Y0.Y3.Increase ~ Treatment, family='binomial', data=rich_h1_df)
summary(b_rich_Y0_Y3_try_03)
# Treatment has an insignificant effect on the response at 5% significance level
# higher AIC than model 2

# compare model 2 with model 1
anova(b_rich_Y0_Y3_try_02,b_rich_Y0_Y3_try_01)
pchisq(1.0711, 2)
# p-value >> 0.05
# insufficient evidence to suggest that model 1 is better

# cannot incorporate polynomial terms because both predictors are factors with values 0 & 1

# Model 2: Goodness of fit test 
c <- confusion(b_rich_Y0_Y3_try_02, rich_h1_df$Y0.Y3.Increase, rich_h1_df, 0.5)

c$Confusion
c$Sensitivity # True Positive Rate
c$Specificity # True Negative Rate

r <- roc(rich_h1_df$Y0.Y3.Increase, predict(b_rich_Y0_Y3_try_02, rich_h1_df, type = 'response'))
auc(r)
plot.roc(r)
# Not so good 

# Model 2 is best here, but it doesn't seem to be good enough 
# There's insufficient evidence to suggest that removal of browsing has caused a greater increase in understorey species richness in T1 compared to T2 from Year 0 to Year 3  
```

# try modelling using Poisson GLM & linear model
```{r}
hist(rich_df$X3)

library(fitdistrplus)
descdist(rich_df$X3)

fitdis1 = fitdist(rich_df$X3, "norm")
fitdis2 = fitdist(rich_df$X3, "pois")
summary(fitdis1); summary(fitdis2)
poisson.lambda = 13.98425	 
plot(fitdis1)
plot(fitdis2)

n=length(rich_df$X3)
samp.pct <- (1:n-0.5)/n
qqplot(x=qpois(samp.pct,poisson.lambda), y=sort(rich_df$X3), xlab = 'Theoretical quantiles', ylab = 'Empirical quantiles', main='Q-Q plot Poisson')
abline(0,1,col="red")

b_rich_pois_Y0_Y3_01 = glm(X3 ~ X0 + (Treatment + Gap + Fenced)^2, data = rich_df, family = 'poisson')
summary(b_rich_pois_Y0_Y3_01)

b_rich_pois_Y0_Y3_02=step(b_rich_pois_Y0_Y3_01)
summary(b_rich_pois_Y0_Y3_02)

anova(b_rich_pois_Y0_Y3_02,b_rich_pois_Y0_Y3_01)
pchisq(3.0705,4)
plot(b_rich_pois_Y0_Y3_02)

b_rich_pois_Y0_Y3_03 = glm(X3 ~ X0 + Treatment * Fenced + Gap, data = rich_df, family = 'poisson')
summary(b_rich_pois_Y0_Y3_03)

anova(b_rich_pois_Y0_Y3_02,b_rich_pois_Y0_Y3_03)
pchisq(1.9724,2)

b_rich_pois_Y0_Y3_04 = glm(X3 ~ X0 + (Treatment + Gap + Fenced)^2, data = rich_df, family = 'gaussian')
summary(b_rich_pois_Y0_Y3_04)
plot(b_rich_pois_Y0_Y3_04)

b_rich_pois_Y0_Y3_05 = step(b_rich_pois_Y0_Y3_04)
summary(b_rich_pois_Y0_Y3_05)
plot(b_rich_pois_Y0_Y3_05)
# same predictors as model2, but model 2 yields lower AIC

b_rich_pois_Y0_Y3_06 = glm(X3 ~ X0 + Treatment + Gap, data = rich_df, family = 'gaussian')
summary(b_rich_pois_Y0_Y3_06)
plot(b_rich_pois_Y0_Y3_06)
```
