---
title: "Browsing_diversity_modelling"
author: "Chayanit Jaroonsophonsak"
date: '2022-08-31'
output: word_document
---

# Import data
```{r}
div_df = read.csv("/Users/tat/Documents/Data Science Sem 3/Project 2/Github/Browsing/data_generated/ShannonDI_calculation.csv")
head(div_df)
```

# H: Removal of browsing will cause a greater increase in understorey species diversity in T1 as compared to T2 
## Visualisation
```{r}
#Year 0 to 3
div_fenced_03=div_df[div_df$Time=="0-3" & div_df$Quadrat_fenced=="True",]
plot(div_fenced_03$Diversity_change, xlab="16 fenced quadrats for each plot: 7, 9, 10, 12", ylab="change in species diversity", main="Change in species diversity from Y0 to Y3 after browsing exclusion", col=factor(div_fenced_03$Treatment))
abline(h=0, lty=2)
legend("bottomright", legend = paste(c("T1","T2")), col=c("black","red"), pch=1)

boxplot(div_fenced_03$Diversity_change~div_fenced_03$Treatment, xlab="Treatment", ylab="Change in species diversity", main="Change in species diversity from Y0 to Y3 after browsing exclusion")

plot(density(div_fenced_03$Diversity_change), main = "Probability density of the change in species diversity from Y0 to Y3", xlab="Change in species diversity")

#Year 3 to 6
div_fenced_36=div_df[div_df$Time=="3-6" & div_df$Quadrat_fenced=="True",]
plot(div_fenced_36$Diversity_change, xlab="16 fenced quadrats for each plot: 7, 9, 10, 12", ylab="change in species diversity", main="Change in species diversity from Y3 to Y6 after browsing exclusion", col=factor(div_fenced_36$Treatment))
abline(h=0, lty=2)
legend("bottomright", legend = paste(c("T1","T2")), col=c("black","red"), pch=1)

boxplot(div_fenced_36$Diversity_change~div_fenced_36$Treatment, xlab="Treatment", ylab="Change in species diversity", main="Change in species diversity from Y3 to Y6 after browsing exclusion")

plot(density(div_fenced_36$Diversity_change), main = "Probability density of the change in species diversity from Y3 to Y6", xlab="Change in species diversity")

#Year 0 to 6
div_fenced_06=div_df[div_df$Time=="0-6" & div_df$Quadrat_fenced=="True",]
plot(div_fenced_06$Diversity_change, xlab="16 fenced quadrats for each plot: 7, 9, 10, 12", ylab="change in species diversity", main="Change in species diversity from Y0 to Y6 after browsing exclusion", col=factor(div_fenced_06$Treatment))
abline(h=0, lty=2)
legend("bottomright", legend = paste(c("T1","T2")), col=c("black","red"), pch=1)

boxplot(div_fenced_06$Diversity_change~div_fenced_06$Treatment, xlab="Treatment", ylab="Change in species diversity", main="Change in species diversity from Y0 to Y6 after browsing exclusion")

plot(density(div_fenced_06$Diversity_change), main = "Probability density of the change in species diversity from Y0 to Y6", xlab="Change in species diversity")
```

## Modelling
### Year 0 to 3 (Choose linear model*)
```{r}
# Linear model
div_03_lm_a=glm(Diversity_change~Treatment+Quadrat_gap, div_fenced_03,family=gaussian)
summary(div_03_lm_a)
par(mfrow = c(2, 2))
plot(div_03_lm_a)
div_03_lm_b=lm(Diversity_change~Quadrat_gap, div_fenced_03)
summary(div_03_lm_b)
anova(div_03_lm_b,div_03_lm_a)
#there's an evidence that Treatment is significant (i.e. gap treatment has more positive impact on species diversity)
#but not sure if the fit is good enough (data has lighter left tail) 
```

```{r}
# Gamma and Inverse Gaussian GLM
#manipulate data by adding a constant to every y, so that they are all positive
d=min(div_fenced_03$Diversity_change)-0.0001 #-0.0001 so that y is not 0 
div_fenced_03$Diversity_change_pos=div_fenced_03$Diversity_change + abs(d)

#visualisation and test if the data is gamma or inverse Gaussian
plot(density(div_fenced_03$Diversity_change_pos))
#Unlikely to be gamma cause it's not positively skewed
gamma.model=fitdist(div_fenced_03$Diversity_change_pos,"gamma",method=c("mle"))
gamma.model
plot(gamma.model)
#looking terrible


#fit the model (Gamma)
div_03_gamma_a=glm(Diversity_change_pos~Treatment+Quadrat_gap, div_fenced_03, family = Gamma)
summary(div_03_gamma_a)
div_03_gamma_b=glm(Diversity_change_pos~Treatment, div_fenced_03, family = Gamma)
summary(div_03_gamma_b)
#goodness of fit
qchisq(0.95,62)
#adequate 
glm.diag.plots((div_03_gamma_b))
#So, choose linear model
```
### try incorporate random effects
```{r}
library(lme4)
div_03_lmm_a=lmer(Diversity_change~Treatment+Quadrat_gap+(1|Plot_number/Quadrat_number), data=div_fenced_03, REML = FALSE)
#Error: number of levels of each grouping factor must be < number of observations

div_03_lmm_b=lmer(Diversity_change~Treatment+Quadrat_gap+(1|Plot_number), data=div_fenced_03, REML = FALSE)
summary(div_03_lmm_b)
coef(div_03_lmm_b)
#Plot_number has insignificant random effect
#Choose normal LM
```

### Year 3 to 6 (Choose linear model)
```{r}
# Linear model
div_36_lm_a=lm(Diversity_change~Treatment+Quadrat_gap, div_fenced_36)
summary(div_36_lm_a) #choose this
plot(div_36_lm_a)
#the fit looks moderately good

div_36_lm_b=lm(Diversity_change~Quadrat_gap, div_fenced_36)
summary(div_36_lm_b)
anova(div_36_lm_b,div_36_lm_a)
#Treatment is not significant
```

```{r}
# Gamma GLM
#manipulate data by adding a constant to every y, so that they are all positive
e=min(div_fenced_36$Diversity_change)-0.0001 #-0.0001 so that y is not 0 
div_fenced_36$Diversity_change_pos=div_fenced_36$Diversity_change + abs(e)

#visualisation and test if the data is gamma
plot(density(div_fenced_36$Diversity_change_pos))

gamma.model2=fitdist(div_fenced_36$Diversity_change_pos,"gamma",method=c("mle"))
gamma.model2
plot(gamma.model2)
#looking quite bad

#fit the model (Gamma)
div_36_gamma_a=glm(Diversity_change_pos~Treatment+Quadrat_gap, div_fenced_36, family = Gamma)
glm.diag.plots(div_36_gamma_a)
summary(div_36_gamma_a)
#goodness of fit
qchisq(0.95,61)
#adequate, but the fit looks worse than LM

div_36_gamma_b=glm(Diversity_change_pos~Quadrat_gap, div_fenced_36, family = Gamma)
summary(div_36_gamma_b)
anova(div_36_gamma_b, div_36_gamma_a)
qchisq(0.95,1)
#Treatment is not significant
```

### try incorporate random effects
```{r}
div_36_lmm_a=lmer(Diversity_change~Treatment+Quadrat_gap+(1|Plot_number/Quadrat_number), data=div_fenced_36, REML = FALSE)
#Error: number of levels of each grouping factor must be < number of observations

div_36_lmm_b=lmer(Diversity_change~Treatment+Quadrat_gap+(1|Plot_number), data=div_fenced_36, REML = FALSE)
summary(div_36_lmm_b)
coef(div_36_lmm_b)
#Plot_number has insignificant random effect
#Choose normal LM
```

### Year 0 to 6 (Choose linear model)
```{r}
# Linear model
div_06_lm_a=glm(Diversity_change~Treatment + Quadrat_gap, div_fenced_06, family = gaussian)
summary(div_06_lm_a) #choose this 
plot(div_06_lm_a)
#quite good fit

div_06_lm_b=lm(Diversity_change~Quadrat_gap, div_fenced_06)
summary(div_06_lm_b)
anova(div_06_lm_b,div_06_lm_a)
#Treatment is not significant
```

```{r}
# Gamma GLM
#manipulate data by adding a constant to every y, so that they are all positive
f=min(div_fenced_06$Diversity_change)-0.000001 #-0.0001 so that y is not 0 
div_fenced_06$Diversity_change_pos=div_fenced_06$Diversity_change + abs(f)

#visualisation and test if the data is gamma
plot(density(div_fenced_06$Diversity_change_pos))

gamma.model3=fitdist(div_fenced_06$Diversity_change_pos,"gamma",method=c("mle"))
gamma.model3
plot(gamma.model3)
#looking quite bad 

#fit the model (Gamma)
div_06_gamma_a=glm(Diversity_change_pos~Treatment+Quadrat_gap, div_fenced_06, family = Gamma)
summary(div_06_gamma_a)
#goodness of fit
qchisq(0.95,61)
#adequate 
glm.diag.plots(div_06_gamma_a)

div_06_gamma_b=glm(Diversity_change_pos~Quadrat_gap, div_fenced_06, family = Gamma)
summary(div_06_gamma_b)
#adequate
#Treatment is not significant
#worse fit than LM
```

### try incorporate random effects
```{r}
div_06_lmm_a=lmer(Diversity_change~Treatment+Quadrat_gap+(1|Plot_number/Quadrat_number), data=div_fenced_06, REML = FALSE)
#Error: number of levels of each grouping factor must be < number of observations

div_06_lmm_b=lmer(Diversity_change~Treatment+Quadrat_gap+(1|Plot_number), data=div_fenced_06, REML = FALSE)
summary(div_06_lmm_b)
coef(div_06_lmm_b)
#Plot_number has insignificant random effect
#Choose normal LM
```

# H: Removal of Browsing will cause Species diversity to be higher across all treatments (T1 & T2)
## Visualisation
```{r}
#Year 0 to 3
div_03_df=div_df[div_df$Time=="0-3",]
plot(div_03_df$Diversity_change, xlab="32 quadrats for each plot: 7, 9, 10, 12", ylab="change in species diversity", main="Change in species diversity in fenced and unfenced quadrats from Y0 to Y3 ", col=factor(div_03_df$Quadrat_fenced))
abline(h=0, lty=2)
legend("bottomright", legend = paste(c("unfenced","fenced")), col=c("black","red"), pch=1)

boxplot(div_03_df$Diversity_change~div_03_df$Quadrat_fenced, xlab="Fenced quadrat?", ylab="Change in species diversity", main="Change in species diversity from Y0 to Y3")

library(ggplot2)
ggplot(div_03_df, aes(x=Diversity_change, color=Quadrat_fenced)) + geom_histogram(fill="white", position="dodge") + theme(legend.position="top")

#Year 3 to 6
div_36_df=div_df[div_df$Time=="3-6",]
plot(div_36_df$Diversity_change, xlab="32 quadrats for each plot: 7, 9, 10, 12", ylab="change in species diversity", main="Change in species diversity in fenced and unfenced quadrats from Y3 to Y6 ", col=factor(div_36_df$Quadrat_fenced))
abline(h=0, lty=2)
legend("bottomright", legend = paste(c("unfenced","fenced")), col=c("black","red"), pch=1)

boxplot(div_36_df$Diversity_change~div_36_df$Quadrat_fenced, xlab="Fenced quadrat?", ylab="Change in species diversity", main="Change in species diversity from Y3 to Y6")

ggplot(div_36_df, aes(x=Diversity_change, color=Quadrat_fenced)) + geom_histogram(fill="white", position="dodge") + theme(legend.position="top")

#Year 0 to 6
div_06_df=div_df[div_df$Time=="0-6",]
plot(div_06_df$Diversity_change, xlab="32 quadrats for each plot: 7, 9, 10, 12", ylab="change in species diversity", main="Change in species diversity in fenced and unfenced quadrats from Y0 to Y6 ", col=factor(div_06_df$Quadrat_fenced))
abline(h=0, lty=2)
legend("bottomright", legend = paste(c("unfenced","fenced")), col=c("black","red"), pch=1)

boxplot(div_06_df$Diversity_change~div_06_df$Quadrat_fenced, xlab="Fenced quadrat?", ylab="Change in species diversity", main="Change in species diversity from Y0 to Y6")

ggplot(div_06_df, aes(x=Diversity_change, color=Quadrat_fenced)) + geom_histogram(fill="white", position="dodge") + theme(legend.position="top")
```

## Modelling
### Year 0 to 3 (Choose Linear model)
```{r}
# Linear model
div_03_lm_h2a=lm(Diversity_change~Quadrat_fenced+Treatment+Quadrat_gap, div_03_df)
summary(div_03_lm_h2a)
plot(div_03_lm_h2a)
AIC(div_03_lm_h2a)

div_03_lm_h2b=lm(Diversity_change~Quadrat_fenced+Treatment, div_03_df)
summary(div_03_lm_h2b)
plot(div_03_lm_h2b) #choose this
AIC(div_03_lm_h2b)

div_03_lm_h2c=lm(Diversity_change~Treatment, div_03_df)
summary(div_03_lm_h2c)
plot(div_03_lm_h2c)
anova(div_03_lm_h2c,div_03_lm_h2b)
#Quadrat_fenced is not significant 
```

```{r}
# Gamma GLM
#manipulate data by adding a constant to every y, so that they are all positive
t=min(div_03_df$Diversity_change)-0.0001 #-0.0001 so that y is not 0 
div_03_df$Diversity_change_pos=div_03_df$Diversity_change + abs(t)

#visualisation and test if the data is gamma or inverse Gaussian
plot(density(div_03_df$Diversity_change_pos))
gamma.model4=fitdist(div_03_df$Diversity_change_pos,"gamma",method=c("mle"))
gamma.model4
plot(gamma.model4)
#looking worse than linear model

#fit the model (Gamma)
div_03_gamma_h2a=glm(Diversity_change_pos~Quadrat_fenced+Treatment+Quadrat_gap, div_03_df, family = Gamma)
summary(div_03_gamma_h2a)
#goodness of fit
qchisq(0.95,124)
#adequate

div_03_gamma_h2b=glm(Diversity_change_pos~Quadrat_fenced+Treatment, div_03_df, family = Gamma)
summary(div_03_gamma_h2b)
#goodness of fit
qchisq(0.95,125)
#adequate
glm.diag.plots(div_06_gamma_b)
#looking worse than LM

div_03_gamma_h2c=glm(Diversity_change_pos~Treatment, div_03_df, family = Gamma)
summary(div_03_gamma_h2c)
#goodness of fit
qchisq(0.95,126)
#adequate

anova(div_03_gamma_h2c,div_03_gamma_h2b)
#Quadrat_gap not significant
```

### try incorporate random effects
```{r}
div_06_lmm_h2a=lmer(Diversity_change~Quadrat_fenced+Treatment+(1|Plot_number/Quadrat_number), data=div_03_df, REML = FALSE)
#Error: number of levels of each grouping factor must be < number of observations

div_06_lmm_h2b=lmer(Diversity_change~Quadrat_fenced+Treatment+(1|Plot_number), data=div_03_df, REML = FALSE)
summary(div_06_lmm_h2b)
coef(div_06_lmm_h2b)
#Plot_number has insignificant random effect
#Choose normal LM
```

### Year 3 to 6 (Choose Linear model) 
```{r}
# Linear model
div_36_lm_h2a=lm(Diversity_change~Quadrat_fenced+Treatment+Quadrat_gap, div_36_df)
summary(div_36_lm_h2a) #choose this
plot(div_36_lm_h2a) #good fit
AIC(div_36_lm_h2a)

div_36_lm_h2b=lm(Diversity_change~Treatment+Quadrat_gap, div_36_df)
summary(div_36_lm_h2b)
plot(div_36_lm_h2b)
AIC(div_36_lm_h2b)

anova(div_36_lm_h2b,div_36_lm_h2a)
#Quadrat_fenced is not significant
```

```{r}
# Gamma GLM
#manipulate data by adding a constant to every y, so that they are all positive
s=min(div_36_df$Diversity_change)-0.0001 #-0.0001 so that y is not 0 
div_36_df$Diversity_change_pos=div_36_df$Diversity_change + abs(s)

#visualisation and test if the data is gamma
plot(density(div_36_df$Diversity_change_pos))
#Unlikely to be gamma cause it's not positively skewed
gamma.model5=fitdist(div_36_df$Diversity_change_pos,"gamma",method=c("mle"))
gamma.model5
plot(gamma.model5)
#looking worse than linear model

#fit the model (Gamma)
div_36_gamma_h2a=glm(Diversity_change_pos~Quadrat_fenced+Treatment+Quadrat_gap, div_36_df, family = Gamma)
summary(div_36_gamma_h2a)
#goodness of fit
qchisq(0.95,124)
#adequate

div_36_gamma_h2b=glm(Diversity_change_pos~Treatment+Quadrat_gap, div_36_df, family = Gamma)
summary(div_36_gamma_h2b)
#goodness of fit
qchisq(0.95,125)
#adequate
glm.diag.plots(div_36_gamma_h2b)

anova(div_36_gamma_h2b,div_36_gamma_h2a)
qchisq(0.95,1)
#Quadrat_gap not significant
#worse fit than LM
```

### try incorporate random effects
```{r}
div_36_lmm_h2a=lmer(Diversity_change~Quadrat_fenced+Treatment+Quadrat_gap+(1|Plot_number/Quadrat_number), data=div_36_df, REML = FALSE)
#Error: number of levels of each grouping factor must be < number of observations

div_36_lmm_h2b=lmer(Diversity_change~Quadrat_fenced+Treatment+Quadrat_gap+(1|Plot_number), data=div_36_df, REML = FALSE)
summary(div_36_lmm_h2b)
coef(div_36_lmm_h2b)
#Plot_number has some random effect (small)
library(RLRsim)
exactRLRT(div_36_lmm_h2b)
#Reference : https://www.ssc.wisc.edu/sscc/pubs/MM/MM_TestEffects.html#test-of-random-parameters
#The variance is not significant
#Choose normal LM (it also has lower AIC)
```

### Year 0 to 6 (Choose Linear model*)
```{r}
# Linear model
div_06_lm_h2a=lm(Diversity_change~Quadrat_fenced+Treatment+Quadrat_gap, div_06_df)
summary(div_06_lm_h2a)
plot(div_06_lm_h2a)

div_06_lm_h2b=lm(Diversity_change~Quadrat_fenced+Quadrat_gap, div_06_df)
summary(div_06_lm_h2b) #choose this
plot(div_06_lm_h2b) 
#data looks a bit heavier tailed than LM
AIC(div_06_lm_h2b)

div_06_lm_h2c=lm(Diversity_change~Quadrat_gap, div_06_df)
summary(div_06_lm_h2c)
anova(div_06_lm_h2c,div_06_lm_h2b)
#Quadrat_fenced is not significant
```

```{r}
# Gamma GLM
#manipulate data by adding a constant to every y, so that they are all positive
r=min(div_06_df$Diversity_change)-0.0001 #-0.0001 so that y is not 0 
div_06_df$Diversity_change_pos=div_06_df$Diversity_change + abs(r)

#visualisation and test if the data is gamma
plot(density(div_06_df$Diversity_change_pos))
gamma.model6=fitdist(div_06_df$Diversity_change_pos,"gamma",method=c("mle"))
gamma.model6
plot(gamma.model6)
#doesn't look that good

#fit the model (Gamma)
div_06_gamma_h2a=glm(Diversity_change_pos~Quadrat_fenced+Treatment+Quadrat_gap, div_06_df, family = Gamma)
summary(div_06_gamma_h2a)
#goodness of fit
qchisq(0.95,124)
#adequate

div_06_gamma_h2b=glm(Diversity_change_pos~Quadrat_fenced+Quadrat_gap, div_06_df, family = Gamma)
summary(div_06_gamma_h2b)
#goodness of fit
qchisq(0.95,125)
#adequate
glm.diag.plots(div_06_gamma_h2b)

div_06_gamma_h2c=glm(Diversity_change_pos~Quadrat_gap, div_06_df, family = Gamma)
summary(div_06_gamma_h2c)
#goodness of fit
qchisq(0.95,126)
#adequate

anova(div_06_gamma_h2c,div_06_gamma_h2b)
qchisq(0.95,1)
#Quadrat_gap not significant
#worse fit than LM
```

### try incorporate random effects
```{r}
div_06_lmm_h2a=lmer(Diversity_change~Quadrat_fenced+Quadrat_gap+(1|Plot_number/Quadrat_number), data=div_06_df, REML = FALSE)
#Error: number of levels of each grouping factor must be < number of observations

div_06_lmm_h2b=lmer(Diversity_change~Quadrat_fenced+Quadrat_gap+(1|Plot_number), data=div_06_df, REML = FALSE)
summary(div_06_lmm_h2b)
coef(div_06_lmm_h2b)
#Plot_number has some random effect (small)
exactRLRT(div_06_lmm_h2b)
#The variance is not significant
#Choose normal LM (it also has lower AIC)
```
