---
title: "Browsing calculations"
author: "Chayanit Jaroonsophonsak"
date: "27/07/2022"
output: word_document
---

Note: Only plots 7(T2), 9(T2), 10(T1), and 12(T1) have fenced quadrats. 
So, we are only able to test the effect of browsing exclusion between these 2 treatments, but not the control

Note2: even quadrats are fenced, odd quadrats are unfenced 

Current issue:
1. Y3-6 richness H1: heavy left tail NB GLM
2. Y0-6 richness H1: heavy left tail NB GLM
3. Y3-6 richness H2: heavy left tail NB GLM
4. Y0-6 richness H2: heavy left tail NB GLM
5. Y0-3 diversity H1: light left tail LM
6. Y0-6 diversity H2: heavy tails LM

# Import data
```{r}
data=read.csv("/Users/tat/Documents/Data Science Sem 3/Project 2/Data/joined_understorey_observations.csv")
head(data)
focused_data=data[data$Plot_number %in% c(7,9,10,12),]#only retain data from plots 7,9,10,12

Y0_data=focused_data[focused_data$Year_monitoring==0,]
Y3_data=focused_data[focused_data$Year_monitoring==3,]
Y6_data=focused_data[focused_data$Year_monitoring==6,]
```

# Species richness
```{r}
#calculate species richness in each quadrat each year 
library(dplyr)
Y0_rich=c()
for (i in c(7,9,10,12))
  {for (j in seq(1,32,1))
    {dataset=Y0_data[Y0_data$Plot_number==i & Y0_data$Quadrat_number==j,]
    count_scientifname <- dataset %>% count(Scientific.Name)
    a=nrow(count_scientifname)
    Y0_rich=append(Y0_rich,a)
  }
}
Y0_rich

Y3_rich=c()
for (i in c(7,9,10,12))
  {for (j in seq(1,32,1))
    {dataset=Y3_data[Y3_data$Plot_number==i & Y3_data$Quadrat_number==j,]
    count_scientifname <- dataset %>% count(Scientific.Name)
    a=nrow(count_scientifname)
    Y3_rich=append(Y3_rich,a)
  }
}
Y3_rich

Y6_rich=c()
for (i in c(7,9,10,12))
  {for (j in seq(1,32,1))
    {dataset=Y6_data[Y6_data$Plot_number==i & Y6_data$Quadrat_number==j,]
    count_scientifname <- dataset %>% count(Scientific.Name)
    a=nrow(count_scientifname) 
    Y6_rich=append(Y6_rich,a)
  }
}
Y6_rich

#Calculate changes in species richness
Y03_rich=Y3_rich-Y0_rich 
Y03_rich 

Y36_rich=Y6_rich-Y3_rich
Y36_rich 

Y06_rich=Y6_rich-Y0_rich
Y06_rich

#Extract in gap/ out of gap information
Quadrat_gap={}
for (i in c(7,9,10,12)){
  for (j in seq(1,32,1)){
    gap=toString(unique(Y0_data[Y0_data$Plot_number==i & Y0_data$Quadrat_number==j,]$Quadrat_gap))
    Quadrat_gap=append(Quadrat_gap,gap)
  }
}
Quadrat_gap
length(Quadrat_gap)


#Generate a dataframe for the changes in species richness 
Time=rep(c('0-3','3-6','0-6'),c(128,128,128))
Plot_number=rep(rep(c(7,9,10,12),c(32,32,32,32)),3)
Treatment=rep(rep(c("Radial","Gap"),c(64,64)),3)
Quadrat_number=rep(seq(1,32,1),12)
Quadrat_fenced=rep(c('False','True'),192)
Quadrat_gap=rep(Quadrat_gap,3)
Richness_change=c(Y03_rich,Y36_rich, Y06_rich)
rich_df=data.frame(Time, Plot_number, Treatment, Quadrat_number, Quadrat_fenced, Quadrat_gap, Richness_change)
write.csv(rich_df,"/Users/tat/Documents/Data Science Sem 3/Project 2/Github/data_generated/richness_calculation.csv", row.names = TRUE)
```

# Species diversity (Shannon's diversity index)
*First idea was to use (num of specific species)/(num of total unique species) but...
* It seems to make more sense(?) if p is calculated by score(species i)/(total score in the quadrat)
reference: http://www.bioline.org.br/pdf?st09043

```{r}
#Calculate Shannon's diversity index

Shannon <- function (vector) {
  
  # Define corresponding abundance value for each score
  Abundance = c(0.5, 3, 7.5, 15, 25, 35, 45, 55, 65, 75, 85) / 100
  names(Abundance) = c('0.5', '1', '2', '3', '4', '5', '6', '7', '8', '9', '10')
  
  scores = as.character(vector)
  
  # Vector containing abundance of each species
  a = Abundance[scores]

  # Vector containing abundance of each species as a proportion of total
  p = a/sum(a)
  
  # Apply formula for Shannon's Diversity Index
  return(-sum(log(p)*p))
}

Y0_div=c()
for (i in c(7,9,10,12))
  {for (j in seq(1,32,1))
    {dataset=Y0_data[Y0_data$Plot_number==i & Y0_data$Quadrat_number==j,]
    shannon_di=Shannon(dataset$Score)
    Y0_div=append(Y0_div,shannon_di)
  }
}
Y0_div

Y3_div=c()
for (i in c(7,9,10,12))
  {for (j in seq(1,32,1))
    {dataset=Y3_data[Y3_data$Plot_number==i & Y3_data$Quadrat_number==j,]
    shannon_di=Shannon(dataset$Score)
    Y3_div=append(Y3_div,shannon_di)
  }
}
Y3_div

Y6_div=c()
for (i in c(7,9,10,12))
  {for (j in seq(1,32,1))
    {dataset=Y6_data[Y6_data$Plot_number==i & Y6_data$Quadrat_number==j,]
    shannon_di=Shannon(dataset$Score)
    Y6_div=append(Y6_div, shannon_di)
  }
}
Y6_div

# Calculate changes in species diversity
Y03_div=Y3_div-Y0_div
Y03_div

Y36_div=Y6_div-Y3_div
Y36_div

Y06_div=Y6_div-Y0_div
Y06_div

#Generate a dataframe for the changes in species diversity
Diversity_change=c(Y03_div,Y36_div,Y06_div)
div_df=data.frame(Time, Plot_number, Treatment, Quadrat_number, Quadrat_fenced, Quadrat_gap, Diversity_change)
write.csv(div_df,"/Users/tat/Documents/Data Science Sem 3/Project 2/Github/data_generated/ShannonDI_calculation.csv", row.names = TRUE)
```

