---
title: "understorey_visualisation"
output: word_document
---

## Import data
```{r}
USdata=read.csv("~/Documents/Data Science sem 2/Group project 1/preprocessed data/understoreydata2020.csv")
head(USdata,10)
```

## Initial data exploration
```{r}
library(dplyr)
#which plots have fenced quadrats?
Which=USdata[USdata$Quadrat_fenced=='True',]
unique(Which$Plot_number)
#because of limited fundting, all quadrats in plot 4 and 11 are unfenced
#no more control plots for fenced quadrats

#which quadrats are fenced? 
Whichq_7=USdata[USdata$Quadrat_fenced=='True'& USdata$Plot_number==7,]
unique(Whichq_7$Quadrat_number)

Whichq_9=USdata[USdata$Quadrat_fenced=='True'& USdata$Plot_number==9,]
unique(Whichq_9$Quadrat_number)

Whichq_10=USdata[USdata$Quadrat_fenced=='True'& USdata$Plot_number==10,]
unique(Whichq_10$Quadrat_number)

Whichq_12=USdata[USdata$Quadrat_fenced=='True'& USdata$Plot_number==12,]
unique(Whichq_12$Quadrat_number)
```

## Count number of species in each quadrat each year
*For some specific quadrat in specific plot, the same species was recorded more than once with different score. Not sure what this means? Below, we will count all of them  

*Issue: TfN says 16 unfenced quadrants in each plot, but in plot 4 and 11, data from 32 quadrants were recorded as unfenced in year 3 and 6. If there’s no fenced quadrants, then which quadrants in these 2 plots should we use for analysis?
Solution: We will choose to analyse the 16 quadrants recorded in year 0 for plot 4 & 11

Unfenced
```{r}
Year0uf_count=c()
for (i in 1:12)
{if (i %in% c(1,2,3,5,6,8))
  {for (j in 1:16)
    {a=nrow(USdata[USdata$Year_monitoring==0 & USdata$Quadrat_fenced=='False' & USdata$Plot_number==i & USdata$Quadrat_number==j,])
    Year0uf_count=append(Year0uf_count,a)}
}
else 
  {for (j in seq(1,32,2))
    {a=nrow(USdata[USdata$Year_monitoring==0 & USdata$Quadrat_fenced=='False' & USdata$Plot_number==i & USdata$Quadrat_number==j,])
    Year0uf_count=append(Year0uf_count,a)}
  }
    }
Year0uf_count

Year3uf_count=c()
for (i in 1:12)
{if (i %in% c(1,2,3,5,6,8))
  {for (j in 1:16)
    {a=nrow(USdata[USdata$Year_monitoring==3 & USdata$Quadrat_fenced=='False' & USdata$Plot_number==i & USdata$Quadrat_number==j,])
    Year3uf_count=append(Year3uf_count,a)}
}
else 
  {for (j in seq(1,32,2))
    {a=nrow(USdata[USdata$Year_monitoring==3 & USdata$Quadrat_fenced=='False' & USdata$Plot_number==i & USdata$Quadrat_number==j,])
    Year3uf_count=append(Year3uf_count,a)}
  }
    }
Year3uf_count
#no data of plot 3 in year 3


Year6uf_count=c()
for (i in 1:12)
{if (i %in% c(1,2,3,5,6,8))
  {for (j in 1:16)
    {a=nrow(USdata[USdata$Year_monitoring==6 & USdata$Quadrat_fenced=='False' & USdata$Plot_number==i & USdata$Quadrat_number==j,])
    Year6uf_count=append(Year6uf_count,a)}
}
else 
  {for (j in seq(1,32,2))
    {a=nrow(USdata[USdata$Year_monitoring==6 & USdata$Quadrat_fenced=='False' & USdata$Plot_number==i & USdata$Quadrat_number==j,])
    Year6uf_count=append(Year6uf_count,a)}
  }
    }
Year6uf_count
```

Fenced (only plot 7, 9, 10, 12)
```{r}
Year0f_count=c()
for (i in c(7,9,10,12))
  {for (j in seq(2,32,2))
    {a=nrow(USdata[USdata$Year_monitoring==0 & USdata$Quadrat_fenced=='True' & USdata$Plot_number==i & USdata$Quadrat_number==j,])
    Year0f_count=append(Year0f_count,a)
    }
  }
Year0f_count

Year3f_count=c()
for (i in c(7,9,10,12))
  {for (j in seq(2,32,2))
    {a=nrow(USdata[USdata$Year_monitoring==3 & USdata$Quadrat_fenced=='True' & USdata$Plot_number==i & USdata$Quadrat_number==j,])
    Year3f_count=append(Year3f_count,a)
    }
  }
Year3f_count

Year6f_count=c()
for (i in c(7,9,10,12))
  {for (j in seq(2,32,2))
    {a=nrow(USdata[USdata$Year_monitoring==6 & USdata$Quadrat_fenced=='True' & USdata$Plot_number==i & USdata$Quadrat_number==j,])
    Year6f_count=append(Year6f_count,a)
    }
  }
Year6f_count
```

## Calculate the change in number of species
Unfenced
```{r}
uf_30=Year3uf_count[-(33:48)]-Year0uf_count[-(33:48)] #exclude plot 3 cause missing data in Y3
uf_30

uf_63=Year6uf_count[-(33:48)]-Year3uf_count[-(33:48)] #exclude plot 3 cause missing data in Y3 
uf_63

uf_60=Year6uf_count-Year0uf_count
uf_60
```

Fenced
```{r}
f_30=Year3f_count-Year0f_count
f_30

f_63=Year6f_count-Year3f_count
f_63

f_60=Year6f_count-Year0f_count
f_60
```

#Plots to compare the change in species number 
Note
T1: plot 1,8,10,12
T2: plot 3,5,7,9
Control: plot 2,4,6,11

Unfenced quadrats with different treatments
```{r}
plot(uf_30, xlab="16 unfenced quadrats for each plot, 1 to 12, excluding plot 3", ylab="change in number of species", main="Change in number of species from Y0 to Y3 (unfenced)", col=c(rep(4,16),rep(6,16),rep(6,16),rep(2,16),rep(6,16),rep(2,16),rep(4,16),rep(2,16),rep(4,16),rep(6,16),rep(4,16)), pch=18)
abline(h=0, lty=2)
legend("top", legend = paste(c("T1","T2","Control")), col = c(4,2,6), pch = 18)

treatments_30=as.data.frame(uf_30)
colnames(treatments_30)=c("change in number of species")
treatments_30$treatment=c(rep("T1",16),rep("Control",16),rep("Control",16),rep("T2",16),rep("Control",16),rep("T2",16),rep("T1",16),rep("T2",16),rep("T1",16),rep("Control",16),rep("T1",16))
treatments_30
boxplot(treatments_30$`change in number of species`~treatments_30$treatment, xlab="treatment", ylab="change in number of species", main="Change in number of species from Y0 to Y3 (unfenced)")

plot(uf_63, xlab="16 unfenced quadrats for each plot, 1 to 12, excluding plot 3", ylab="change in number of species", main="Change in number of species from Y3 to Y6 (unfenced)", col=c(rep(4,16),rep(6,16),rep(6,16),rep(2,16),rep(6,16),rep(2,16),rep(4,16),rep(2,16),rep(4,16),rep(6,16),rep(4,16)), pch=18)
abline(h=0, lty=2)
legend("bottom", legend = paste(c("T1","T2","Control")), col = c(4,2,6), pch = 18)

treatments_63=as.data.frame(uf_63)
colnames(treatments_63)=c("change in number of species")
treatments_63$treatment=c(rep("T1",16),rep("Control",16),rep("Control",16),rep("T2",16),rep("Control",16),rep("T2",16),rep("T1",16),rep("T2",16),rep("T1",16),rep("Control",16),rep("T1",16))
treatments_63
boxplot(treatments_63$`change in number of species`~treatments_63$treatment, xlab="treatment", ylab="change in number of species", main="Change in number of species from Y3 to Y6 (unfenced)")

plot(uf_60, xlab="16 unfenced quadrats for each plot, 1 to 12", ylab="change in number of species", main="Change in number of species from Y0 to Y6 (unfenced)", col=c(rep(4,16),rep(6,16),rep(2,16),rep(6,16),rep(2,16),rep(6,16),rep(2,16),rep(4,16),rep(2,16),rep(4,16),rep(6,16),rep(4,16)), pch=18)
abline(h=0, lty=2)
legend("bottomright", legend = paste(c("T1","T2","Control")), col = c(4,2,6), pch = 18)

treatments_60=as.data.frame(uf_60)
colnames(treatments_60)=c("change in number of species")
treatments_60$treatment=c(rep("T1",16),rep("Control",16),rep("T2",16),rep("Control",16),rep("T2",16),rep("Control",16),rep("T2",16),rep("T1",16),rep("T2",16),rep("T1",16),rep("Control",16),rep("T1",16))
treatments_60
boxplot(treatments_60$`change in number of species`~treatments_60$treatment, xlab="treatment", ylab="change in number of species", main="Change in number of species from Y0 to Y6 (unfenced)")
```

Fenced quadrats with different treatments
```{r}
plot(f_30, xlab="16 fenced quadrats for plot 7,9,10,12 respectively", ylab="change in number of species", main="Change in number of species from Y0 to Y3 (fenced)", col=c(rep(2,16),rep(2,16),rep(4,16),rep(4,16)), pch=18)
abline(h=0, lty=2)
legend("bottom", legend = paste(c("T1","T2","Control")), col = c(4,2,6), pch = 18)

treatments_30_f=as.data.frame(f_30)
colnames(treatments_30_f)=c("change in number of species")
treatments_30_f$treatment=c(rep("T2",32),rep("T1",32))
treatments_30_f
boxplot(treatments_30_f$`change in number of species`~treatments_30_f$treatment, xlab="treatment", ylab="change in number of species", main="Change in number of species from Y0 to Y3 (fenced)")

plot(f_63, xlab="16 fenced quadrats for plot 7,9,10,12 respectively", ylab="change in number of species", main="Change in number of species from Y3 to Y6 (fenced)", col=c(rep(2,16),rep(2,16),rep(4,16),rep(4,16)), pch=18)
abline(h=0, lty=2)
legend("topright", legend = paste(c("T1","T2","Control")), col = c(4,2,6), pch = 18)

treatments_63_f=as.data.frame(f_63)
colnames(treatments_63_f)=c("change in number of species")
treatments_63_f$treatment=c(rep("T2",32),rep("T1",32))
treatments_63_f
boxplot(treatments_63_f$`change in number of species`~treatments_63_f$treatment, xlab="treatment", ylab="change in number of species", main="Change in number of species from Y3 to Y6 (fenced)")

plot(f_60, xlab="16 fenced quadrats for plot 7,9,10,12 respectively", ylab="change in number of species", main="Change in number of species from Y0 to Y6 (fenced)", col=c(rep(2,16),rep(2,16),rep(4,16),rep(4,16)), pch=18)
abline(h=0, lty=2)
legend("bottom", legend = paste(c("T1","T2","Control")), col = c(4,2,6), pch = 18)

treatments_60_f=as.data.frame(f_60)
colnames(treatments_60_f)=c("change in number of species")
treatments_60_f$treatment=c(rep("T2",32),rep("T1",32))
treatments_60_f
boxplot(treatments_60_f$`change in number of species`~treatments_60_f$treatment, xlab="treatment", ylab="change in number of species", main="Change in number of species from Y0 to Y6 (fenced)")
```

Compare fenced vs unfenced quadrats in plot 7,9,10,12
```{r}
f_uf_30=as.data.frame(c(f_30,uf_30[c(81:96,113:128,129:144,161:176)])) #quadrats in plot 7, 9, 10, 12 (fenced and unfenced)
colnames(f_uf_30)=c("change in number of species")
f_uf_30$fencedVSunfenced=rep(c("Fenced","Unfenced"),c(64,64))
f_uf_30
boxplot(f_uf_30$`change in number of species`~f_uf_30$fencedVSunfenced, xlab="method", ylab="change in number of species", main="Change in number of species from Y0 to Y3")

f_uf_63=as.data.frame(c(f_63,uf_63[c(81:96,113:128,129:144,161:176)]))
colnames(f_uf_63)=c("change in number of species")
f_uf_63$fencedVSunfenced=rep(c("Fenced","Unfenced"),c(64,64))
f_uf_63
boxplot(f_uf_63$`change in number of species`~f_uf_63$fencedVSunfenced, xlab="method", ylab="change in number of species", main="Change in number of species from Y3 to Y6")

f_uf_60=as.data.frame(c(f_60,uf_60[c(97:112,129:144,145:160,177:192)]))
colnames(f_uf_60)=c("change in number of species")
f_uf_60$fencedVSunfenced=rep(c("Fenced","Unfenced"),c(64,64))
f_uf_60
boxplot(f_uf_60$`change in number of species`~f_uf_60$fencedVSunfenced, xlab="method", ylab="change in number of species", main="Change in number of species from Y0 to Y6")
```

## Calculate the sum of scores in each quadrat 
Unfenced
```{r}
Year0uf_scores=c()
for (i in 1:12)
{if (i %in% c(1,2,3,5,6,8))
  {for (j in 1:16)
    {a=sum(USdata[USdata$Year_monitoring==0 & USdata$Quadrat_fenced=='False' & USdata$Plot_number==i & USdata$Quadrat_number==j,]$Score)
    Year0uf_scores=append(Year0uf_scores,a)}
}
else 
  {for (j in seq(1,32,2))
    {a=sum(USdata[USdata$Year_monitoring==0 & USdata$Quadrat_fenced=='False' & USdata$Plot_number==i & USdata$Quadrat_number==j,]$Score)
    Year0uf_scores=append(Year0uf_scores,a)}
  }
    }
Year0uf_scores

Year3uf_scores=c()
for (i in 1:12)
{if (i %in% c(1,2,3,5,6,8))
  {for (j in 1:16)
    {a=sum(USdata[USdata$Year_monitoring==3 & USdata$Quadrat_fenced=='False' & USdata$Plot_number==i & USdata$Quadrat_number==j,]$Score)
    Year3uf_scores=append(Year3uf_scores,a)}
}
else 
  {for (j in seq(1,32,2))
  {a=sum(USdata[USdata$Year_monitoring==3 & USdata$Quadrat_fenced=='False' & USdata$Plot_number==i & USdata$Quadrat_number==j,]$Score)
    Year3uf_scores=append(Year3uf_scores,a)}
  }
    }
Year3uf_scores
#no data of plot 3 in year 3

Year6uf_scores=c()
for (i in 1:12)
{if (i %in% c(1,2,3,5,6,8))
  {for (j in 1:16)
  {a=sum(USdata[USdata$Year_monitoring==6 & USdata$Quadrat_fenced=='False' & USdata$Plot_number==i & USdata$Quadrat_number==j,]$Score)
    Year6uf_scores=append(Year6uf_scores,a)}
}
else 
  {for (j in seq(1,32,2))
  {a=sum(USdata[USdata$Year_monitoring==6 & USdata$Quadrat_fenced=='False' & USdata$Plot_number==i & USdata$Quadrat_number==j,]$Score)
    Year6uf_scores=append(Year6uf_scores,a)}
  }
    }
Year6uf_scores
```

Fenced
```{r}
Year0f_scores=c()
for (i in c(7,9,10,12))
  {for (j in seq(2,32,2))
    {a=sum(USdata[USdata$Year_monitoring==0 & USdata$Quadrat_fenced=='True' & USdata$Plot_number==i & USdata$Quadrat_number==j,]$Score)
    Year0f_scores=append(Year0f_scores,a)}
  }
Year0f_scores

Year3f_scores=c()
for (i in c(7,9,10,12))
  {for (j in seq(2,32,2))
    {a=sum(USdata[USdata$Year_monitoring==3 & USdata$Quadrat_fenced=='True' & USdata$Plot_number==i & USdata$Quadrat_number==j,]$Score)
    Year3f_scores=append(Year3f_scores,a)
    }
  }
Year3f_scores

Year6f_scores=c()
for (i in c(7,9,10,12))
  {for (j in seq(2,32,2))
    {a=sum(USdata[USdata$Year_monitoring==6 & USdata$Quadrat_fenced=='True' & USdata$Plot_number==i & USdata$Quadrat_number==j,]$Score)
    Year6f_scores=append(Year6f_scores,a)
    }
  }
Year6f_scores
```

## Calculate the score changes in each quadrat
```{r}
uf_score_30=Year3uf_scores[-(33:48)]-Year0uf_scores[-(33:48)] #exclude plot 3 cause missing data in Y3
uf_score_30

uf_score_63=Year6uf_scores[-(33:48)]-Year3uf_scores[-(33:48)] #exclude plot 3 cause missing data in Y3 
uf_score_63

uf_score_60=Year6uf_scores-Year0uf_scores
uf_score_60
```

Fenced
```{r}
f_score_30=Year3f_scores-Year0f_scores
f_score_30

f_score_63=Year6f_scores-Year3f_scores
f_score_63

f_score_60=Year6f_scores-Year0f_scores
f_score_60
```

#Plots to compare the change in scores
Unfenced quadrats with different treatments
```{r}
plot(uf_score_30, xlab="16 unfenced quadrats for each plot, 1 to 12, excluding plot 3", ylab="change in sum of scores", main="Change in sum of scores from Y0 to Y3 (unfenced)", col=c(rep(4,16),rep(6,16),rep(6,16),rep(2,16),rep(6,16),rep(2,16),rep(4,16),rep(2,16),rep(4,16),rep(6,16),rep(4,16)), pch=18)
abline(h=0, lty=2)
legend("top", legend = paste(c("T1","T2","Control")), col = c(4,2,6), pch = 18)

treatment_score_30=as.data.frame(c(uf_score_30))
colnames(treatment_score_30)=c("change in score")
treatment_score_30$treatment=c(rep("T1",16),rep("Control",16),rep("Control",16),rep("T2",16),rep("Control",16),rep("T2",16),rep("T1",16),rep("T2",16),rep("T1",16),rep("Control",16),rep("T1",16))
treatment_score_30
boxplot(treatment_score_30$`change in score`~treatment_score_30$treatment, xlab="treatment", ylab="change in sum of score", main="Change in sum of scores from Y0 to Y3 (unfenced)")

plot(uf_score_63, xlab="16 unfenced quadrats for each plot, 1 to 12, excluding plot 3", ylab="change in sum of scores", main="Change in sum of scores from Y3 to Y6 (unfenced)", col=c(rep(4,16),rep(6,16),rep(6,16),rep(2,16),rep(6,16),rep(2,16),rep(4,16),rep(2,16),rep(4,16),rep(6,16),rep(4,16)), pch=18)
abline(h=0, lty=2)
legend("bottom", legend = paste(c("T1","T2","Control")), col = c(4,2,6), pch = 18)

treatment_score_63=as.data.frame(c(uf_score_63))
colnames(treatment_score_63)=c("change in score")
treatment_score_63$treatment=c(rep("T1",16),rep("Control",16),rep("Control",16),rep("T2",16),rep("Control",16),rep("T2",16),rep("T1",16),rep("T2",16),rep("T1",16),rep("Control",16),rep("T1",16))
treatment_score_63
boxplot(treatment_score_63$`change in score`~treatment_score_63$treatment, xlab="treatment", ylab="change in sum of score", main="Change in sum of scores from Y3 to Y6 (unfenced)")

plot(uf_score_60, xlab="16 unfenced quadrats for each plot, 1 to 12", ylab="change in sum of scores", main="Change in sum of scores from Y0 to Y6 (unfenced)", col=c(rep(4,16),rep(6,16),rep(2,16),rep(6,16),rep(2,16),rep(6,16),rep(2,16),rep(4,16),rep(2,16),rep(4,16),rep(6,16),rep(4,16)), pch=18)
abline(h=0, lty=2)
legend("bottomleft", legend = paste(c("T1","T2","Control")), col = c(4,2,6), pch = 18)

treatment_score_60=as.data.frame(c(uf_score_60))
colnames(treatment_score_60)=c("change in score")
treatment_score_60$treatment=c(rep("T1",16),rep("Control",16),rep("T2",16),rep("Control",16),rep("T2",16),rep("Control",16),rep("T2",16),rep("T1",16),rep("T2",16),rep("T1",16),rep("Control",16),rep("T1",16))
treatment_score_60
boxplot(treatment_score_60$`change in score`~treatment_score_60$treatment, xlab="treatment", ylab="change in sum of score", main="Change in sum of scores from Y0 to Y6 (unfenced)")
```

Fenced quadrats with different treatments
```{r}
plot(f_score_30, xlab="16 fenced quadrats for plot 7,9,10,12 respectively", ylab="change in sum of scores", main="Change in sum of scores from Y0 to Y3 (fenced)", col=c(rep(2,16),rep(2,16),rep(4,16),rep(4,16)), pch=18)
abline(h=0, lty=2)
legend("bottomright", legend = paste(c("T1","T2","Control")), col = c(4,2,6), pch = 18)

treatment_score_30_f=as.data.frame(c(f_score_30))
colnames(treatment_score_30_f)=c("change in score")
treatment_score_30_f$treatment=c(rep("T2",32),rep("T1",32))
treatment_score_30_f
boxplot(treatment_score_30_f$`change in score`~treatment_score_30_f$treatment, xlab="treatment",ylab="change in sum of score", main="Change in sum of scores from Y0 to Y3 (fenced)")

plot(f_score_63, xlab="16 fenced quadrats for plot 7,9,10,12 respectively", ylab="change in sum of scores", main="Change in sum of scores from Y3 to Y6 (fenced)", col=c(rep(2,16),rep(2,16),rep(4,16),rep(4,16)), pch=18)
abline(h=0, lty=2)
legend("bottomright", legend = paste(c("T1","T2","Control")), col = c(4,2,6), pch = 18)

treatment_score_63_f=as.data.frame(c(f_score_63))
colnames(treatment_score_63_f)=c("change in score")
treatment_score_63_f$treatment=c(rep("T2",32),rep("T1",32))
treatment_score_63_f
boxplot(treatment_score_63_f$`change in score`~treatment_score_63_f$treatment, xlab="treatment",ylab="change in sum of score", main="Change in sum of scores from Y3 to Y6 (fenced)")

plot(f_score_60, xlab="16 fenced quadrats for plot 7,9,10,12 respectively", ylab="change in sum of scores", main="Change in sum of scores from Y0 to Y6 (fenced)", col=c(rep(2,16),rep(2,16),rep(4,16),rep(4,16)), pch=18)
abline(h=0, lty=2)
legend("bottomleft", legend = paste(c("T1","T2","Control")), col = c(4,2,6), pch = 18)

treatment_score_60_f=as.data.frame(c(f_score_60))
colnames(treatment_score_60_f)=c("change in score")
treatment_score_60_f$treatment=c(rep("T2",32),rep("T1",32))
treatment_score_60_f
boxplot(treatment_score_60_f$`change in score`~treatment_score_60_f$treatment, xlab="treatment",ylab="change in sum of score", main="Change in sum of scores from Y0 to Y6 (fenced)")
```

Compare fenced vs unfenced quadrats in plot 7,9,10,12
```{r}
f_uf_score_30=as.data.frame(c(f_score_30,uf_score_30[c(81:96,113:128,129:144,161:176)]))
colnames(f_uf_score_30)=c("change in score")
f_uf_score_30$fencedVSunfenced=rep(c("Fenced","Unfenced"),c(64,64))
f_uf_score_30
boxplot(f_uf_score_30$`change in score`~f_uf_score_30$fencedVSunfenced, xlab="method", ylab="change in sum of scores", main="Change in sum of scores from Y0 to Y3")
 
f_uf_score_63=as.data.frame(c(f_score_63,uf_score_63[c(81:96,113:128,129:144,161:176)]))
colnames(f_uf_score_63)=c("change in score")
f_uf_score_63$fencedVSunfenced=rep(c("Fenced","Unfenced"),c(64,64))
f_uf_score_63
boxplot(f_uf_score_63$`change in score`~f_uf_score_63$fencedVSunfenced, xlab="method", ylab="change in sum of scores", main="Change in sum of scores from Y3 to Y6")

f_uf_score_60=as.data.frame(c(f_score_60,uf_score_60[c(97:112,129:144,145:160,177:192)]))
colnames(f_uf_score_60)=c("change in score")
f_uf_score_60$fencedVSunfenced=rep(c("Fenced","Unfenced"),c(64,64))
f_uf_score_60
boxplot(f_uf_score_60$`change in score`~f_uf_score_60$fencedVSunfenced, xlab="method", ylab="change in sum of scores", main="Change in sum of scores from Y0 to Y6")
```

