---
title: "model_analysis"
output: word_document
date: "2022-08-18"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/Desktop/capstone/analysis/Juan/CWD")
library(MASS)
library(survival)
library(fitdistrplus)

CWDdata = read.csv("output.csv")
CWDdata[is.na(CWDdata)] <- 0
CWDdata$difference = CWDdata$year.6 - CWDdata$year.0

```

problems:
currently, since the CWD is a discrete variable, so we first fit the data via some discrete model families also tried some continuous models as well in order to find better models. We build models with guassian, poissson, quasipoisson, negative binomial and gamma, but all the models display the same problems, they model assumptions are violated and no obvious better models from the diagnosis plots, had no idea on choosing the best models, might need some suggestions.

## visualisation 
```{r visualisation, echo = TRUE}
hist(CWDdata$difference)
plot(CWDdata$difference, col = factor(CWDdata$treatment))
legend("topleft", legend = c("Gap", "Control", "Radial"), col=factor(CWDdata$treatment), pch=1)
boxplot(CWDdata$difference~CWDdata$treatment, xlab="Treatment", ylab="CWD difference", main="CWD difference against different treatments")

```

## Gussian model fitting
reform data with following two precedures:
(1)adding a constant.
(2)removal of plot 3; 
```{r data_redefinition, echo=FALSE}
CWDdata["abs_diff"] = CWDdata["difference"] + abs(min(CWDdata["difference"])) + 1
CWDdatarm3 = CWDdata[-c(3),]
```

The scale-location plot shows an increasing trend, it states the heteroskedasticity, that is the spread of the residuals is not equal, and the cook’s distance shows that the point 3 are a significant influential point. Treatment radial is statistically significant in this case, and treatment Gap is not. the outstanding outliers become 8,9,12, besides, QQplot and scale location plots state that violation of the model assumption, even though the Gap treatment become significant now, and AIC is smaller than before.
```{r guassian_remove_plot3, echo=FALSE}
#guassian model + plot3 removed
guass.model = glm(difference ~ treatment , family = gaussian, data = CWDdatarm3)
summary(guass.model)


par(mfrow = c(2,2))
plot(guass.model)

pure.guass.model <- glm(difference ~ 1 , family = gaussian, data = CWDdatarm3)
anova( pure.guass.model, guass.model, test = "F")
# the plot here shows that the plot 8,9,12 are always outliers, since the F-stat is <0.05, we will reject the null at 5% significance level, so the treatment is useful in predicting the difference, both of them are positive, so an increasing in the difference is indicated
```



## Poisson model fitting

The poisson model cannot be used directly, since the one of the difference in CWD is -6.5, so if we want to fit the poisson model, we have to convert all different to positive by adding a constant in r.(pros: no information is lost + we only care the direction rather than the magnitude) this time the scale-location model still shows an increasing trend. All the coefficient is statistically significant and positive. 
```{r poisson_abs_diff, echo=FALSE}
# model using poisson + abs_diff
#Poisson GLM

#visualisation and test if the data is poisson
hist(CWDdatarm3$abs_diff)

new_abs_diff <- ceiling(CWDdatarm3$abs_diff) # scale it to fit the poisson model
poisson.model1=fitdist(new_abs_diff , "pois")
poisson.model1
poisson.lambda=poisson.model1$estimate
plot(poisson.model1)

n=length(new_abs_diff)
samp.pct <- (1:n-0.5)/n
qqplot(x=qpois(samp.pct,poisson.lambda), y=sort(new_abs_diff), xlab = 'Theoretical quantiles', ylab = 'Empirical quantiles', main='Q-Q plot Poisson')
abline(0,1,col="red")
#fit the model
model0 = glm(new_abs_diff ~ treatment, family = poisson, data = CWDdatarm3)
summary(model0)
par(mfrow = c(2,2))
plot(model0)
#goodness of fit
qchisq(0.95,8)
library(msme)
P__disp(model0)
#adequate, but the dispersion is far greater than 1, so we need to fit the quasipoisson model


model1 <- glm(new_abs_diff ~ treatment, family = quasipoisson, data = CWDdatarm3)
summary(model1)
par(mfrow = c(2,2))
plot(model1)
#goodness of fit
qchisq(0.95,model1$df.residual)
library(msme)
P__disp(model1)


```

## Negative binomial fitting
```{r negative_binomial, echo = TRUE}
# model using negative binomial + abs_diff
#Negative Binomial GLM

#visualisation and test if the data is poisson
nb.model1=fitdist(new_abs_diff, "nbinom")
nb.model1
nb.size=nb.model1$estimate[1]
nb.mu = nb.model1$estimate[2]
nb.prob=nb.size/(nb.size+nb.mu) 
plot(nb.model1)

n=length(new_abs_diff)
samp.pct <- (1:n-0.5)/n
qqplot(x=qnbinom(samp.pct,nb.size, nb.prob), y=sort(new_abs_diff), xlab = 'Theoretical quantiles', ylab = 'Empirical quantiles', main='Q-Q plot Negative Binomial')
abline(0,1,col="red")
#fit the model
model2 = glm.nb(new_abs_diff ~ treatment, data = CWDdatarm3)
summary(model2)
par(mfrow = c(2,2))
plot(model2)
#goodness of fit
qchisq(0.95,8)
library(msme)
P__disp(model2)
# both coefficient is positive and significant, but the pearson coefficient is less than 15.51, so we fail to reject them from this perpective
```

## Gamma model fitting

The Gamma also takes no negative values, we first fit the inverse link function, the outliers are 9(Radial), 6(C),11(C), and no significant points are presented, but the normal QQ plot is left-skewed, and the scale-location is showed an increasing trend . All the coefficients from the model are negative but not statistically significant. But the AIC is the lowest among all the models.
```{r gamma_diff_inverse, echo=FALSE}
## Gamma + abs_diff + inverse
new_abs_diff <- ceiling(CWDdatarm3$abs_diff) # scale it to fit the gamma model
gamma.model1=fitdist(new_abs_diff , "gamma")
gamma.model1
gamma.alpha=gamma.model1$estimate[1]
gamma.beta =gamma.model1$estimate[2]
plot(gamma.model1)

n=length(new_abs_diff)
samp.pct <- (1:n-0.5)/n
qqplot(x=qgamma(samp.pct,gamma.alpha, gamma.beta), y=sort(new_abs_diff), xlab = 'Theoretical quantiles', ylab = 'Empirical quantiles', main='Q-Q plot gamma')
abline(0,1,col="red")

model3 = glm(new_abs_diff ~ treatment, family = Gamma(link = "inverse"), data = CWDdatarm3)
summary(model3)

par(mfrow = c(2,2))
plot(model3)
#goodness of fit
qchisq(0.95,10)
library(msme)
P__disp(model3)
# model fit is ok, but worse than the poisson, and fail to reject the null

```