---
title: "Understorey_richness"
author: "Juan Dai"
date: "2022-09-18"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/Desktop/capstone/analysis/Juan/Understorey_priority2&low")
library(MASS)
library(survival)
library(fitdistrplus)
library(lme4)
```

poisson, negative binomial

```{r data_loading, echo=TRUE}
richness_df <- read.csv('./dataset/tables/species_richness_quadrats.csv')
richness_df_plot <- read.csv('./contingency tables/output/richness/species_richness_plots.csv')

# Treatments, Plot numbers, Quadrat numbers, Life forms as factors
richness_df$Treatment <- factor(richness_df$Treatment)
richness_df$Plot.Number <- factor(richness_df$Plot.Number)
richness_df$Quadrat.Number <- factor(richness_df$Quadrat.Number)

# Convert Fenced, Gap columns to boolean
richness_df$Gap <- richness_df$Gap=='True'
richness_df$Fenced <- richness_df$Fenced=='True'

richness_df_plot$Gap <- richness_df_plot$Gap=='True'
richness_df_plot$Fenced <- richness_df_plot$Fenced=='True'

# Remove null rows
richness_df <- na.omit(richness_df)
#richness_df_plot <- na.omit(richness_df_plot) 
# what is the purpose of the richness_df_plot
```

## Data preparation
```{r data, echo=TRUE}
richness_df$Y0.Y3.Difference <- richness_df$X3 - richness_df$X0
richness_df$Y3.Y6.Difference <- richness_df$X6 - richness_df$X3
richness_df$Y0.Y6.Difference <- richness_df$X6 - richness_df$X0
richness_df$Y0.Y3.Abs_Difference <- richness_df$Y0.Y3.Difference + abs(min(richness_df$Y0.Y3.Difference)) + 1
richness_df$Y3.Y6.Abs_Difference <- richness_df$Y3.Y6.Difference + abs(min(richness_df$Y3.Y6.Difference)) + 1
richness_df$Y0.Y6.Abs_Difference <- richness_df$Y0.Y6.Difference + abs(min(richness_df$Y0.Y6.Difference)) + 1
```

# modelling
```{r 0to3_gaussian, echo=True}
# guassian modeling
# model using gaussian
Y0.Y3.guas.model=fitdist(richness_df$Y0.Y3.Difference, "norm")
Y0.Y3.guas.model
guas.mean=Y0.Y3.guas.model$estimate[1]
guas.sd =Y0.Y3.guas.model$estimate[2]
plot(Y0.Y3.guas.model)

n=length(richness_df$Y0.Y3.Difference)
samp.pct <- (1:n-0.5)/n
qqplot(x=qnorm(samp.pct,guas.mean, guas.sd), y=sort(richness_df$Y0.Y3.Difference), xlab = 'Theoretical quantiles', ylab = 'Empirical quantiles', main='Q-Q plot Guassian')
abline(0,1,col="red")

Y0.Y3.guass.model = glm(Y0.Y3.Difference ~ Treatment  + Gap + Quadrat.Number + Fenced, family = gaussian, data = richness_df)
summary(Y0.Y3.guass.model)
par(mfrow = c(2,2))

plot(Y0.Y3.guass.model)
Y0.Y3.pure.guass.model <- glm(Y0.Y3.Difference ~ 1, family = gaussian, data = richness_df)
anova( Y0.Y3.pure.guass.model, Y0.Y3.guass.model, test = "F")
# In model fitting, the treatment radial, gap are statistically significant, but when we use F-stats to test the H0: all the coefficient is zero, it states that we fail to reject null at 5% significance level
# Pr(>F) = 0.5084
```
```{r model_selection_guassian, echo=TRUE}
Y0.Y3.guass.model_selected <-step(Y0.Y3.guass.model, scope=~.)
summary(Y0.Y3.guass.model_selected)
# only treatment and gap is significant, all other factor is not significant in predicting the yr_richness_difference, AIC decreases from 1670 to 1625
anova(Y0.Y3.guass.model_selected, Y0.Y3.pure.guass.model,test = "F")
# the selected model states that the gap and treatment is related to the richness
#Pr(>F) = 0.0006474
```

```{r glm_fit_gamma, echo=TRUE}
Y0.Y3.gamma.model1=fitdist(richness_df$Y0.Y3.Abs_Difference, "gamma")
Y0.Y3.gamma.model1
gamma.alpha=Y0.Y3.gamma.model1$estimate[1]
gamma.beta =Y0.Y3.gamma.model1$estimate[2]
plot(Y0.Y3.gamma.model1)

n=length(richness_df$Y0.Y3.Abs_Difference)
samp.pct <- (1:n-0.5)/n
qqplot(x=qgamma(samp.pct,gamma.alpha, gamma.beta), y=sort(richness_df$Y0.Y3.Abs_Difference), xlab = 'Theoretical quantiles', ylab = 'Empirical quantiles', main='Q-Q plot gamma')
abline(0,1,col="red")


Y0.Y3.gamma.model2 = glm(Y0.Y3.Abs_Difference ~ Treatment  + Gap + Quadrat.Number + Fenced, family = Gamma, data = richness_df)
summary(Y0.Y3.gamma.model2)
par(mfrow = c(2,2))
plot(Y0.Y3.gamma.model2)
```

```{r model_selection_gamma, echo=TRUE}
Y0.Y3.gamma.model_selected <-step(Y0.Y3.gamma.model2, scope=~.)
summary(Y0.Y3.gamma.model_selected)
# AIC decrease from 1702 to 1653
Y0.Y3.empty_gamma_model <- glm(Y0.Y3.Abs_Difference ~ 1, family = Gamma, data = richness_df)
anova(Y0.Y3.gamma.model_selected, Y0.Y3.empty_gamma_model, test = "F")
# the model is significant at 5%, suggest that the coefficient for treatment and gap is not zero, Pr(>F) = 0.0006044
```

```{r fit_poisson_model, echo=TRUE}
Y0.Y3.poisson.model1=fitdist(richness_df$Y0.Y3.Abs_Difference, "pois")
Y0.Y3.poisson.model1
poisson.lambda=Y0.Y3.poisson.model1$estimate
plot(Y0.Y3.poisson.model1)

n=length(richness_df$Y0.Y3.Abs_Difference)
samp.pct <- (1:n-0.5)/n
qqplot(x=qpois(samp.pct,poisson.lambda), y=sort(richness_df$Y0.Y3.Abs_Difference), xlab = 'Theoretical quantiles', ylab = 'Empirical quantiles', main='Q-Q plot Poisson')
abline(0,1,col="red")



Y0.Y3.poisson.model2 = glm(Y0.Y3.Abs_Difference~ Treatment  + Gap + Quadrat.Number + Fenced, family = poisson, data = richness_df)
summary(Y0.Y3.poisson.model2)
par(mfrow = c(2,2))
plot(Y0.Y3.poisson.model2)
```

```{r model_selection_poisson, echo=TRUE}
Y0.Y3.poisson.model_selected <-step(Y0.Y3.poisson.model2, scope=~.)
summary(Y0.Y3.poisson.model_selected)
#AIC decrease from 1708 to 1674
Y0.Y3.empty_poisson_model <- glm(Y0.Y3.Abs_Difference ~ 1, family = poisson, data = richness_df)
anova(Y0.Y3.poisson.model_selected, Y0.Y3.empty_poisson_model, test = "Chi")
# the model is significant at 5%, suggest that the coefficient for treatment and gap is not zero, Pr(>F) = 8.643e-07
```

```{r fit_negative_binomial, echo=TRUE}
#Negative Binomial GLM

#visualisation and test if the data is negative binomial
Y0.Y3.nb.model1=fitdist(richness_df$Y0.Y3.Abs_Difference, "nbinom")
Y0.Y3.nb.model1
nb.size=Y0.Y3.nb.model1$estimate[1]
nb.mu = Y0.Y3.nb.model1$estimate[2]
nb.prob=nb.size/(nb.size+nb.mu) 
plot(Y0.Y3.nb.model1)

n=length(richness_df$Y0.Y3.Abs_Difference)
samp.pct <- (1:n-0.5)/n
qqplot(x=qnbinom(samp.pct,nb.size, nb.prob), y=sort(richness_df$Y0.Y3.Abs_Difference), xlab = 'Theoretical quantiles', ylab = 'Empirical quantiles', main='Q-Q plot Negative Binomial')
abline(0,1,col="red")
#fit the model
Y0.Y3.nb.model2 = glm.nb(Y0.Y3.Abs_Difference~ Treatment  + Gap + Quadrat.Number + Fenced, data = richness_df)
summary(Y0.Y3.nb.model2 )
par(mfrow = c(2,2))
plot(Y0.Y3.nb.model2)
```

```{r model_selection_negative_binomial, echo=TRUE}
Y0.Y3.nb.model_selected <-step(Y0.Y3.nb.model2, scope=~.)
summary(Y0.Y3.nb.model_selected)
# AIC decrease from 1674 to 1629
Y0.Y3.empty_nb_model <- glm.nb(Y0.Y3.Abs_Difference ~ 1,  data = richness_df)
anova(Y0.Y3.nb.model_selected, Y0.Y3.empty_nb_model, test = "LR")
# the model is significant at 5%, suggest that the coefficient for treatment and gap is not zero, Pr(>z) = 0.0008629232
```

**selected based on AIC values, we will choose the gaussian model, because it gives lowest AIC value of 1625.**



## diffrent modeling via prediction
```{r model_prediction_gaussian, echo=TRUE}
Y0.Y3.guass.model2 = glm(X3 ~ X0 + Treatment  + Gap + Quadrat.Number + Fenced, family = gaussian, data = richness_df)
summary(Y0.Y3.guass.model2)
par(mfrow = c(2,2))
plot(Y0.Y3.guass.model2)

Y0.Y3.gamma.model3 = glm(X3 ~ X0 + Treatment  + Gap + Quadrat.Number + Fenced, family = Gamma, data = richness_df)
summary(Y0.Y3.gamma.model3)
par(mfrow = c(2,2))

plot(Y0.Y3.gamma.model3)

Y0.Y3.poisson.model3 = glm(X3 ~ X0 + Treatment  + Gap + Quadrat.Number + Fenced, family = poisson, data = richness_df)
summary(Y0.Y3.poisson.model3)
par(mfrow = c(2,2))

plot(Y0.Y3.poisson.model3)
```

```{r model_selection_prediction_gaussian, echo=TRUE}
Y0.Y3.select_guass.model2 <- step(Y0.Y3.guass.model2, scope=~.)
summary(Y0.Y3.select_guass.model2)

Y0.Y3.select_gamma.model3 <- step(Y0.Y3.gamma.model3, scope=~.)
summary(Y0.Y3.select_gamma.model3)

Y0.Y3.select_poisson.model3 <- step(Y0.Y3.poisson.model3, scope=~.)
summary(Y0.Y3.select_poisson.model3)
# according to the AIC, we will choose the poisson model, since it has the lowest AIC of 1493.1
```
## the result from gamma is contradicted with the result from gaussian and poisson.


**Therefore, using the prediction with poisson model, it has the lowest AIC of 1493.1**
```{r model_selection_prediction_gaussian_log_glmer, echo=TRUE}
Y0.Y3.glmer.poiss1= glmer(X3 ~ X0 + Treatment  + Gap +  Fenced+ (1|Plot.Number), family = poisson, data = richness_df)
summary(Y0.Y3.glmer.poiss1)
print(Y0.Y3.glmer.poiss1, correlation=TRUE)

#diagnosis plots
plot(Y0.Y3.glmer.poiss1)
plot(Y0.Y3.glmer.poiss1, resid(., scaled=TRUE) ~ fitted(.), abline = 0,pch=16,col=richness_df$Plot.Number,xlab="Fitted values",ylab="Standardised residuals")


qqnorm(resid(Y0.Y3.glmer.poiss1),pch=16,col=richness_df$Plot.Number)
qqline(resid(Y0.Y3.glmer.poiss1))


anova(Y0.Y3.glmer.poiss1, Y0.Y3.select_poisson.model3)


Y0.Y3.glmer.poiss2= glmer(X3 ~ X0 + Treatment  + Gap +  Fenced  +(X0|Plot.Number), family = poisson, data = richness_df)
summary(Y0.Y3.glmer.poiss2)
par(mfrow = c(2,2))
plot(Y0.Y3.glmer.poiss2)
anova(Y0.Y3.glmer.poiss2,Y0.Y3.select_poisson.model3,  test = "Chi")

Y0.Y3.glmer.poiss3= glmer(X3 ~ X0 + Treatment  + Gap +  Fenced + (1|Quadrat.Number), family = poisson, data = richness_df)
summary(Y0.Y3.glmer.poiss3)
print(Y0.Y3.glmer.poiss3, correlation=TRUE)
par(mfrow = c(2,2))
plot(Y0.Y3.glmer.poiss3)
anova(Y0.Y3.glmer.poiss3, Y0.Y3.select_poisson.model3)


# according to the Chisq, we will choose the poisson model without the random effect, since all the anova states no random effect,

```






##Yr3 to Yr6
# modelling
```{r gaussian_Y36, echo=True}
# guassian modeling
# model using gaussian
Y3.Y6.guas.model=fitdist(richness_df$Y3.Y6.Difference, "norm")
Y3.Y6.guas.model
guas.mean=Y3.Y6.guas.model$estimate[1]
guas.sd =Y3.Y6.guas.model$estimate[2]
plot(Y3.Y6.guas.model)

n=length(richness_df$Y3.Y6.Difference)
samp.pct <- (1:n-0.5)/n
qqplot(x=qnorm(samp.pct,guas.mean, guas.sd), y=sort(richness_df$Y3.Y6.Difference), xlab = 'Theoretical quantiles', ylab = 'Empirical quantiles', main='Q-Q plot Guassian')
abline(0,1,col="red")

Y3.Y6.guass.model = glm(Y3.Y6.Difference ~ Treatment  + Gap + Quadrat.Number + Fenced, family = gaussian, data = richness_df)
summary(Y3.Y6.guass.model)
par(mfrow = c(2,2))

plot(Y3.Y6.guass.model)
Y3.Y6.pure.guass.model <- glm(Y3.Y6.Difference ~ 1, family = gaussian, data = richness_df)
anova(Y3.Y6.pure.guass.model, Y3.Y6.guass.model, test = "F")
# In model fitting, the treatment radial, gap are statistically significant, also when we use F-stats to test the H0: all the coefficient is zero, it states that we reject null at 5% significance level
# Pr(>F) = 0.03462
```


```{r model_selection_guassian_Y36, echo=TRUE}
Y3.Y6.guass.model_selected <-step(Y3.Y6.guass.model, scope=~.)
summary(Y3.Y6.guass.model_selected)
# only treatment and gap is significant, all other factor is not significant in predicting the yr_richness_difference, AIC decreases from 1685 to 1650
anova(Y3.Y6.guass.model_selected, Y3.Y6.pure.guass.model,test = "F")
# the selected model states that the gap and treatment is related to the richness
#Pr(>F) = 3.422e-06
```

```{r glm_fit_gamma_Y36, echo=TRUE}
Y3.Y6.gamma.model1=fitdist(richness_df$Y3.Y6.Abs_Difference, "gamma")
Y3.Y6.gamma.model1
gamma.alpha=Y3.Y6.gamma.model1$estimate[1]
gamma.beta =Y3.Y6.gamma.model1$estimate[2]
plot(Y3.Y6.gamma.model1)

n=length(richness_df$Y3.Y6.Abs_Difference)
samp.pct <- (1:n-0.5)/n
qqplot(x=qgamma(samp.pct,gamma.alpha, gamma.beta), y=sort(richness_df$Y3.Y6.Abs_Difference), xlab = 'Theoretical quantiles', ylab = 'Empirical quantiles', main='Q-Q plot gamma')
abline(0,1,col="red")


Y3.Y6.gamma.model2 = glm(Y3.Y6.Abs_Difference ~ Treatment  + Gap + Quadrat.Number + Fenced, family = Gamma, data = richness_df)
summary(Y3.Y6.gamma.model2)
par(mfrow = c(2,2))
plot(Y3.Y6.gamma.model2)
# the model fit is worse than the gaussian fit, from the qq-plot
```

```{r model_selection_gamma_Y36, echo=TRUE}
Y3.Y6.gamma.model_selected <-step(Y3.Y6.gamma.model2, scope=~.)
summary(Y3.Y6.gamma.model_selected)
# AIC decrease from 1738 to 1695
Y3.Y6.empty_gamma_model <- glm(Y3.Y6.Abs_Difference ~ 1, family = Gamma, data = richness_df)
anova(Y3.Y6.gamma.model_selected, Y3.Y6.empty_gamma_model, test = "F")
# the model is significant at 5%, suggest that the coefficient for treatment and gap is not zero, Pr(>F) = 4.893e-06
```

```{r fit_poisson_model_Y36, echo=TRUE}
Y3.Y6.poisson.model1=fitdist(richness_df$Y3.Y6.Abs_Difference, "pois")
Y3.Y6.poisson.model1
poisson.lambda=Y3.Y6.poisson.model1$estimate
plot(Y3.Y6.poisson.model1)

n=length(richness_df$Y3.Y6.Abs_Difference)
samp.pct <- (1:n-0.5)/n
qqplot(x=qpois(samp.pct,poisson.lambda), y=sort(richness_df$Y3.Y6.Abs_Difference), xlab = 'Theoretical quantiles', ylab = 'Empirical quantiles', main='Q-Q plot Poisson')
abline(0,1,col="red")



Y3.Y6.poisson.model2 = glm(Y3.Y6.Abs_Difference~ Treatment  + Gap + Quadrat.Number + Fenced, family = poisson, data = richness_df)
summary(Y3.Y6.poisson.model2)
par(mfrow = c(2,2))
plot(Y3.Y6.poisson.model2)
```

```{r model_selection_poisson_Y36, echo=TRUE}
Y3.Y6.poisson.model_selected <-step(Y3.Y6.poisson.model2, scope=~.)
summary(Y3.Y6.poisson.model_selected)
#AIC decrease from 1730 to 1713
Y3.Y6.empty_poisson_model <- glm(Y3.Y6.Abs_Difference ~ 1, family = poisson, data = richness_df)
anova(Y3.Y6.poisson.model_selected, Y3.Y6.empty_poisson_model, test = "Chi")
# the model is significant at 5%, suggest that the coefficient for treatment and gap is not zero, Pr(>F) = 1.121e-10
```

**we still choose gaussian model, since it has lowest AIC of 1650**

## diffrent modeling via prediction for yr3 to yr6
```{r model_prediction_gaussian_Y36, echo=TRUE}
Y3.Y6.guass.model2 = glm(X6 ~ X3 + Treatment  + Gap + Quadrat.Number + Fenced, family = gaussian, data = richness_df)
summary(Y3.Y6.guass.model2)
par(mfrow = c(2,2))
plot(Y3.Y6.guass.model2)

Y3.Y6.gamma.model3 = glm(X6 ~ X3 + Treatment  + Gap + Quadrat.Number + Fenced, family = Gamma, data = richness_df)
summary(Y3.Y6.gamma.model3)
par(mfrow = c(2,2))

plot(Y3.Y6.gamma.model3)

Y3.Y6.poisson.model3 = glm(X6 ~ X3 + Treatment  + Gap + Quadrat.Number + Fenced, family = poisson, data = richness_df)
summary(Y3.Y6.poisson.model3)
par(mfrow = c(2,2))

plot(Y3.Y6.poisson.model3)
```

```{r model_selection_prediction_gaussian_Y36, echo=TRUE}
Y3.Y6.select_guass.model2 <- step(Y3.Y6.guass.model2, scope=~.)
summary(Y3.Y6.select_guass.model2)

Y3.Y6.select_gamma.model3 <- step(Y3.Y6.gamma.model3, scope=~.)
summary(Y3.Y6.select_gamma.model3)

Y3.Y6.select_poisson.model3 <- step(Y3.Y6.poisson.model3, scope=~.)
summary(Y3.Y6.select_poisson.model3)

# if we use the AIC to choose the model, we will choose the gamma model, it yields the lowest AIC of 1538
```


##Yr0 to Yr6
# modelling
```{r gaussian_Y06, echo=True}
# guassian modeling
# model using gaussian
Y0.Y6.guas.model=fitdist(richness_df$Y0.Y6.Difference, "norm")
Y0.Y6.guas.model
guas.mean=Y0.Y6.guas.model$estimate[1]
guas.sd =Y0.Y6.guas.model$estimate[2]
plot(Y0.Y6.guas.model)

n=length(richness_df$Y0.Y6.Difference)
samp.pct <- (1:n-0.5)/n
qqplot(x=qnorm(samp.pct,guas.mean, guas.sd), y=sort(richness_df$Y0.Y6.Difference), xlab = 'Theoretical quantiles', ylab = 'Empirical quantiles', main = 'Q-Q plot Guassian')
abline(0,1,col="red")

Y0.Y6.guass.model = glm(Y0.Y6.Difference ~ Treatment  + Gap + Quadrat.Number + Fenced, family = gaussian, data = richness_df)
summary(Y0.Y6.guass.model)
par(mfrow = c(2,2))

plot(Y0.Y6.guass.model)
Y0.Y6.pure.guass.model <- glm(Y0.Y6.Difference ~ 1, family = gaussian, data = richness_df)
anova(Y0.Y6.pure.guass.model, Y0.Y6.guass.model, test = "F")
# In model fitting, the intercept, gap and some quadrates number are statistically significant, but when we use F-stats to test the H0: all the coefficient is zero, it states that we fail to reject null at 5% significance level
# Pr(>F) = 0.2032
```


```{r model_selection_guassian_Y06, echo=TRUE}
Y0.Y6.guass.model_selected <-step(Y0.Y6.guass.model, scope=~.)
summary(Y0.Y6.guass.model_selected)
# only treatment radial is significant, all other factor is not significant in predicting the yr_richness_difference, AIC decreases from 1642 to 1610
anova(Y0.Y6.guass.model_selected, Y0.Y6.pure.guass.model,test = "F")
# the selected model states that the treatment and gap is related to the richness, although only treatment radial is significant
#Pr(>F) = 0.004403
```

```{r glm_fit_gamma_Y06, echo=TRUE}
Y0.Y6.gamma.model1=fitdist(richness_df$Y0.Y6.Abs_Difference, "gamma")
Y0.Y6.gamma.model1
gamma.alpha=Y0.Y6.gamma.model1$estimate[1]
gamma.beta =Y0.Y6.gamma.model1$estimate[2]
plot(Y0.Y6.gamma.model1)

n=length(richness_df$Y0.Y6.Abs_Difference)
samp.pct <- (1:n-0.5)/n
qqplot(x=qgamma(samp.pct,gamma.alpha, gamma.beta), y=sort(richness_df$Y0.Y6.Abs_Difference), xlab = 'Theoretical quantiles', ylab = 'Empirical quantiles', main='Q-Q plot gamma')
abline(0,1,col="red")


Y0.Y6.gamma.model2 = glm(Y0.Y6.Abs_Difference ~ Treatment  + Gap + Quadrat.Number + Fenced, family = Gamma, data = richness_df)
summary(Y0.Y6.gamma.model2)
par(mfrow = c(2,2))
plot(Y0.Y6.gamma.model2)
# the model fit is worse than the gaussian fit, from the qq-plot, it dispalys a curvature
```

```{r model_selection_gamma_Y06, echo=TRUE}
Y0.Y6.gamma.model_selected <-step(Y0.Y6.gamma.model2, scope=~.)
summary(Y0.Y6.gamma.model_selected)
# AIC decrease from 1701 to 1661
Y0.Y6.empty_gamma_model <- glm(Y0.Y6.Abs_Difference ~ 1, family = Gamma, data = richness_df)
anova(Y0.Y6.gamma.model_selected, Y0.Y6.empty_gamma_model, test = "F")
# the model is significant at 5%, suggest that the coefficient for treatment is not zero, Pr(>F) = 0.004019
```

```{r fit_poisson_model_Y06, echo=TRUE}
Y0.Y6.poisson.model1=fitdist(richness_df$Y0.Y6.Abs_Difference, "pois")
Y0.Y6.poisson.model1
poisson.lambda=Y0.Y6.poisson.model1$estimate
plot(Y0.Y6.poisson.model1)

n=length(richness_df$Y0.Y6.Abs_Difference)
samp.pct <- (1:n-0.5)/n
qqplot(x=qpois(samp.pct,poisson.lambda), y=sort(richness_df$Y0.Y6.Abs_Difference), xlab = 'Theoretical quantiles', ylab = 'Empirical quantiles', main='Q-Q plot Poisson')
abline(0,1,col="red")



Y0.Y6.poisson.model2 = glm(Y0.Y6.Abs_Difference~ Treatment  + Gap + Quadrat.Number + Fenced, family = poisson, data = richness_df)
summary(Y0.Y6.poisson.model2)
par(mfrow = c(2,2))
plot(Y0.Y6.poisson.model2)
```

```{r model_selection_poisson_Y06, echo=TRUE}
Y0.Y6.poisson.model_selected <-step(Y0.Y6.poisson.model2, scope=~.)
summary(Y0.Y6.poisson.model_selected)
#AIC decrease from 1663 to 1640
Y0.Y6.empty_poisson_model <- glm(Y0.Y6.Abs_Difference ~ 1, family = poisson, data = richness_df)
anova(Y0.Y6.poisson.model_selected, Y0.Y6.empty_poisson_model, test = "Chi")
# the model is significant at 5%, suggest that the coefficient for treatment is not zero at 5% significance level, Pr(>F) = 0.0003047
```

**Based on the AIC, we will choose guassian model, since it give the lowest AIC of 1610 compared to gamma and poisson**


## diffrent modeling via prediction for yr0 to yr6 (not include X3)
```{r model_prediction_gaussian_Y06, echo=TRUE}
Y0.Y6.guass.model2 = glm(X6 ~ X0 + Treatment  + Gap + Quadrat.Number + Fenced, family = gaussian, data = richness_df)
summary(Y0.Y6.guass.model2)
par(mfrow = c(2,2))
plot(Y0.Y6.guass.model2)

Y0.Y6.gamma.model3 = glm(X6 ~ X0 + Treatment  + Gap + Quadrat.Number + Fenced, family = Gamma, data = richness_df)
summary(Y0.Y6.gamma.model3)
par(mfrow = c(2,2))

plot(Y0.Y6.gamma.model3)

Y0.Y6.poisson.model3 = glm(X6 ~ X0 + Treatment  + Gap + Quadrat.Number + Fenced, family = poisson, data = richness_df)
summary(Y0.Y6.poisson.model3)
par(mfrow = c(2,2))

plot(Y0.Y6.poisson.model3)
```

```{r model_selection_prediction_gaussian_Y06, echo=TRUE}
Y0.Y6.select_guass.model2 <- step(Y0.Y6.guass.model2, scope=~.)
summary(Y0.Y6.select_guass.model2)

Y0.Y6.select_gamma.model3 <- step(Y0.Y6.gamma.model3, scope=~.)
summary(Y0.Y6.select_gamma.model3)

Y0.Y6.select_poisson.model3 <- step(Y0.Y6.poisson.model3, scope=~.)
summary(Y0.Y6.select_poisson.model3)

# if we use the AIC to choose the model, we will choose the poisson model, it yields the lowest AIC of 1514, in the poisson model, all the treatment, X0, Gap and Fence are significant in predicting the X6 richness at 5% significance level.
#both treatment factor are negative.
```

## diffrent modeling via prediction for yr0 to yr6 (include X3)
```{r model_prediction_gaussian_Y036, echo=TRUE}
Y0.Y6.guass.model3 = glm(X6 ~ X0 + X3 + Treatment  + Gap + Quadrat.Number + Fenced, family = gaussian, data = richness_df)
summary(Y0.Y6.guass.model3)
par(mfrow = c(2,2))
plot(Y0.Y6.guass.model3)

Y0.Y6.gamma.model4 = glm(X6 ~ X0 + X3+ Treatment  + Gap + Quadrat.Number + Fenced, family = Gamma, data = richness_df)
summary(Y0.Y6.gamma.model4)
par(mfrow = c(2,2))

plot(Y0.Y6.gamma.model4)

Y0.Y6.poisson.model4 = glm(X6 ~ X0 + X3 + Treatment  + Gap + Quadrat.Number + Fenced, family = poisson, data = richness_df)
summary(Y0.Y6.poisson.model4)
par(mfrow = c(2,2))

plot(Y0.Y6.poisson.model4)
```

```{r model_selection_prediction_gaussian_Y036, echo=TRUE}
Y0.Y6.select_guass.model3 <- step(Y0.Y6.guass.model3, scope=~.)
summary(Y0.Y6.select_guass.model3)

Y0.Y6.select_gamma.model4 <- step(Y0.Y6.gamma.model4, scope=~.)
summary(Y0.Y6.select_gamma.model4)

Y0.Y6.select_poisson.model4 <- step(Y0.Y6.poisson.model4, scope=~.)
summary(Y0.Y6.select_poisson.model4)

# if we use the AIC to choose the model, we will choose the poisson model, it yields the lowest AIC of 1508, in the poisson model, both X0 and X3, treatment, Gap and Fence are significant in predicting the richness of X6 at 5% significance level.
#Both treatment state a negative effect on richness increase
```

```{r model_selection_prediction_gaussian_log_glmer, echo=TRUE}
Y0.Y6.glmer.poiss1= glmer(X6 ~ X0 + X3 + Treatment  + Gap + Quadrat.Number + Fenced+ (1|Plot.Number), family = poisson, data = richness_df)
summary(Y0.Y6.glmer.poiss1)
print(Y0.Y6.glmer.poiss1, correlation=TRUE)

#diagnosis plots
plot(Y0.Y6.glmer.poiss1)
plot(Y0.Y6.glmer.poiss1, resid(., scaled=TRUE) ~ fitted(.), abline = 0,pch=16,col=richness_df$Plot.Number,xlab="Fitted values",ylab="Standardised residuals")


qqnorm(resid(Y0.Y6.glmer.poiss1),pch=16,col=richness_df$Plot.Number)
qqline(resid(Y0.Y6.glmer.poiss1))


anova(Y0.Y6.glmer.poiss1, Y0.Y3.select_poisson.model3)


Y0.Y6.glmer.poiss2= glmer(X6 ~ X0 + X3 + Treatment  + Gap + Quadrat.Number + Fenced + (1|Plot.Number) + (X0|Plot.Number), family = poisson, data = richness_df)
summary(Y0.Y6.glmer.poiss2)
par(mfrow = c(2,2))
plot(Y0.Y6.glmer.poiss2)
anova(Y0.Y6.glmer.poiss2,Y0.Y6.glmer.poiss1,  test = "Chi")

Y0.Y6.glmer.poiss3= glmer(X6 ~ X0 + X3 + Treatment  + Gap + Quadrat.Number + Fenced + (1|Plot.Number)+ (1|Quadrat.Number), family = poisson, data = richness_df)
summary(Y0.Y6.glmer.poiss3)
print(Y0.Y6.glmer.poiss3, correlation=TRUE)
par(mfrow = c(2,2))
plot(Y0.Y6.glmer.poiss3)
anova(Y0.Y6.glmer.poiss3, Y0.Y6.glmer.poiss1)


# according to the Chisq, we will choose the poisson model with the random effect of (1|plot.number)

```