---
title: "DBH_visualization"
output: html_document
---

# data preprocessing
```{r}
# load original data
DBHdata = read.csv("~/Desktop/capstone/Datascience/DataScienceProject/Tree DBH/Final Models/dataset/all_plots.csv", stringsAsFactors = FALSE)
```

```{r}
# Skip this, it will output a csv!!!!!!!!!!!!
# output sub-csv from original data "all_plots.csv", which contains treegrowth = DBHyear6 - DBHyear0 <= 0
DBHdata$Tree.growth = DBHdata$DBH.year.6 - DBHdata$DBH.year.0
write.csv(DBHdata[DBHdata$Tree.growth <= 0 & !is.na(DBHdata$Tree.growth), ], "nonpositive treegrowth_original data.csv", row.names = FALSE)
# Skip this, it will output a csv!!!!!!!!!!!!
```

```{r}
summary(DBHdata)
sapply(DBHdata, typeof)
```

Currently, "Class.year.0" and "Class.year.6" are of type "logical", we don't modify them as they will not be used in the following preliminary analyses.
The six hollow columns need to be converted to integer type, but first, we need to check values for each column and deal with incorrect values, NAs and blank.

check incorrect values
```{r}
sapply(DBHdata[c(2,3,4,5,8,9,10,11,12,13,14,15)], unique)
```
```{r}
# assign correct values in the first 4 hollow columns: "0/0" should be "0", "yes" and "Yes " should both be "Yes"
DBHdata[DBHdata$Hollows.Class.1.year.0 == "0/0", 10:12] = "0"
DBHdata[DBHdata$Hollows.Class.1.year.6 == "yes" | DBHdata$Hollows.Class.1.year.6 == "Yes ", "Hollows.Class.1.year.6"] = "Yes"

# As "Hollows.Class.1.year.6" can't be converted to integer (not sure what value substitute "Yes"), further conversion for integer-representation: "0.0" to "0", "1.0" to "1", "2.0" to "2"
DBHdata[DBHdata$Hollows.Class.1.year.6 == "0.0", "Hollows.Class.1.year.6"] = "0"
DBHdata[DBHdata$Hollows.Class.1.year.6 == "1.0", "Hollows.Class.1.year.6"] = "1"
DBHdata[DBHdata$Hollows.Class.1.year.6 == "2.0", "Hollows.Class.1.year.6"] = "2"

sapply(DBHdata[10:15], unique)
```

check NAs and blank
```{r}
# blank "Tree.Number"
DBHdata[DBHdata$Tree.Number == "", ]
DBHdata = DBHdata[DBHdata$Tree.Number != "", ] # doesn't have any other information, remove rows
```
```{r}
# blank "Large"
DBHdata[DBHdata$Large == "", "Tree.Number"]
DBHdata[DBHdata$Large == "", "Large"] = "No" # assume "No", where there's no "Yes"
```
```{r}
# blank "Alive.at.year.6"
DBHdata[DBHdata$Alive.at.year.6 == "", 1:7]
# 20 rows with blank "Alive.at.year.6" have valid values for "DBH.year.6", assume they have "Yes" for "Alive.at.year.6"
DBHdata[DBHdata$Alive.at.year.6 == "" & !(is.na(DBHdata$DBH.year.6)), "Alive.at.year.6"] = "Yes"

# the rest rows with blank "Alive.at.year.6" have NA for "DBH.year.6", not sure whether the trees are not alive, or just missing records in year6, so no modification.
```
```{r}
# NA in "DBH.year.6", no modification
DBHdata[is.na(DBHdata$DBH.year.6), 1:7]

# 15 NAs for "DBH.year.6", 14 rows are the rows where "Alive.at.year.6" is blank, and one exception.
```
```{r}
# NA in "DBH.year.0", no modification
DBHdata[is.na(DBHdata$DBH.year.0), 1:7] # 7 NAs for "DBH.year.0"
```
```{r}
# assign 0 for NA in the 6 hollow-columns
DBHdata[DBHdata$Hollows.Class.1.year.0=="", "Hollows.Class.1.year.0"]="0"
DBHdata[DBHdata$Hollows.Class.2.year.0=="", "Hollows.Class.2.year.0"]="0"
DBHdata[DBHdata$Hollows.Class.3.year.0=="", "Hollows.Class.3.year.0"]="0"
DBHdata[DBHdata$Hollows.Class.1.year.6=="", "Hollows.Class.1.year.6"]="0"
DBHdata[is.na(DBHdata$Hollows.Class.2.year.6), "Hollows.Class.2.year.6"] = 0
DBHdata[is.na(DBHdata$Hollows.Class.3.year.6), "Hollows.Class.3.year.6"] = 0
sapply(DBHdata[10:15], unique)
```

assign correct data type for hollow columns
```{r}
DBHdata[c(10,11,12,14,15)] = sapply(DBHdata[c(10,11,12,14,15)], as.integer)
sapply(DBHdata, typeof)
```

prepare data for boxplot
```{r}
# prepare data for boxplot
DBHdata$Tree.growth = DBHdata$DBH.year.6 - DBHdata$DBH.year.0
T1data = DBHdata[DBHdata$Treatment == "T1: Gap", ]
T2data = DBHdata[DBHdata$Treatment == "T2: Radial", ]
Controldata = DBHdata[DBHdata$Treatment == "Control", ]
```






# basic analysis
boxplot DBH in year0 VS year6 for different plot in T1, T2, Control
```{r}
at = c(1, 3.5, 6, 8.5, 1.5, 4, 6.5, 9)
boxplot(T1data$DBH.year.0~T1data$Plot, at = c(1, 3.5, 6, 8.5), boxwex = 0.5, xlab = "T1: Gap plot number", ylab = "DBH in year0 VS year6", xlim = c(0, 10), ylim = c(15, 100), col = "light grey")
boxplot(T1data$DBH.year.6~T1data$Plot, add = TRUE, at = c(1.5, 4, 6.5, 9), boxwex = 0.5, xaxt = "n", col = "dark grey")
plotnum = c(1, 8, 10, 12)
for (i in c(1, 2, 3, 4)) {
  text(x = at[i]-0.5, y = fivenum(T1data[T1data$Plot == plotnum[i], "DBH.year.0"]), labels = fivenum(T1data[T1data$Plot == plotnum[i], "DBH.year.0"]), cex = 0.6)
  text(x = at[i+4]+0.5, y = fivenum(T1data[T1data$Plot == plotnum[i], "DBH.year.6"]), labels = fivenum(T1data[T1data$Plot == plotnum[i], "DBH.year.6"]), cex = 0.6)
}
legend("topright", legend = c("year0", "year6"), fill = c("light grey", "dark grey"))

boxplot(T2data$DBH.year.0~T2data$Plot, at = c(1, 3.5, 6, 8.5), boxwex = 0.5, xlab = "T2: Radial plot number", ylab = "DBH in year0 VS year6", xlim = c(0, 10), ylim = c(15, 100), col = "light grey")
boxplot(T2data$DBH.year.6~T2data$Plot, add = TRUE, at = c(1.5, 4, 6.5, 9), boxwex = 0.5, xaxt = "n", col = "dark grey")
plotnum = c(3, 5, 7, 9)
for (i in c(1, 2, 3, 4)) {
  text(x = at[i]-0.5, y = fivenum(T2data[T2data$Plot == plotnum[i], "DBH.year.0"]), labels = fivenum(T2data[T2data$Plot == plotnum[i], "DBH.year.0"]), cex = 0.6)
  text(x = at[i+4]+0.5, y = fivenum(T2data[T2data$Plot == plotnum[i], "DBH.year.6"]), labels = round(fivenum(T2data[T2data$Plot == plotnum[i], "DBH.year.6"]), 2), cex = 0.6)
}
legend("topright", legend = c("year0", "year6"), fill = c("light grey", "dark grey"))

boxplot(Controldata$DBH.year.0~Controldata$Plot, at = c(1, 3.5, 6, 8.5), boxwex = 0.5, xlab = "Control treatment plot number", ylab = "DBH in year0 VS year6", xlim = c(0, 10), ylim = c(15, 100), col = "light grey")
boxplot(Controldata$DBH.year.6~Controldata$Plot, add = TRUE, at = c(1.5, 4, 6.5, 9), boxwex = 0.5, xaxt = "n", col = "dark grey")
plotnum = c(2, 4, 6, 11)
for (i in c(1, 2, 3, 4)) {
  text(x = at[i]-0.5, y = fivenum(Controldata[Controldata$Plot == plotnum[i], "DBH.year.0"]), labels = round(fivenum(Controldata[Controldata$Plot == plotnum[i], "DBH.year.0"]), 2), cex = 0.6)
  text(x = at[i+4]+0.5, y = fivenum(Controldata[Controldata$Plot == plotnum[i], "DBH.year.6"]), labels = round(fivenum(Controldata[Controldata$Plot == plotnum[i], "DBH.year.6"]), 2), cex = 0.6)
}
legend("topright", legend = c("year0", "year6"), fill = c("light grey", "dark grey"))
```






# preliminary result
boxplot DBH in year0 VS year6 for T1, T2, Control
```{r}
at = c(1, 3.5, 6, 1.5, 4, 6.5)
boxplot(DBHdata$DBH.year.0~DBHdata$Treatment, boxwex = 0.5, at = c(1, 3.5, 6), xlab = "treatment", ylab = "DBH in year0 VS year6", xlim = c(0, 7.5), col = "light grey")
boxplot(DBHdata$DBH.year.6~DBHdata$Treatment, boxwex = 0.5, add = TRUE, at = c(1.5, 4, 6.5), xaxt = "n", col = "dark grey")
treatment = c("Control", "T1: Gap", "T2: Radial")
for (i in c(1, 2, 3)) {
  text(x = at[i]-0.5, y = fivenum(DBHdata[DBHdata$Treatment == treatment[i], "DBH.year.0"]), labels = fivenum(DBHdata[DBHdata$Treatment == treatment[i], "DBH.year.0"]), cex = 0.6)
  text(x = at[i+3]+0.5, y = fivenum(DBHdata[DBHdata$Treatment == treatment[i], "DBH.year.6"]), labels = round(fivenum(DBHdata[DBHdata$Treatment == treatment[i], "DBH.year.6"]), 2), cex = 0.6)
}
legend("topleft", legend = c("year0", "year6"), fill = c("light grey", "dark grey"))
```

boxplot tree growth (DBH in year6 - DBH in year0) for T1, T2, Control
```{r}
at = c(1, 2.2, 3.4)
treatment = c("Control", "T1: Gap", "T2: Radial")

boxplot(DBHdata$Tree.growth~DBHdata$Treatment, at = c(1, 2.2, 3.4), xlab = "treatment", ylab = "tree growth = DBHyear6 - DBHyear0", xlim = c(0.5, 4), col = "light grey")
for (i in c(1, 2, 3)) {
  text(x = at[i]+0.5, y = fivenum(DBHdata[DBHdata$Treatment == treatment[i], "Tree.growth"]), labels = round(fivenum(DBHdata[DBHdata$Treatment == treatment[i], "Tree.growth"]), 2), cex = 0.6)
}

boxplot(DBHdata$Tree.growth~DBHdata$Treatment, at = c(1, 2.2, 3.4), xlab = "treatment", ylab = "tree growth = DBHyear6 - DBHyear0 (zoom in)", xlim = c(0.5, 4), ylim = c(-3, 8), col = "light grey")
for (i in c(1, 2, 3)) {
  text(x = at[i]+0.5, y = fivenum(DBHdata[DBHdata$Treatment == treatment[i], "Tree.growth"]), labels = round(fivenum(DBHdata[DBHdata$Treatment == treatment[i], "Tree.growth"]), 2), cex = 0.6)
}
```

boxplot DBH for large trees VS other trees in T1, T2, Control
```{r}
boxplot(T1data$DBH.year.0~T1data$Large, at = c(1, 2.5), boxwex = 0.5, xlab = "T1: Gap Large tree", ylab = "DBH in year0 VS year6", xlim = c(0.5, 3.5), ylim = c(15, 70), col = "light grey")
boxplot(T1data$DBH.year.6~T1data$Large, boxwex = 0.5, at = c(1.5, 3), add = TRUE, col = "dark grey")
text(x = 0.65, y = fivenum(T1data[T1data$Large == "No", "DBH.year.0"]), labels = fivenum(T1data[T1data$Large == "No", "DBH.year.0"]), cex = 0.6)
text(x = 1.85, y = fivenum(T1data[T1data$Large == "No", "DBH.year.6"]), labels = fivenum(T1data[T1data$Large == "No", "DBH.year.6"]), cex = 0.6)
text(x = 2.15, y = fivenum(T1data[T1data$Large == "Yes", "DBH.year.0"]), labels = fivenum(T1data[T1data$Large == "Yes", "DBH.year.0"]), cex = 0.6)
text(x = 3.35, y = fivenum(T1data[T1data$Large == "Yes", "DBH.year.6"]), labels = fivenum(T1data[T1data$Large == "Yes", "DBH.year.6"]), cex = 0.6)
legend("topleft", legend = c("year0", "year6"), fill = c("light grey", "dark grey"))

boxplot(T2data$DBH.year.0~T2data$Large, at = c(1, 2.5), boxwex = 0.5, xlab = "T2: Radial Large tree", ylab = "DBH in year0 VS year6", xlim = c(0.5, 3.5), col = "light grey")
boxplot(T2data$DBH.year.6~T2data$Large, boxwex = 0.5, at = c(1.5, 3), add = TRUE, col = "dark grey")
text(x = 0.65, y = fivenum(T2data[T2data$Large == "No", "DBH.year.0"]), labels = fivenum(T2data[T2data$Large == "No", "DBH.year.0"]), cex = 0.6)
text(x = 1.85, y = fivenum(T2data[T2data$Large == "No", "DBH.year.6"]), labels = fivenum(T2data[T2data$Large == "No", "DBH.year.6"]), cex = 0.6)
text(x = 2.15, y = fivenum(T2data[T2data$Large == "Yes", "DBH.year.0"]), labels = fivenum(T2data[T2data$Large == "Yes", "DBH.year.0"]), cex = 0.6)
text(x = 3.35, y = fivenum(T2data[T2data$Large == "Yes", "DBH.year.6"]), labels = round(fivenum(T2data[T2data$Large == "Yes", "DBH.year.6"]), 2), cex = 0.6)
legend("topleft", legend = c("year0", "year6"), fill = c("light grey", "dark grey"))

boxplot(Controldata$DBH.year.0~Controldata$Large, at = c(1, 2.5), boxwex = 0.5, xlab = "Control treatment Large tree", ylab = "DBH in year0 VS year6", xlim = c(0.5, 3.5), ylim = c(15, 70), col = "light grey")
boxplot(Controldata$DBH.year.6~Controldata$Large, boxwex = 0.5, at = c(1.5, 3), add = TRUE, col = "dark grey")
text(x = 0.65, y = fivenum(Controldata[Controldata$Large == "No", "DBH.year.0"]), labels = fivenum(Controldata[Controldata$Large == "No", "DBH.year.0"]), cex = 0.6)
text(x = 1.85, y = fivenum(Controldata[Controldata$Large == "No", "DBH.year.6"]), labels = fivenum(Controldata[Controldata$Large == "No", "DBH.year.6"]), cex = 0.6)
text(x = 2.15, y = fivenum(Controldata[Controldata$Large == "Yes", "DBH.year.0"]), labels = round(fivenum(Controldata[Controldata$Large == "Yes", "DBH.year.0"]), 2), cex = 0.6)
text(x = 3.35, y = fivenum(Controldata[Controldata$Large == "Yes", "DBH.year.6"]), labels = fivenum(Controldata[Controldata$Large == "Yes", "DBH.year.6"]), cex = 0.6)
legend("topleft", legend = c("year0", "year6"), fill = c("light grey", "dark grey"))
```

boxplot tree growth (DBH in year6 - DBH in year0) for large trees VS other trees in T1, T2, Control
```{r}
boxplot(T1data$Tree.growth~T1data$Large, xlab = "Large tree", ylab = "tree growth = DBHyear6 - DBHyear0 (zoom in)", ylim = c(-3, 8), col = "light blue", at = c(0.5, 1), boxwex = 0.5, xlim = c(0, 5))
boxplot(T2data$Tree.growth~T2data$Large, ylim = c(-3, 8), col = "pink", at = c(2, 2.5), boxwex = 0.5, add = TRUE)
boxplot(Controldata$Tree.growth~Controldata$Large, ylim = c(-3, 8), col = "light grey", at = c(3.5, 4), boxwex = 0.5, add = TRUE)

text(x = 0.15, y = fivenum(T1data[T1data$Large == "No", "Tree.growth"]), labels = round(fivenum(T1data[T1data$Large == "No", "Tree.growth"]), 2), cex = 0.6)
text(x = 1.35, y = fivenum(T1data[T1data$Large == "Yes", "Tree.growth"]), labels = fivenum(T1data[T1data$Large == "Yes", "Tree.growth"]), cex = 0.6)

text(x = 1.65, y = fivenum(T2data[T2data$Large == "No", "Tree.growth"]), labels = round(fivenum(T2data[T2data$Large == "No", "Tree.growth"]), 2), cex = 0.6)
text(x = 2.85, y = fivenum(T2data[T2data$Large == "Yes", "Tree.growth"]), labels = round(fivenum(T2data[T2data$Large == "Yes", "Tree.growth"]), 2), cex = 0.6)

text(x = 3.15, y = fivenum(Controldata[Controldata$Large == "No", "Tree.growth"]), labels = round(fivenum(Controldata[Controldata$Large == "No", "Tree.growth"]), 2), cex = 0.6)
text(x = 4.35, y = fivenum(Controldata[Controldata$Large == "Yes", "Tree.growth"]), labels = round(fivenum(Controldata[Controldata$Large == "Yes", "Tree.growth"]), 2), cex = 0.6)

legend("topright", legend = c("T1: Gap", "T2: Radial", "Control"), fill = c("light blue", "pink", "light grey"))
```






# other plots
boxplot tree growth (DBH in year6 - DBH in year0) for different plots in T1, T2, Control
```{r}
boxplot(T1data$DBH.year.6-T1data$DBH.year.0~T1data$Plot, xlab = "plot number", ylab = "tree growth = DBHyear6 - DBHyear0 (zoom in)", ylim = c(-3, 8), xlim = c(0, 8), at = c(0.5, 1, 1.5, 2), boxwex = 0.45, col = "light blue")

boxplot(T2data$DBH.year.6-T2data$DBH.year.0~T2data$Plot, ylim = c(-3, 8), at = c(3, 3.5, 4, 4.5), boxwex = 0.45, add = TRUE, col = "pink")

boxplot(Controldata$DBH.year.6-Controldata$DBH.year.0~Controldata$Plot, ylim = c(-3, 8), at = c(5.5, 6, 6.5, 7), boxwex = 0.45, add = TRUE, col = "light grey")
legend("topright", legend = c("T1: Gap", "T2: Radial", "Control"), fill = c("light blue", "pink", "light grey"))
```

```{r}
# Skip this, it will output a csv!!!!!!!!!!!!
# output sub-csv from original data "all_plots.csv", which contains non-zero hollows
write.csv(DBHdata[DBHdata$Hollows.Class.1.year.6 != "0" | DBHdata$Hollows.Class.1.year.0 > 0 | DBHdata$Hollows.Class.2.year.0 > 0 | DBHdata$Hollows.Class.3.year.0 > 0 | DBHdata$Hollows.Class.2.year.6 > 0 | DBHdata$Hollows.Class.3.year.6 > 0, ], "nonzero hollows.csv", row.names = FALSE)
# Skip this, it will output a csv!!!!!!!!!!!!
```






# further analysis
# relative tree growth
relative growth = tree growth/DBHyear0*100, plot relative growth for each treatment, each plot, large VS other trees
```{r}
DBHdata$Relative.growth = DBHdata$Tree.growth/DBHdata$DBH.year.0*100
T1data = DBHdata[DBHdata$Treatment == "T1: Gap", ]
T2data = DBHdata[DBHdata$Treatment == "T2: Radial", ]
Controldata = DBHdata[DBHdata$Treatment == "Control", ]

summary(DBHdata$Relative.growth)
DBHdata[is.na(DBHdata$Tree.growth) | is.na(DBHdata$DBH.year.0), ]
```

relative growth for each plot
```{r}
boxplot(DBHdata$Relative.growth~DBHdata$Plot, xlab = "plot number", ylab = "relative tree growth(%)")

boxplot(T1data$Relative.growth~T1data$Plot, xlab = "plot number", ylab = "relative tree growth(%) (zoom in)", ylim = c(-10, 30), xlim = c(0, 8), at = c(0.5, 1, 1.5, 2), boxwex = 0.45, col = "light blue")
boxplot(T2data$Relative.growth~T2data$Plot, ylim = c(-10, 30), at = c(3, 3.5, 4, 4.5), boxwex = 0.45, add = TRUE, col = "pink")
boxplot(Controldata$Relative.growth~Controldata$Plot, ylim = c(-10, 30), at = c(5.5, 6, 6.5, 7), boxwex = 0.45, add = TRUE, col = "light grey")
legend("topright", legend = c("T1: Gap", "T2: Radial", "Control"), fill = c("light blue", "pink", "light grey"))
```

relative growth for each treatment
```{r}
boxplot(DBHdata$Relative.growth~DBHdata$Treatment, xlab = "treatment", ylab = "relative tree growth(%)")
boxplot(DBHdata$Relative.growth~DBHdata$Treatment, ylim = c(-10, 30), xlab = "treatment", ylab = "relative tree growth(%) (zoom in)")
```

relative growth for large VS other trees
```{r}
boxplot(T1data$Relative.growth~T1data$Large, xlab = "Large tree", ylab = "relative tree growth(%) (zoom in)", ylim = c(-10, 30), col = "light blue", at = c(0.5, 1), boxwex = 0.5, xlim = c(0, 5))
boxplot(T2data$Relative.growth~T2data$Large, ylim = c(-10, 30), col = "pink", at = c(2, 2.5), boxwex = 0.5, add = TRUE)
boxplot(Controldata$Relative.growth~Controldata$Large, ylim = c(-10, 30), col = "light grey", at = c(3.5, 4), boxwex = 0.5, add = TRUE)

text(x = 0.15, y = fivenum(T1data[T1data$Large == "No", "Relative.growth"]), labels = round(fivenum(T1data[T1data$Large == "No", "Relative.growth"]), 2), cex = 0.6)
text(x = 1.35, y = fivenum(T1data[T1data$Large == "Yes", "Relative.growth"]), labels = round(fivenum(T1data[T1data$Large == "Yes", "Relative.growth"]), 2), cex = 0.6)

text(x = 1.65, y = fivenum(T2data[T2data$Large == "No", "Relative.growth"]), labels = round(fivenum(T2data[T2data$Large == "No", "Relative.growth"]), 2), cex = 0.6)
text(x = 2.85, y = fivenum(T2data[T2data$Large == "Yes", "Relative.growth"]), labels = round(fivenum(T2data[T2data$Large == "Yes", "Relative.growth"]), 2), cex = 0.6)

text(x = 3.15, y = fivenum(Controldata[Controldata$Large == "No", "Relative.growth"]), labels = round(fivenum(Controldata[Controldata$Large == "No", "Relative.growth"]), 2), cex = 0.6)
text(x = 4.35, y = fivenum(Controldata[Controldata$Large == "Yes", "Relative.growth"]), labels = round(fivenum(Controldata[Controldata$Large == "Yes", "Relative.growth"]), 2), cex = 0.6)

legend("topright", legend = c("T1: Gap", "T2: Radial", "Control"), fill = c("light blue", "pink", "light grey"))
```

Finding:
T2 has significant effect in increasing relative tree growth for large trees, while relative tree growth for large trees in T1 and Control seems to be less than not-large trees. Comparing to Tree.growth in the previous plots, relative tree growth has very different pattern.