---
title: "Understorey_richness"
author: "Juan Dai"
date: "2022-09-18"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/Desktop/capstone/analysis/Juan/Understorey_priority2&low")
library(MASS)
library(survival)
library(fitdistrplus)
library(lme4)
```

poisson, negative binomial

```{r data_loading, echo=TRUE}
richness_df <- read.csv('./dataset/tables/species_richness_quadrats.csv')
richness_df_plot <- read.csv('./contingency tables/output/richness/species_richness_plots.csv')

# Treatments, Plot numbers, Quadrat numbers, Life forms as factors
richness_df$Treatment <- factor(richness_df$Treatment)
richness_df$Plot.Number <- factor(richness_df$Plot.Number)
richness_df$Quadrat.Number <- factor(richness_df$Quadrat.Number)

# Convert Fenced, Gap columns to boolean
richness_df$Gap <- richness_df$Gap=='True'
richness_df$Fenced <- richness_df$Fenced=='True'

richness_df_plot$Gap <- richness_df_plot$Gap=='True'
richness_df_plot$Fenced <- richness_df_plot$Fenced=='True'

# Remove null rows
richness_df <- na.omit(richness_df)
#richness_df_plot <- na.omit(richness_df_plot) 
# what is the purpose of the richness_df_plot
```

## Data preparation
```{r data, echo=TRUE}
richness_df$Y0.Y3.Difference <- richness_df$X3 - richness_df$X0
richness_df$Y3.Y6.Difference <- richness_df$X6 - richness_df$X3
richness_df$Y0.Y6.Difference <- richness_df$X6 - richness_df$X0
richness_df$Y0.Y3.Abs_Difference <- richness_df$Y0.Y3.Difference + abs(min(richness_df$Y0.Y3.Difference)) + 1
richness_df$Y3.Y6.Abs_Difference <- richness_df$Y3.Y6.Difference + abs(min(richness_df$Y3.Y6.Difference)) + 1
richness_df$Y0.Y6.Abs_Difference <- richness_df$Y0.Y6.Difference + abs(min(richness_df$Y0.Y6.Difference)) + 1
```

# modelling
```{r 0to3_gaussian, echo=True}
# guassian modeling
# model using gaussian
Y0.Y3.guas.model=fitdist(richness_df$Y0.Y3.Difference, "norm")
Y0.Y3.guas.model
guas.mean=Y0.Y3.guas.model$estimate[1]
guas.sd =Y0.Y3.guas.model$estimate[2]
plot(Y0.Y3.guas.model)

n=length(richness_df$Y0.Y3.Difference)
samp.pct <- (1:n-0.5)/n
qqplot(x=qnorm(samp.pct,guas.mean, guas.sd), y=sort(richness_df$Y0.Y3.Difference), xlab = 'Theoretical quantiles', ylab = 'Empirical quantiles', main='Q-Q plot Guassian')
abline(0,1,col="red")

Y0.Y3.guass.model = glm(Y0.Y3.Difference ~  Fenced+ Treatment* Gap, family = gaussian, data = richness_df)
summary(Y0.Y3.guass.model)
par(mfrow = c(2,2))

plot(Y0.Y3.guass.model)
Y0.Y3.pure.guass.model <- glm(Y0.Y3.Difference ~ 1, family = gaussian, data = richness_df)
anova( Y0.Y3.pure.guass.model, Y0.Y3.guass.model, test = "F")

```
```{r model_selection_guassian, echo=TRUE}
Y0.Y3.guass.model_selected <-step(Y0.Y3.guass.model, scope=~.)
summary(Y0.Y3.guass.model_selected)
# only treatment and gap are significant at 5% significance level, all other factor is not significant in predicting the yr_richness_difference, AICdoes not change, stays at 1617
anova(Y0.Y3.guass.model_selected, Y0.Y3.pure.guass.model,test = "F")
# the selected model states that the gap and treatment is related to the richness, and exists interaction effect between treatment and Gap
#Pr(>F) = 2.339e-05
```

```{r glm_fit_gamma, echo=TRUE}
Y0.Y3.gamma.model1=fitdist(richness_df$Y0.Y3.Abs_Difference, "gamma")
Y0.Y3.gamma.model1
gamma.alpha=Y0.Y3.gamma.model1$estimate[1]
gamma.beta =Y0.Y3.gamma.model1$estimate[2]
plot(Y0.Y3.gamma.model1)

n=length(richness_df$Y0.Y3.Abs_Difference)
samp.pct <- (1:n-0.5)/n
qqplot(x=qgamma(samp.pct,gamma.alpha, gamma.beta), y=sort(richness_df$Y0.Y3.Abs_Difference), xlab = 'Theoretical quantiles', ylab = 'Empirical quantiles', main='Q-Q plot gamma')
abline(0,1,col="red")


Y0.Y3.gamma.model2 = glm(Y0.Y3.Abs_Difference ~ Treatment * Gap +  Fenced, family = Gamma, data = richness_df)
summary(Y0.Y3.gamma.model2)
par(mfrow = c(2,2))
plot(Y0.Y3.gamma.model2)
```

```{r model_selection_gamma, echo=TRUE}
Y0.Y3.gamma.model_selected <-step(Y0.Y3.gamma.model2, scope=~.)
summary(Y0.Y3.gamma.model_selected)
# AIC does not change, stays at 1651
Y0.Y3.empty_gamma_model <- glm(Y0.Y3.Abs_Difference ~ 1, family = Gamma, data = richness_df)
anova(Y0.Y3.gamma.model_selected, Y0.Y3.empty_gamma_model, test = "F")
# the model is significant at 5%, suggest that the coefficient for treatment and gap, fenced is not zero, Pr(>F) = 0.0001253
```

```{r fit_poisson_model, echo=TRUE}
Y0.Y3.poisson.model1=fitdist(richness_df$Y0.Y3.Abs_Difference, "pois")
Y0.Y3.poisson.model1
poisson.lambda=Y0.Y3.poisson.model1$estimate
plot(Y0.Y3.poisson.model1)

n=length(richness_df$Y0.Y3.Abs_Difference)
samp.pct <- (1:n-0.5)/n
qqplot(x=qpois(samp.pct,poisson.lambda), y=sort(richness_df$Y0.Y3.Abs_Difference), xlab = 'Theoretical quantiles', ylab = 'Empirical quantiles', main='Q-Q plot Poisson')
abline(0,1,col="red")



Y0.Y3.poisson.model2 = glm(Y0.Y3.Abs_Difference~ Treatment* Gap +  Fenced, family = poisson, data = richness_df)
summary(Y0.Y3.poisson.model2)
par(mfrow = c(2,2))
plot(Y0.Y3.poisson.model2)
```

```{r model_selection_poisson, echo=TRUE}
Y0.Y3.poisson.model_selected <-step(Y0.Y3.poisson.model2, scope=~.)
summary(Y0.Y3.poisson.model_selected)
#AIC decrease from 1662
Y0.Y3.empty_poisson_model <- glm(Y0.Y3.Abs_Difference ~ 1, family = poisson, data = richness_df)
anova(Y0.Y3.poisson.model_selected, Y0.Y3.empty_poisson_model, test = "Chi")
# the model is significant at 5%, suggest that the coefficient for fenced, treatment and gap is not zero, Pr(>F) = 3.19e-09
```

```{r fit_negative_binomial, echo=TRUE}
#Negative Binomial GLM

#visualisation and test if the data is negative binomial
Y0.Y3.nb.model1=fitdist(richness_df$Y0.Y3.Abs_Difference, "nbinom")
Y0.Y3.nb.model1
nb.size=Y0.Y3.nb.model1$estimate[1]
nb.mu = Y0.Y3.nb.model1$estimate[2]
nb.prob=nb.size/(nb.size+nb.mu) 
plot(Y0.Y3.nb.model1)

n=length(richness_df$Y0.Y3.Abs_Difference)
samp.pct <- (1:n-0.5)/n
qqplot(x=qnbinom(samp.pct,nb.size, nb.prob), y=sort(richness_df$Y0.Y3.Abs_Difference), xlab = 'Theoretical quantiles', ylab = 'Empirical quantiles', main='Q-Q plot Negative Binomial')
abline(0,1,col="red")
#fit the model
Y0.Y3.nb.model2 = glm.nb(Y0.Y3.Abs_Difference~ Treatment* Gap +  Fenced, data = richness_df)
summary(Y0.Y3.nb.model2 )
par(mfrow = c(2,2))
plot(Y0.Y3.nb.model2)
```

```{r model_selection_negative_binomial, echo=TRUE}
Y0.Y3.nb.model_selected <-step(Y0.Y3.nb.model2, scope=~.)
summary(Y0.Y3.nb.model_selected)
# AIC decrease from 1625
Y0.Y3.empty_nb_model <- glm.nb(Y0.Y3.Abs_Difference ~ 1,  data = richness_df)
anova(Y0.Y3.nb.model_selected, Y0.Y3.empty_nb_model, test = "LR")
# the model is significant at 5%, suggest that the coefficient for treatment and gap is not zero, Pr(>z) = 9.075586e-05
```

**selected based on AIC values, we will choose the gaussian model, because it gives lowest AIC value of 1617.**



## diffrent modeling via prediction
```{r model_prediction_gaussian, echo=TRUE}
Y0.Y3.guass.model2 = glm(X3 ~ X0 + Treatment* Gap + Fenced, family = gaussian, data = richness_df)
summary(Y0.Y3.guass.model2)
par(mfrow = c(2,2))
plot(Y0.Y3.guass.model2)

Y0.Y3.gamma.model3 = glm(X3 ~ X0 + Treatment* Gap + Fenced, family = Gamma, data = richness_df)
summary(Y0.Y3.gamma.model3)
par(mfrow = c(2,2))

plot(Y0.Y3.gamma.model3)

Y0.Y3.poisson.model3 = glm(X3 ~ X0 + Treatment* Gap + Fenced, family = poisson, data = richness_df)
summary(Y0.Y3.poisson.model3)
par(mfrow = c(2,2))

plot(Y0.Y3.poisson.model3)
```

```{r model_selection_prediction_gaussian, echo=TRUE}
Y0.Y3.select_guass.model2 <- step(Y0.Y3.guass.model2, scope=~.)
summary(Y0.Y3.select_guass.model2)

Y0.Y3.select_gamma.model3 <- step(Y0.Y3.gamma.model3, scope=~.)
summary(Y0.Y3.select_gamma.model3)

Y0.Y3.select_poisson.model3 <- step(Y0.Y3.poisson.model3, scope=~.)
summary(Y0.Y3.select_poisson.model3)
# according to the AIC, we will choose the poisson model, since it has the low AIC of 1489.7, but it has better dispersion paramter is 1.
```
# select the poisson model with only X0, Treatment and Gap


**Therefore, using the prediction with poisson model, it has the lowest AIC of 1493.1**
```{r model_selection_prediction_gaussian_log_glmer, echo=TRUE}
Y0.Y3.glmer.poiss1= glmer(X3 ~ X0 + Treatment * Gap + (1|Plot.Number), family = poisson, data = richness_df)
summary(Y0.Y3.glmer.poiss1)
print(Y0.Y3.glmer.poiss1, correlation=TRUE)

#diagnosis plots
plot(Y0.Y3.glmer.poiss1)
plot(Y0.Y3.glmer.poiss1, resid(., scaled=TRUE) ~ fitted(.), abline = 0,pch=16,col=richness_df$Plot.Number,xlab="Fitted values",ylab="Standardised residuals")


qqnorm(resid(Y0.Y3.glmer.poiss1),pch=16,col=richness_df$Plot.Number)
qqline(resid(Y0.Y3.glmer.poiss1))


anova(Y0.Y3.glmer.poiss1, Y0.Y3.select_poisson.model3)
#(1 | Plot.Number) is not significant, Pr(>Chisq) = 0.5524

Y0.Y3.glmer.poiss2= glmer(X3 ~ X0 + Treatment * Gap  +(X0|Plot.Number), family = poisson, data = richness_df)
summary(Y0.Y3.glmer.poiss2)
par(mfrow = c(2,2))
plot(Y0.Y3.glmer.poiss2)
anova(Y0.Y3.glmer.poiss2,Y0.Y3.select_poisson.model3,  test = "Chi")
#(X0|Plot.Number) is not significant, Pr(>Chisq) = 0.1778

Y0.Y3.glmer.poiss3= glmer(X3 ~ X0 + Treatment * Gap +  Fenced + (1|Quadrat.Number), family = poisson, data = richness_df)
summary(Y0.Y3.glmer.poiss3)
print(Y0.Y3.glmer.poiss3, correlation=TRUE)
par(mfrow = c(2,2))
plot(Y0.Y3.glmer.poiss3)
anova(Y0.Y3.glmer.poiss3, Y0.Y3.select_poisson.model3)
# (1 | Quadrat.Number) is not significant, Pr(>Chisq) = 0.7585

# according to the Chisq, we will choose the poisson model without the random effect, since all the anova states no random effect,

```
From year 0 to year 3, we may not use the random effect.





##Yr3 to Yr6
# modelling
```{r gaussian_Y36, echo=True}
# guassian modeling
# model using gaussian
Y3.Y6.guas.model=fitdist(richness_df$Y3.Y6.Difference, "norm")
Y3.Y6.guas.model
guas.mean=Y3.Y6.guas.model$estimate[1]
guas.sd =Y3.Y6.guas.model$estimate[2]
plot(Y3.Y6.guas.model)

n=length(richness_df$Y3.Y6.Difference)
samp.pct <- (1:n-0.5)/n
qqplot(x=qnorm(samp.pct,guas.mean, guas.sd), y=sort(richness_df$Y3.Y6.Difference), xlab = 'Theoretical quantiles', ylab = 'Empirical quantiles', main='Q-Q plot Guassian')
abline(0,1,col="red")

Y3.Y6.guass.model = glm(Y3.Y6.Difference ~ Treatment * Gap  + Fenced, family = gaussian, data = richness_df)
summary(Y3.Y6.guass.model)
par(mfrow = c(2,2))

plot(Y3.Y6.guass.model)
Y3.Y6.pure.guass.model <- glm(Y3.Y6.Difference ~ 1, family = gaussian, data = richness_df)
anova(Y3.Y6.pure.guass.model, Y3.Y6.guass.model, test = "F")
# In model fitting, fenced, gap, and the interactions between treatment and gap are statistically significant, also when we use F-stats to test the H0: all the coefficient is zero, it states that we reject null at 5% significance level
# Pr(>F) = 4.811e-08
```


```{r model_selection_guassian_Y36, echo=TRUE}
Y3.Y6.guass.model_selected <-step(Y3.Y6.guass.model, scope=~.)
summary(Y3.Y6.guass.model_selected)
# only treatment and gap is significant, all other factor is not significant in predicting the yr_richness_difference, AIC stays at 1637.7
anova(Y3.Y6.guass.model_selected, Y3.Y6.pure.guass.model,test = "F")
# the selected model states that the gap and fenced and the interaction between gap and treatment is related to the richness
#Pr(>F) = 4.811e-08
```

```{r glm_fit_gamma_Y36, echo=TRUE}
Y3.Y6.gamma.model1=fitdist(richness_df$Y3.Y6.Abs_Difference, "gamma")
Y3.Y6.gamma.model1
gamma.alpha=Y3.Y6.gamma.model1$estimate[1]
gamma.beta =Y3.Y6.gamma.model1$estimate[2]
plot(Y3.Y6.gamma.model1)

n=length(richness_df$Y3.Y6.Abs_Difference)
samp.pct <- (1:n-0.5)/n
qqplot(x=qgamma(samp.pct,gamma.alpha, gamma.beta), y=sort(richness_df$Y3.Y6.Abs_Difference), xlab = 'Theoretical quantiles', ylab = 'Empirical quantiles', main='Q-Q plot gamma')
abline(0,1,col="red")


Y3.Y6.gamma.model2 = glm(Y3.Y6.Abs_Difference ~ Treatment * Gap  + Fenced, family = Gamma, data = richness_df)
summary(Y3.Y6.gamma.model2)
par(mfrow = c(2,2))
plot(Y3.Y6.gamma.model2)
# the model fit is worse than the gaussian fit, from the qq-plot
```

```{r model_selection_gamma_Y36, echo=TRUE}
Y3.Y6.gamma.model_selected <-step(Y3.Y6.gamma.model2, scope=~.)
summary(Y3.Y6.gamma.model_selected)
# AIC stays at 1679.9
Y3.Y6.empty_gamma_model <- glm(Y3.Y6.Abs_Difference ~ 1, family = Gamma, data = richness_df)
anova(Y3.Y6.gamma.model_selected, Y3.Y6.empty_gamma_model, test = "F")
# the model is significant at 5%, suggest that the coefficient for gap, fenced and the interaction between treatment and gap are not zero, Pr(>F) = 1.957e-09
```

```{r fit_poisson_model_Y36, echo=TRUE}
Y3.Y6.poisson.model1=fitdist(richness_df$Y3.Y6.Abs_Difference, "pois")
Y3.Y6.poisson.model1
poisson.lambda=Y3.Y6.poisson.model1$estimate
plot(Y3.Y6.poisson.model1)

n=length(richness_df$Y3.Y6.Abs_Difference)
samp.pct <- (1:n-0.5)/n
qqplot(x=qpois(samp.pct,poisson.lambda), y=sort(richness_df$Y3.Y6.Abs_Difference), xlab = 'Theoretical quantiles', ylab = 'Empirical quantiles', main='Q-Q plot Poisson')
abline(0,1,col="red")



Y3.Y6.poisson.model2 = glm(Y3.Y6.Abs_Difference~ Treatment * Gap  + Fenced, family = poisson, data = richness_df)
summary(Y3.Y6.poisson.model2)
par(mfrow = c(2,2))
plot(Y3.Y6.poisson.model2)
```

```{r model_selection_poisson_Y36, echo=TRUE}
Y3.Y6.poisson.model_selected <-step(Y3.Y6.poisson.model2, scope=~.)
summary(Y3.Y6.poisson.model_selected)
#AIC stays at 1561.1
Y3.Y6.empty_poisson_model <- glm(Y3.Y6.Abs_Difference ~ 1, family = poisson, data = richness_df)
anova(Y3.Y6.poisson.model_selected, Y3.Y6.empty_poisson_model, test = "Chi")
# the model is significant at 5%, suggest that the coefficient for gap, fenced and interaction between tratemnt and gap are not zero, Pr(>Chi) = < 2.2e-16 
```

**we still choose poisson model, since it has lowest AIC of 1561.1**

## diffrent modeling via prediction for yr3 to yr6
```{r model_prediction_gaussian_Y36, echo=TRUE}
Y3.Y6.guass.model2 = glm(X6 ~ X3 + Treatment  * Gap  + Fenced, family = gaussian, data = richness_df)
summary(Y3.Y6.guass.model2)
par(mfrow = c(2,2))
plot(Y3.Y6.guass.model2)

Y3.Y6.gamma.model3 = glm(X6 ~ X3 + Treatment * Gap  + Fenced, family = Gamma, data = richness_df)
summary(Y3.Y6.gamma.model3)
par(mfrow = c(2,2))

plot(Y3.Y6.gamma.model3)

Y3.Y6.poisson.model3 = glm(X6 ~ X3 + Treatment * Gap + + Fenced, family = poisson, data = richness_df)
summary(Y3.Y6.poisson.model3)
par(mfrow = c(2,2))

plot(Y3.Y6.poisson.model3)
```

```{r model_selection_prediction_gaussian_Y36, echo=TRUE}
Y3.Y6.select_guass.model2 <- step(Y3.Y6.guass.model2, scope=~.)
summary(Y3.Y6.select_guass.model2)

Y3.Y6.select_gamma.model3 <- step(Y3.Y6.gamma.model3, scope=~.)
summary(Y3.Y6.select_gamma.model3)

Y3.Y6.select_poisson.model3 <- step(Y3.Y6.poisson.model3, scope=~.)
summary(Y3.Y6.select_poisson.model3)

# if we use the AIC to choose the model, we will choose the gamma model, it yields the relatively AIC of 1535.6 and dispersion parameter as 1
```


##Yr0 to Yr6
# modelling
```{r gaussian_Y06, echo=True}
# guassian modeling
# model using gaussian
Y0.Y6.guas.model=fitdist(richness_df$Y0.Y6.Difference, "norm")
Y0.Y6.guas.model
guas.mean=Y0.Y6.guas.model$estimate[1]
guas.sd =Y0.Y6.guas.model$estimate[2]
plot(Y0.Y6.guas.model)

n=length(richness_df$Y0.Y6.Difference)
samp.pct <- (1:n-0.5)/n
qqplot(x=qnorm(samp.pct,guas.mean, guas.sd), y=sort(richness_df$Y0.Y6.Difference), xlab = 'Theoretical quantiles', ylab = 'Empirical quantiles', main = 'Q-Q plot Guassian')
abline(0,1,col="red")

Y0.Y6.guass.model = glm(Y0.Y6.Difference ~ Treatment * Gap  + Fenced, family = gaussian, data = richness_df)
summary(Y0.Y6.guass.model)
par(mfrow = c(2,2))

plot(Y0.Y6.guass.model)
Y0.Y6.pure.guass.model <- glm(Y0.Y6.Difference ~ 1, family = gaussian, data = richness_df)
anova(Y0.Y6.pure.guass.model, Y0.Y6.guass.model, test = "F")
# In model fitting, the intercept, gap and some quadrates number are statistically significant, but when we use F-stats to test the H0: all the coefficient is zero, it states that we reject null at 5% significance level
# Pr(>F) = 0.02073 
```


```{r model_selection_guassian_Y06, echo=TRUE}
Y0.Y6.guass.model_selected <-step(Y0.Y6.guass.model, scope=~.)
summary(Y0.Y6.guass.model_selected)
# only treatment radial is significant, all other factor is not significant in predicting the yr_richness_difference, AIC decreases from 1614.7 to 1610
anova(Y0.Y6.guass.model_selected, Y0.Y6.pure.guass.model,test = "F")
# the selected model states that the treatment and gap is related to the richness, although only treatment radial is significant
#Pr(>F) = 0.004403
```

```{r glm_fit_gamma_Y06, echo=TRUE}
Y0.Y6.gamma.model1=fitdist(richness_df$Y0.Y6.Abs_Difference, "gamma")
Y0.Y6.gamma.model1
gamma.alpha=Y0.Y6.gamma.model1$estimate[1]
gamma.beta =Y0.Y6.gamma.model1$estimate[2]
plot(Y0.Y6.gamma.model1)

n=length(richness_df$Y0.Y6.Abs_Difference)
samp.pct <- (1:n-0.5)/n
qqplot(x=qgamma(samp.pct,gamma.alpha, gamma.beta), y=sort(richness_df$Y0.Y6.Abs_Difference), xlab = 'Theoretical quantiles', ylab = 'Empirical quantiles', main='Q-Q plot gamma')
abline(0,1,col="red")


Y0.Y6.gamma.model2 = glm(Y0.Y6.Abs_Difference ~ Treatment  * Gap  + Fenced, family = Gamma, data = richness_df)
summary(Y0.Y6.gamma.model2)
par(mfrow = c(2,2))
plot(Y0.Y6.gamma.model2)
# the model fit is worse than the gaussian fit, from the qq-plot, it dispalys a curvature
```

```{r model_selection_gamma_Y06, echo=TRUE}
Y0.Y6.gamma.model_selected <-step(Y0.Y6.gamma.model2, scope=~.)
summary(Y0.Y6.gamma.model_selected)
# AIC decrease from 1666 to 1661
Y0.Y6.empty_gamma_model <- glm(Y0.Y6.Abs_Difference ~ 1, family = Gamma, data = richness_df)
anova(Y0.Y6.gamma.model_selected, Y0.Y6.empty_gamma_model, test = "F")
# the model is significant at 5%, suggest that the coefficient for treatment is not zero, Pr(>F) = 0.004019
```

```{r fit_poisson_model_Y06, echo=TRUE}
Y0.Y6.poisson.model1=fitdist(richness_df$Y0.Y6.Abs_Difference, "pois")
Y0.Y6.poisson.model1
poisson.lambda=Y0.Y6.poisson.model1$estimate
plot(Y0.Y6.poisson.model1)

n=length(richness_df$Y0.Y6.Abs_Difference)
samp.pct <- (1:n-0.5)/n
qqplot(x=qpois(samp.pct,poisson.lambda), y=sort(richness_df$Y0.Y6.Abs_Difference), xlab = 'Theoretical quantiles', ylab = 'Empirical quantiles', main='Q-Q plot Poisson')
abline(0,1,col="red")



Y0.Y6.poisson.model2 = glm(Y0.Y6.Abs_Difference~ Treatment * Gap + Fenced, family = poisson, data = richness_df)
summary(Y0.Y6.poisson.model2)
par(mfrow = c(2,2))
plot(Y0.Y6.poisson.model2)
```

```{r model_selection_poisson_Y06, echo=TRUE}
Y0.Y6.poisson.model_selected <-step(Y0.Y6.poisson.model2, scope=~.)
summary(Y0.Y6.poisson.model_selected)
#AIC decrease from 1664 to 1640
Y0.Y6.empty_poisson_model <- glm(Y0.Y6.Abs_Difference ~ 1, family = poisson, data = richness_df)
anova(Y0.Y6.poisson.model_selected, Y0.Y6.empty_poisson_model, test = "Chi")
# the model is significant at 5%, suggest that the coefficient for treatment is not zero at 5% significance level, Pr(>F) = 0.0003047
```

**Based on the AIC, we will choose guassian model, since it give the lowest AIC of 1610 compared to gamma and poisson**


## diffrent modeling via prediction for yr0 to yr6 (not include X3)
```{r model_prediction_gaussian_Y06, echo=TRUE}
Y0.Y6.guass.model2 = glm(X6 ~ X0 + Treatment * Gap  + Fenced, family = gaussian, data = richness_df)
summary(Y0.Y6.guass.model2)
par(mfrow = c(2,2))
plot(Y0.Y6.guass.model2)

Y0.Y6.gamma.model3 = glm(X6 ~ X0 + Treatment * Gap + Fenced, family = Gamma, data = richness_df)
summary(Y0.Y6.gamma.model3)
par(mfrow = c(2,2))

plot(Y0.Y6.gamma.model3)

Y0.Y6.poisson.model3 = glm(X6 ~ X0 + Treatment * Gap  + Fenced, family = poisson, data = richness_df)
summary(Y0.Y6.poisson.model3)
par(mfrow = c(2,2))

plot(Y0.Y6.poisson.model3)
```

```{r model_selection_prediction_gaussian_Y06, echo=TRUE}
Y0.Y6.select_guass.model2 <- step(Y0.Y6.guass.model2, scope=~.)
summary(Y0.Y6.select_guass.model2)

Y0.Y6.select_gamma.model3 <- step(Y0.Y6.gamma.model3, scope=~.)
summary(Y0.Y6.select_gamma.model3)

Y0.Y6.select_poisson.model3 <- step(Y0.Y6.poisson.model3, scope=~.)
summary(Y0.Y6.select_poisson.model3)

# if we use the AIC to choose the model, we will choose the poisson model, it yields the lowest AIC of 1513, in the poisson model, all the treatment, X0, Gap and Fence, and the interaction term between treatment and gap are significant in predicting the X6 richness at 5% significance level.
#both treatment factor are negative.
```

## diffrent modeling via prediction for yr0 to yr6 (include X3)
```{r model_prediction_gaussian_Y036, echo=TRUE}
Y0.Y6.guass.model3 = glm(X6 ~ X0 + X3 + Treatment * Gap  + Fenced, family = gaussian, data = richness_df)
summary(Y0.Y6.guass.model3)
par(mfrow = c(2,2))
plot(Y0.Y6.guass.model3)

Y0.Y6.gamma.model4 = glm(X6 ~ X0 + X3+ Treatment * Gap  + Fenced, family = Gamma, data = richness_df)
summary(Y0.Y6.gamma.model4)
par(mfrow = c(2,2))

plot(Y0.Y6.gamma.model4)

Y0.Y6.poisson.model4 = glm(X6 ~ X0 + X3 + Treatment * Gap  + Fenced, family = poisson, data = richness_df)
summary(Y0.Y6.poisson.model4)
par(mfrow = c(2,2))

plot(Y0.Y6.poisson.model4)
```

```{r model_selection_prediction_gaussian_Y036, echo=TRUE}
Y0.Y6.select_guass.model3 <- step(Y0.Y6.guass.model3, scope=~.)
summary(Y0.Y6.select_guass.model3)

Y0.Y6.select_gamma.model4 <- step(Y0.Y6.gamma.model4, scope=~.)
summary(Y0.Y6.select_gamma.model4)

Y0.Y6.select_poisson.model4 <- step(Y0.Y6.poisson.model4, scope=~.)
summary(Y0.Y6.select_poisson.model4)

# if we use the AIC to choose the model, we will choose the poisson model, it yields the lowest AIC of 1504, in the poisson model, both X0 and X3, treatment, Gap and Fence, and interaction between treatment and gap are significant in predicting the richness of X6 at 5% significance level.
#Both treatment state a negative effect on richness increase
```

```{r model_selection_prediction_gaussian_log_glmer, echo=TRUE}
Y0.Y6.glmer.poiss1= glmer(X6 ~ X0 + X3 + Treatment * Gap + Fenced+ (1|Plot.Number), family = poisson, data = richness_df)
summary(Y0.Y6.glmer.poiss1)
print(Y0.Y6.glmer.poiss1, correlation=TRUE)

#diagnosis plots
plot(Y0.Y6.glmer.poiss1)
plot(Y0.Y6.glmer.poiss1, resid(., scaled=TRUE) ~ fitted(.), abline = 0,pch=16,col=richness_df$Plot.Number,xlab="Fitted values",ylab="Standardised residuals")

qqnorm(resid(Y0.Y6.glmer.poiss1),pch=16,col=richness_df$Plot.Number)
qqline(resid(Y0.Y6.glmer.poiss1))


anova(Y0.Y6.glmer.poiss1, Y0.Y6.poisson.model4)
# (1|Plot.Number) is significant, p-value = 1.183e-11

Y0.Y6.glmer.poiss2= glmer(X6 ~ X0 + X3 + Treatment * Gap + Fenced + (1|Plot.Number) + (X0|Plot.Number), family = poisson, data = richness_df)
summary(Y0.Y6.glmer.poiss2)
par(mfrow = c(2,2))
plot(Y0.Y6.glmer.poiss2)
anova(Y0.Y6.glmer.poiss2,Y0.Y6.glmer.poiss1,  test = "Chi")
#(X0|Plot.Number) is not significant, p-value = 0.8019

Y0.Y6.glmer.poiss3= glmer(X6 ~ X0 + X3 + Treatment * Gap  + Fenced + (1|Plot.Number)+ (1|Quadrat.Number), family = poisson, data = richness_df)
summary(Y0.Y6.glmer.poiss3)
print(Y0.Y6.glmer.poiss3, correlation=TRUE)
par(mfrow = c(2,2))
plot(Y0.Y6.glmer.poiss3)
anova(Y0.Y6.glmer.poiss3, Y0.Y6.glmer.poiss1)
#(1|Quadrat.Number) is not significant, with p-value = 1

# according to the Chisq, we will choose the poisson model with the random effect of (1|plot.number)

```