---
title: "DBH models 2"
output: html_document
---

This file contains models using DBH.year.6 as response variable. It aims at all hypotheses except for the last one (- The rate of natural mortality will be greater in control plots as compared to T1 and T2.)



# data preparation
```{r}
setwd("C:/Users/YUHONG/Documents/R/DataScienceProject")
# load data
data = read.csv("datasets/all_plots.csv", stringsAsFactors = FALSE)
```
```{r}
summary(data)
sapply(data, typeof)
sapply(data[c(2,3,4,5,8,9,10,11,12,13,14,15)], unique)
```

only extract DBH-related data for DBH modelling
```{r}
DBHdata = data[1:9]
```

check NAs and blank
```{r}
# blank "Large", assume "No" where it's blank
DBHdata[DBHdata$Large == "", "Tree.Number"]
DBHdata[DBHdata$Large == "", "Large"] = "No"
```
```{r}
# remove rows with NA in "DBH.year.6"
DBHdata[is.na(DBHdata$DBH.year.6), 1:7] # 9 NAs for "DBH.year.6"
DBHdata = DBHdata[!is.na(DBHdata$DBH.year.6), ]
```
```{r}
# remove rows with NA in "DBH.year.0"
DBHdata[is.na(DBHdata$DBH.year.0), 1:7] # 2 NAs for "DBH.year.0"
DBHdata = DBHdata[!is.na(DBHdata$DBH.year.0), ]
```
```{r}
# NA in "Class.year.0" and NA in "Class.year.6", no modification
DBHdata[is.na(DBHdata$Class.year.0) | is.na(DBHdata$Class.year.6), ]
# both in Plot5(T2: Radial), Tree.Number 148, 149, they have extremely large DBH.
```

make Treatment and Large factor
```{r}
DBHdata$Treatment = factor(DBHdata$Treatment)
DBHdata$Large = factor(DBHdata$Large)
```

DBH.year.6 <= DBH.year.0
```{r}
DBHdata[DBHdata$DBH.year.6 <= DBHdata$DBH.year.0, ]
DBHpositive = DBHdata[DBHdata$DBH.year.6 > DBHdata$DBH.year.0, ]
```



# data analysis
```{r}
par(mfrow=c(2,2))
plot(DBHdata$DBH.year.0, DBHdata$DBH.year.6, type = "p")
plot(log(DBHdata$DBH.year.0), log(DBHdata$DBH.year.6), type = "p")
plot(DBHpositive$DBH.year.0, DBHpositive$DBH.year.6, type = "p")
plot(log(DBHpositive$DBH.year.0), log(DBHpositive$DBH.year.6), type = "p")

plot(DBHdata$DBH.year.0, DBHdata$DBH.year.6, type = "p", col = DBHdata$Treatment)
plot(log(DBHdata$DBH.year.0), log(DBHdata$DBH.year.6), type = "p", col = DBHdata$Treatment)
plot(DBHpositive$DBH.year.0, DBHpositive$DBH.year.6, type = "p", col = DBHpositive$Treatment)
plot(log(DBHpositive$DBH.year.0), log(DBHpositive$DBH.year.6), type = "p", col = DBHpositive$Treatment)

plot(DBHdata$DBH.year.0, DBHdata$DBH.year.6, type = "p", col = DBHdata$Large)
plot(log(DBHdata$DBH.year.0), log(DBHdata$DBH.year.6), type = "p", col = DBHdata$Large)
plot(DBHpositive$DBH.year.0, DBHpositive$DBH.year.6, type = "p", col = DBHpositive$Large)
plot(log(DBHpositive$DBH.year.0), log(DBHpositive$DBH.year.6), type = "p", col = DBHpositive$Large)
```
```{r}
# points outside the main group
plot(DBHdata$DBH.year.0, DBHdata$DBH.year.6/DBHdata$DBH.year.0, type = "p", col = factor(DBHdata$DBH.year.6/DBHdata$DBH.year.0 > 1.25 | DBHdata$DBH.year.6/DBHdata$DBH.year.0 < 0.9 | DBHdata$DBH.year.0 > 80))
plot(DBHdata$DBH.year.0, DBHdata$DBH.year.6, type = "p", col = factor(DBHdata$DBH.year.6/DBHdata$DBH.year.0 > 1.25 | DBHdata$DBH.year.6/DBHdata$DBH.year.0 < 0.9 | DBHdata$DBH.year.0 > 80))
DBHdata[DBHdata$DBH.year.6/DBHdata$DBH.year.0 > 1.25 | DBHdata$DBH.year.6/DBHdata$DBH.year.0 < 0.9 | DBHdata$DBH.year.0 > 80, ]

# remove these outliers
DBHdatarm = DBHdata[DBHdata$DBH.year.6/DBHdata$DBH.year.0 < 1.25 & DBHdata$DBH.year.6/DBHdata$DBH.year.0 > 0.9 & DBHdata$DBH.year.0 < 80, ]
DBHpositiverm = DBHpositive[DBHpositive$DBH.year.6/DBHpositive$DBH.year.0 < 1.25 & DBHpositive$DBH.year.0 < 80, ]
```
```{r}
cor(DBHdata$DBH.year.0, DBHdata$DBH.year.6)
cor(log(DBHdata$DBH.year.0), log(DBHdata$DBH.year.6))

cor(DBHpositive$DBH.year.0, DBHpositive$DBH.year.6)
cor(log(DBHpositive$DBH.year.0), log(DBHpositive$DBH.year.6))

cor(DBHdatarm$DBH.year.0, DBHdatarm$DBH.year.6)
cor(log(DBHdatarm$DBH.year.0), log(DBHdatarm$DBH.year.6))

cor(DBHpositiverm$DBH.year.0, DBHpositiverm$DBH.year.6)
cor(log(DBHpositiverm$DBH.year.0), log(DBHpositiverm$DBH.year.6))
```
```{r}
shapiro.test(DBHdata$DBH.year.6)
shapiro.test(log(DBHdata$DBH.year.6))

shapiro.test(DBHpositive$DBH.year.6)
shapiro.test(log(DBHpositive$DBH.year.6))

shapiro.test(DBHdatarm$DBH.year.6)
shapiro.test(log(DBHdatarm$DBH.year.6))

shapiro.test(DBHpositiverm$DBH.year.6)
shapiro.test(log(DBHpositiverm$DBH.year.6)) # closest to normal
```
```{r}
library(car)
leveneTest(DBH.year.6~Treatment, data = DBHdata)
leveneTest(DBH.year.6~Large, data = DBHdata)

leveneTest(log(DBH.year.6)~Treatment, data = DBHdata) # constant variance
leveneTest(log(DBH.year.6)~Large,data = DBHdata) # different variance

leveneTest(DBH.year.6~Treatment, data = DBHpositive) # constant variance
leveneTest(DBH.year.6~Large, data = DBHpositive) # constant variance

leveneTest(log(DBH.year.6)~Treatment, data = DBHpositive) # 2nd most constant variance
leveneTest(log(DBH.year.6)~Large, data = DBHpositive) # 3rd most different variance

leveneTest(DBH.year.6~Treatment, data = DBHdatarm)
leveneTest(DBH.year.6~Large, data = DBHdatarm) # constant variance

leveneTest(log(DBH.year.6)~Treatment, data = DBHdatarm) # 3rd most constant variance
leveneTest(log(DBH.year.6)~Large,data = DBHdatarm) # 2nd most different variance

leveneTest(DBH.year.6~Treatment, data = DBHpositiverm) # constant variance
leveneTest(DBH.year.6~Large, data = DBHpositiverm) # constant variance

leveneTest(log(DBH.year.6)~Treatment, data = DBHpositiverm) # most constant variance
leveneTest(log(DBH.year.6)~Large, data = DBHpositiverm) # most different variance
```
```{r}
library(fitdistrplus)
descdist(log(DBHdata$DBH.year.6))
descdist(log(DBHpositive$DBH.year.6)) # close to normal
descdist(log(DBHdatarm$DBH.year.6)) # could be considered as normal
descdist(log(DBHpositiverm$DBH.year.6)) # could be considered as normal

descdist(DBHdatarm$DBH.year.6)
descdist(DBHpositiverm$DBH.year.6)
```

log(DBHpositive$DBH.year.6), log(DBHdatarm$DBH.year.6) and log(DBHpositiverm$DBH.year.6) share same properties:
1. close to normal distribution, or could be considered as normal distribution.
2. constant variance across Treatment groups(top3 largest Levene's test Pr(>F)value), but very different variance across Large groups(top3 smallest Levene's test Pr(>F)value).
  In other word, they have this Large-group-variance-difference situation most significantly. This is actually not a very bad thing if we can make the model capture this Large-group-variance-difference.



# fit glm models
gaussian glm for DBHdata$DBH.year.6, could be considered as a baseline model
```{r}
# best model derived from DBH.year.6~DBH.year.0*Treatment*Large
glm0 = glm(DBH.year.6~DBH.year.0+Treatment+Large+DBH.year.0:Large+Treatment:Large, data = DBHdata)
summary(glm0)

tss = sum((DBHdata$DBH.year.6-mean(DBHdata$DBH.year.6))^2) #TSS
rss = sum((DBHdata$DBH.year.6-glm0$fitted.values)^2) #RSS
(tss-rss)/tss #R^2 = 0.9467556

par(mfrow=c(2,2))
plot(glm0)
par(mfrow=c(1,1))
plot(DBHdata$DBH.year.6, glm0$fitted.values, type = "p", col = DBHdata$Treatment)
legend("bottomright", legend = c("Control", "T1", "T2"), fill = c("black", "red", "green"))
abline(0,1)
plot(DBHdata$DBH.year.6, glm0$fitted.values, type = "p", col = DBHdata$Large)
legend("bottomright", legend = c("not-large", "large"), fill = c("black", "red"))
abline(0,1)
```

gaussian glm for DBHpositiverm$DBH.year.6. It has constant variance across all groups and has lognormal distribution.
```{r}
glm1.0 = glm(DBH.year.6~DBH.year.0+Treatment+Large+DBH.year.0:Large+Treatment:Large, data = DBHpositiverm)
summary(glm1.0)

tss = sum((DBHpositiverm$DBH.year.6-mean(DBHpositiverm$DBH.year.6))^2) #TSS
rss = sum((DBHpositiverm$DBH.year.6-glm1.0$fitted.values)^2) #RSS
(tss-rss)/tss #R^2 = 0.9801139

par(mfrow=c(2,2))
plot(glm1.0)
par(mfrow=c(1,1))
plot(DBHpositiverm$DBH.year.6, glm1.0$fitted.values, type = "p", col = DBHpositiverm$Treatment)
legend("bottomright", legend = c("Control", "T1", "T2"), fill = c("black", "red", "green"))
abline(0,1)
plot(DBHpositiverm$DBH.year.6, glm1.0$fitted.values, type = "p", col = DBHpositiverm$Large)
legend("bottomright", legend = c("not-large", "large"), fill = c("black", "red"))
abline(0,1)
```

gaussian glm for log(DBHpositive$DBH.year.6)
```{r}
glm1.1 = glm(log(DBH.year.6)~log(DBH.year.0)+Treatment+Large+log(DBH.year.0):Large+Treatment:Large, data = DBHpositive)
summary(glm1.1)

tss = sum((log(DBHpositive$DBH.year.6)-mean(log(DBHpositive$DBH.year.6)))^2) #TSS
rss = sum((log(DBHpositive$DBH.year.6)-glm1.1$fitted.values)^2) #RSS
(tss-rss)/tss #R^2 = 0.9554377

par(mfrow=c(2,2))
plot(glm1.1)
par(mfrow=c(1,1))
plot(log(DBHpositive$DBH.year.6), glm1.1$fitted.values, type = "p", col = DBHpositive$Treatment)
legend("bottomright", legend = c("Control", "T1", "T2"), fill = c("black", "red", "green"))
abline(0,1)
plot(log(DBHpositive$DBH.year.6), glm1.1$fitted.values, type = "p", col = DBHpositive$Large)
legend("bottomright", legend = c("not-large", "large"), fill = c("black", "red"))
abline(0,1)
```

gaussian glm for log(DBHdatarm$DBH.year.6)
```{r}
glm1.2 = glm(log(DBH.year.6)~log(DBH.year.0)+Treatment+Large+log(DBH.year.0):Large+Treatment:Large, data = DBHdatarm)
summary(glm1.2)
anova(glm1.2)

tss = sum((log(DBHdatarm$DBH.year.6)-mean(log(DBHdatarm$DBH.year.6)))^2) #TSS
rss = sum((log(DBHdatarm$DBH.year.6)-glm1.2$fitted.values)^2) #RSS
(tss-rss)/tss #R^2 = 0.9783684

par(mfrow=c(2,2))
plot(glm1.2)
par(mfrow=c(1,1))
plot(log(DBHdatarm$DBH.year.6), glm1.2$fitted.values, type = "p", col = DBHdatarm$Treatment)
legend("bottomright", legend = c("Control", "T1", "T2"), fill = c("black", "red", "green"))
abline(0,1)
plot(log(DBHdatarm$DBH.year.6), glm1.2$fitted.values, type = "p", col = DBHdatarm$Large)
legend("bottomright", legend = c("not-large", "large"), fill = c("black", "red"))
abline(0,1)
```

gaussian glm for log(DBHpositiverm$DBH.year.6)
```{r}
glm1.3 = glm(log(DBH.year.6)~log(DBH.year.0)+Treatment+Large+log(DBH.year.0):Large+Treatment:Large, data = DBHpositiverm)
summary(glm1.3)
anova(glm1.3)

tss = sum((log(DBHpositiverm$DBH.year.6)-mean(log(DBHpositiverm$DBH.year.6)))^2) #TSS
rss = sum((log(DBHpositiverm$DBH.year.6)-glm1.3$fitted.values)^2) #RSS
(tss-rss)/tss #R^2 = 0.9826943

par(mfrow=c(2,2))
plot(glm1.3)
par(mfrow=c(1,1))
plot(log(DBHpositiverm$DBH.year.6), glm1.3$fitted.values, type = "p", col = DBHpositiverm$Treatment)
legend("bottomright", legend = c("Control", "T1", "T2"), fill = c("black", "red", "green"))
abline(0,1)
plot(log(DBHpositiverm$DBH.year.6), glm1.3$fitted.values, type = "p", col = DBHpositiverm$Large)
legend("bottomright", legend = c("not-large", "large"), fill = c("black", "red"))
abline(0,1)
```

# ------------------------------------------------------------------------------------------
- All models fit the data very well with very high R^2 values.This is due to the very strong linear correlation between the response variable DBH.year.6 & log(DBH.year.6) and one predictor DBH.year.0 & log(DBH.year.0).
- However, even though with the presence of DBH.year.0, the other two categorical predictors (Treatment, Large) still are very significant.
- The result formula is the same for all models. In some models, intercept is not significant, we are keeping it at present.
- The estimated parameters satisfy all the hypotheses (in a similar way as previous models):
  1. T2>T1>Control, large>not-large
  2. being T2, and especially being T2&Large has great increase in tree growth.
  3. the estimated parameter for term DBH.year.0:LargeYes is a small negative value.
  
# ------------------------------------------------------------------------------------------
gaussian glm for log(DBHdatarm$DBH.year.6) seems to be the "best" model so far. Although it has the 3rd largest R^2 value (0.9783684 < 0.9801139 < 0.9826943), it has the smallest residuals and a normal distribution.
- Guess (not sure): there might be some improvement if we can take in account Large-group-variance-difference? but HOW?
- Need very convincing explanation for the removed points.

gaussian glm for DBHpositiverm$DBH.year.6 might also be a considerable model. It has constant variance across all groups and has lognormal distribution. It has the 2nd largest R^2 value (0.9801139). Although its residuals have larger range than other log() models, compared to the range of predicted values, they are actually not that very large.
- Still need improvement if choose this model, and need very convincing explanation for the removed points.
# ------------------------------------------------------------------------------------------






# random effect
```{r}
par(mfrow=c(3,4))
for (i in 1:12) {
  plot(DBHdata[DBHdata$Plot == i, "DBH.year.0"], DBHdata[DBHdata$Plot == i, "DBH.year.6"], xlab = paste("Plot", i, "DBH.year.0"), ylab = paste("Plot", i, "DBH.year.6"), xlim = c(20,70), ylim = c(20,70), col = DBHdata[DBHdata$Plot == i, "Treatment"])
  abline(0,1)
}

for (i in 1:12) {
  plot(log(DBHdata[DBHdata$Plot == i, "DBH.year.0"]), log(DBHdata[DBHdata$Plot == i, "DBH.year.6"]), xlab = paste("Plot", i, "log(DBH.year.0)"), ylab = paste("Plot", i, "log(DBH.year.6)"), xlim = c(3.0,4.3), ylim = c(3.0, 4.3), col = DBHdata[DBHdata$Plot == i, "Treatment"])
  abline(0,1)
}
```

random effect model for log(DBHdatarm$DBH.year.0)
```{r}
mixed1.0 = lmer(log(DBH.year.6)~log(DBH.year.0)+Large+Treatment+log(DBH.year.0):Large+Treatment:Large+(1|Plot), data = DBHdatarm)
summary(mixed1.0)
VarCorr(mixed1.0)
y.hat <- fitted(mixed1.0) # Fitted values
int.hat <- ranef(mixed1.0)[[1]][[1]] # Predicted intercepts
res.hat <- residuals(mixed1.0) # Estimated residuals
tss = sum((log(DBHdatarm$DBH.year.6)-mean(log(DBHdatarm$DBH.year.6)))^2) #TSS
rss = sum((log(DBHdatarm$DBH.year.6)-y.hat)^2) #RSS
(tss-rss)/tss #R^2 = 0.9800758
```
```{r}
qqnorm(int.hat, main="Random Intercepts"); qqline(int.hat)
qqnorm(res.hat, main="Residuals"); qqline(res.hat)
plot(y.hat, res.hat, xlab="Fitted Values", ylab="Residuals")
abline(h=0, lty=2)
plot(log(DBHdatarm$DBH.year.6), y.hat, col = DBHdatarm$Treatment)
abline(0, 1)
plot(log(DBHdatarm$DBH.year.6), y.hat, col = DBHdatarm$Large)
abline(0, 1)
```

random effect model for log(DBHpositiverm$DBH.year.0)
```{r}
mixed1.1 = lmer(log(DBH.year.6)~log(DBH.year.0)+Large+Treatment+log(DBH.year.0):Large+Treatment:Large+(1|Plot), data = DBHpositiverm)
summary(mixed1.1)
VarCorr(mixed1.1)
y.hat <- fitted(mixed1.1) # Fitted values
int.hat <- ranef(mixed1.1)[[1]][[1]] # Predicted intercepts
res.hat <- residuals(mixed1.1) # Estimated residuals
tss = sum((log(DBHpositiverm$DBH.year.6)-mean(log(DBHpositiverm$DBH.year.6)))^2) #TSS
rss = sum((log(DBHpositiverm$DBH.year.6)-y.hat)^2) #RSS
(tss-rss)/tss #R^2 = 0.9839044
```
```{r}
qqnorm(int.hat, main="Random Intercepts"); qqline(int.hat)
qqnorm(res.hat, main="Residuals"); qqline(res.hat)
plot(y.hat, res.hat, xlab="Fitted Values", ylab="Residuals")
abline(h=0, lty=2)
plot(log(DBHpositiverm$DBH.year.6), y.hat, col = DBHpositiverm$Treatment)
abline(0, 1)
plot(log(DBHpositiverm$DBH.year.6), y.hat, col = DBHpositiverm$Large)
abline(0, 1)
```

# ------------------------------------------------------------------------------------------
random effect is not significant.
From the 12 scatter plots, it seems that there's no random intercept or random slope.
But I think it's still necessary to keep the random intercept term as a reference to human errors.
# ------------------------------------------------------------------------------------------






# "In Gap" term
add "InGap" term
```{r}
DBHdata$InGap = (DBHdata$Treatment == "T2: Radial" & DBHdata$Large == "Yes")
DBHdatarm$InGap = (DBHdatarm$Treatment == "T2: Radial" & DBHdatarm$Large == "Yes")
DBHpositive$InGap = (DBHpositive$Treatment == "T2: Radial" & DBHpositive$Large == "Yes")
DBHpositiverm$InGap = (DBHpositiverm$Treatment == "T2: Radial" & DBHpositiverm$Large == "Yes")
```
make InGap factor
```{r}
DBHdata$InGap = factor(DBHdata$InGap)
DBHdatarm$InGap = factor(DBHdatarm$InGap)
DBHpositive$InGap = factor(DBHpositive$InGap)
DBHpositiverm$InGap = factor(DBHpositiverm$InGap)
```
```{r}
plot(DBHdata$DBH.year.0, DBHdata$DBH.year.6, type = "p", col = DBHdata$InGap)
legend("bottomright", legend = c("not InGap", "InGap"), fill = c("black", "red"))

plot(DBHdata[DBHdata$Treatment == "T2: Radial", "DBH.year.0"], DBHdata[DBHdata$Treatment == "T2: Radial", "DBH.year.6"], type = "p", col = DBHdata[DBHdata$Treatment == "T2: Radial", "InGap"])
legend("bottomright", legend = c("T2 not InGap", "InGap"), fill = c("black", "red"))

plot(DBHdata[DBHdata$Treatment == "T1: Gap" | DBHdata$InGap == TRUE, "DBH.year.0"], DBHdata[DBHdata$Treatment == "T1: Gap" | DBHdata$InGap == TRUE, "DBH.year.6"], type = "p", col = DBHdata[DBHdata$Treatment == "T1: Gap" | DBHdata$InGap == TRUE, "InGap"])
legend("bottomright", legend = c("T1", "InGap"), fill = c("black", "red"))

plot(DBHdata[DBHdata$Treatment == "Control" | DBHdata$InGap == TRUE, "DBH.year.0"], DBHdata[DBHdata$Treatment == "Control" | DBHdata$InGap == TRUE, "DBH.year.6"], type = "p", col = DBHdata[DBHdata$Treatment == "Control" | DBHdata$InGap == TRUE, "InGap"])
legend("bottomright", legend = c("Control", "InGap"), fill = c("black", "red"))
```
```{r}
leveneTest(DBH.year.6~InGap, data = DBHdata)
leveneTest(log(DBH.year.6)~InGap, data = DBHdata)

leveneTest(DBH.year.6~InGap, data = DBHdatarm)
leveneTest(log(DBH.year.6)~InGap, data = DBHdatarm)

leveneTest(DBH.year.6~InGap, data = DBHpositive)
leveneTest(log(DBH.year.6)~InGap, data = DBHpositive)

leveneTest(DBH.year.6~InGap, data = DBHpositiverm)
leveneTest(log(DBH.year.6)~InGap, data = DBHpositiverm)
```
# ------------------------------------------------------------------------------------------
Problem with log(DBH.year.6)~log(DBH.year.0)+Treatment+Large+log(DBH.year.0):InGap + Treatment:InGap
1. Treatment:InGap is a useless term? because T1:InGap = NA
2. Large is not significant when there's InGap term
# ------------------------------------------------------------------------------------------

glm using InGap for log(DBHdatarm$DBH.year.6)
```{r}
glm2.0 = glm(log(DBH.year.6)~log(DBH.year.0)+Treatment+InGap+log(DBH.year.0):InGap, data = DBHdatarm)
summary(glm2.0)

tss = sum((log(DBHdatarm$DBH.year.6)-mean(log(DBHdatarm$DBH.year.6)))^2) #TSS
rss = sum((log(DBHdatarm$DBH.year.6)-glm2.0$fitted.values)^2) #RSS
(tss-rss)/tss #R^2 = 0.9786269

par(mfrow=c(2,2))
plot(glm2.0)
par(mfrow=c(1,1))
plot(log(DBHdatarm$DBH.year.6), glm2.0$fitted.values, type = "p", col = DBHdatarm$Treatment)
legend("bottomright", legend = c("Control", "T1", "T2"), fill = c("black", "red", "green"))
abline(0,1)
plot(log(DBHdatarm$DBH.year.6), glm2.0$fitted.values, type = "p", col = DBHdatarm$Large)
legend("bottomright", legend = c("not-large", "large"), fill = c("black", "red"))
abline(0,1)
plot(log(DBHdatarm$DBH.year.6), glm2.0$fitted.values, type = "p", col = DBHdatarm$InGap)
legend("bottomright", legend = c("not In Gap", "In Gap"), fill = c("black", "red"))
abline(0,1)
```

random effect using InGap for log(DBHdatarm$DBH.year.6)
```{r}
mixed2.0 = lmer(log(DBH.year.6)~log(DBH.year.0)+Treatment+InGap+log(DBH.year.0):InGap+(1|Plot), data = DBHdatarm)
summary(mixed2.0)
VarCorr(mixed2.0)
y.hat <- fitted(mixed2.0) # Fitted values
int.hat <- ranef(mixed2.0)[[1]][[1]] # Predicted intercepts
res.hat <- residuals(mixed2.0) # Estimated residuals
tss = sum((log(DBHdatarm$DBH.year.6)-mean(log(DBHdatarm$DBH.year.6)))^2) #TSS
rss = sum((log(DBHdatarm$DBH.year.6)-y.hat)^2) #RSS
(tss-rss)/tss #R^2 = 0.9801595
```
```{r}
qqnorm(int.hat, main="Random Intercepts"); qqline(int.hat)
qqnorm(res.hat, main="Residuals"); qqline(res.hat)
plot(y.hat, res.hat, xlab="Fitted Values", ylab="Residuals")
abline(h=0, lty=2)
plot(log(DBHdatarm$DBH.year.6), y.hat, col = DBHdatarm$Treatment)
abline(0, 1)
plot(log(DBHdatarm$DBH.year.6), y.hat, col = DBHdatarm$Large)
abline(0, 1)
plot(log(DBHdatarm$DBH.year.6), y.hat, col = DBHdatarm$InGap)
abline(0, 1)
```

glm using InGap for log(DBHpositiverm$DBH.year.6)
```{r}
glm2.1 = glm(log(DBH.year.6)~log(DBH.year.0)+Treatment+InGap+log(DBH.year.0):InGap, data = DBHpositiverm)
summary(glm2.1)

tss = sum((log(DBHpositiverm$DBH.year.6)-mean(log(DBHpositiverm$DBH.year.6)))^2) #TSS
rss = sum((log(DBHpositiverm$DBH.year.6)-glm2.1$fitted.values)^2) #RSS
(tss-rss)/tss #R^2 = 0.982808

par(mfrow=c(2,2))
plot(glm2.1)
par(mfrow=c(1,1))
plot(log(DBHpositiverm$DBH.year.6), glm2.1$fitted.values, type = "p", col = DBHpositiverm$Treatment)
legend("bottomright", legend = c("Control", "T1", "T2"), fill = c("black", "red", "green"))
abline(0,1)
plot(log(DBHpositiverm$DBH.year.6), glm2.1$fitted.values, type = "p", col = DBHpositiverm$Large)
legend("bottomright", legend = c("not-large", "large"), fill = c("black", "red"))
abline(0,1)
plot(log(DBHpositiverm$DBH.year.6), glm2.1$fitted.values, type = "p", col = DBHpositiverm$InGap)
legend("bottomright", legend = c("not In Gap", "In Gap"), fill = c("black", "red"))
abline(0,1)
```

random effect using InGap for log(DBHpositiverm$DBH.year.6)
```{r}
mixed2.1 = lmer(log(DBH.year.6)~log(DBH.year.0)+Treatment+InGap+log(DBH.year.0):InGap+(1|Plot), data = DBHpositiverm)
summary(mixed2.1)
VarCorr(mixed2.1)
y.hat <- fitted(mixed2.1) # Fitted values
int.hat <- ranef(mixed2.1)[[1]][[1]] # Predicted intercepts
res.hat <- residuals(mixed2.1) # Estimated residuals
tss = sum((log(DBHpositiverm$DBH.year.6)-mean(log(DBHpositiverm$DBH.year.6)))^2) #TSS
rss = sum((log(DBHpositiverm$DBH.year.6)-y.hat)^2) #RSS
(tss-rss)/tss #R^2 = 0.9840615
```
```{r}
qqnorm(int.hat, main="Random Intercepts"); qqline(int.hat)
qqnorm(res.hat, main="Residuals"); qqline(res.hat)
plot(y.hat, res.hat, xlab="Fitted Values", ylab="Residuals")
abline(h=0, lty=2)
plot(log(DBHpositiverm$DBH.year.6), y.hat, col = DBHpositiverm$Treatment)
abline(0, 1)
plot(log(DBHpositiverm$DBH.year.6), y.hat, col = DBHpositiverm$Large)
abline(0, 1)
plot(log(DBHpositiverm$DBH.year.6), y.hat, col = DBHpositiverm$InGap)
abline(0, 1)
```






# Interpretation
# ------------------------------------------------------------------------------------------
ACHIEVEMENTS:
1. All models can explain priority1

PROBLEMS:
1. Cannot priority2(tree growth for large trees) because "Large" is insignificant when there's "InGap" term.
    -> use scatter plot of tree.growth vs DBH.year.0 to explain
2. Cannot say bigger DBH.year.0/tree size gives bigger tree growth (priority3.1), 
   INSTEAD, we can say bigger DBH.year.0/tree size results in bigger DBH.year.6)
    -> use scatter plot of tree.growth vs DBH.year.0 to explain
3. Coefficients for InGapTRUE and log(DBH.year.0):InGapTRUE form a constrain for InGap tree growth > not InGap tree growth.
    -> not sure how to interpret these constrains
    -> guess: much fewer data points over constrain, so the model might not be very accurate
# ------------------------------------------------------------------------------------------

```{r}
plot(DBHdatarm$DBH.year.0, DBHdatarm$DBH.year.6, col = DBHdatarm$InGap)
abline(v = 50.5221)
```
```{r}
DBHdatarm[DBHdatarm$DBH.year.0 > 50.5221, ]
```






# Futher work
1. examine the 11 points classified as outliers in the original datasheet








