---
title: "final result"
output:
  word_document: default
  html_document: default
---

This file contains the final model and all its visualisations. The content is identical to Tree model result interpretation in the report.

# load packages
```{r}
setwd("~/Desktop/capstone/Datascience/DataScienceProject/Tree DBH")
library(lme4)
library(merTools)
library(car)
```

# data preparation
```{r}
# load data
data = read.csv("~/Desktop/capstone/Datascience/DataScienceProject/Tree DBH/dataset/all_plots.csv", stringsAsFactors = FALSE)

# only extract DBH-related data for DBH modelling
DBHdata = data[1:9]
```

check NAs and blank
```{r}
# blank "Large", assume "No" where it's blank
DBHdata[DBHdata$Large == "", "Tree.Number"]
DBHdata[DBHdata$Large == "", "Large"] = "No"
```
```{r}
# remove rows with NA in "DBH.year.6"
DBHdata[is.na(DBHdata$DBH.year.6), 1:7] # 9 NAs for "DBH.year.6"
DBHdata = DBHdata[!is.na(DBHdata$DBH.year.6), ]
```
#------------------------------------------------------------------------------------------
Referring to lower priority hypotheses3.2 (The rate of natural mortality will be greater in control plots as compared to T1 and T2.): 
NOT satisfied: all blank Alive.at.year.6 are in T2:Radial.
CANNOT make any conclusion, need more data, revisit this hypothesis in the future.
#-------------------------------------------------------------------------------------------
```{r}
# remove rows with NA in "DBH.year.0"
DBHdata[is.na(DBHdata$DBH.year.0), 1:7] # 2 NAs for "DBH.year.0"
DBHdata = DBHdata[!is.na(DBHdata$DBH.year.0), ]
```
```{r}
# NA in "Class.year.0" and NA in "Class.year.6", no modification
DBHdata[is.na(DBHdata$Class.year.0) | is.na(DBHdata$Class.year.6), ]
# both in Plot5(T2: Radial), Tree.Number 148, 149, they have extremely large DBH.
```

make Treatment and Large factor
```{r}
DBHdata$Treatment = factor(DBHdata$Treatment)
DBHdata$Large = factor(DBHdata$Large)
```

list rows where DBH.year.6 <= DBH.year.0 (tree growth <= 0)
```{r}
DBHdata[DBHdata$DBH.year.6 <= DBHdata$DBH.year.0, ]
```






# data analysis
```{r}
# visualise and list points outside the main group
plot(DBHdata$DBH.year.0, DBHdata$DBH.year.6, type = "p", col = factor(DBHdata$DBH.year.6/DBHdata$DBH.year.0 > 1.25 | DBHdata$DBH.year.6/DBHdata$DBH.year.0 < 0.9 | DBHdata$DBH.year.0 > 80))
plot(DBHdata$DBH.year.0, DBHdata$DBH.year.6/DBHdata$DBH.year.0, type = "p", col = factor(DBHdata$DBH.year.6/DBHdata$DBH.year.0 > 1.25 | DBHdata$DBH.year.6/DBHdata$DBH.year.0 < 0.9 | DBHdata$DBH.year.0 > 80), ylab = "gradient")
DBHdata[DBHdata$DBH.year.6/DBHdata$DBH.year.0 > 1.25 | DBHdata$DBH.year.6/DBHdata$DBH.year.0 < 0.9 | DBHdata$DBH.year.0 > 80, ]

# remove these outliers
DBHdatarm = DBHdata[DBHdata$DBH.year.6/DBHdata$DBH.year.0 < 1.25 & DBHdata$DBH.year.6/DBHdata$DBH.year.0 > 0.9 & DBHdata$DBH.year.0 < 80, ]
```

```{r}
# strong correlation between DBH.year.6 and DBH.year.0
cor(DBHdata$DBH.year.0, DBHdata$DBH.year.6)
cor(log(DBHdata$DBH.year.0), log(DBHdata$DBH.year.6))

cor(DBHdatarm$DBH.year.0, DBHdatarm$DBH.year.6)
cor(log(DBHdatarm$DBH.year.0), log(DBHdatarm$DBH.year.6))
```






# random effect
```{r}
par(mfrow=c(3,4))
for (i in 1:12) {
  plot(log(DBHdata[DBHdata$Plot == i, "DBH.year.0"]), log(DBHdata[DBHdata$Plot == i, "DBH.year.6"]), xlab = paste("Plot", i, "log(DBH.year.0)"), ylab = paste("Plot", i, "log(DBH.year.6)"), xlim = c(3.0,4.3), ylim = c(3.0, 4.3), col = DBHdata[DBHdata$Plot == i, "Treatment"])
  abline(0,1)
}
```
There's no obvious random intercept or random slope, but we still keep the random intercept term as a reference to human error.






# "In Gap" term
```{r}
# add "InGap" term
DBHdatarm$InGap = (DBHdatarm$Treatment == "T2: Radial" & DBHdatarm$Large == "Yes")

# make InGap factor
DBHdatarm$InGap = factor(DBHdatarm$InGap)
```






# current final model
mixed effect model for log(DBHdatarm$DBH.year.6) using InGap
```{r}
mixed2.0 = lmer(log(DBH.year.6)~log(DBH.year.0)+Treatment+InGap+log(DBH.year.0):InGap+(1|Plot), data = DBHdatarm)
summary(mixed2.0)
VarCorr(mixed2.0)
y.hat <- fitted(mixed2.0) # Fitted values
int.hat <- ranef(mixed2.0)[[1]][[1]] # Predicted intercepts
res.hat <- residuals(mixed2.0) # Estimated residuals
tss = sum((log(DBHdatarm$DBH.year.6)-mean(log(DBHdatarm$DBH.year.6)))^2) # TSS
rss = sum((log(DBHdatarm$DBH.year.6)-y.hat)^2) # RSS
(tss-rss)/tss # R^2 = 0.9801595
```
```{r}
qqnorm(int.hat, main="Random Intercepts"); qqline(int.hat)
qqnorm(res.hat, main="Normal Q-Q"); qqline(res.hat)
plot(y.hat, res.hat, xlab="Fitted Values", ylab="Residuals", main = "Residuals vs Fitted")
abline(h=0, lty=2)
plot(log(DBHdatarm$DBH.year.6), y.hat, col = DBHdatarm$Treatment)
abline(0, 1)
plot(log(DBHdatarm$DBH.year.6), y.hat, col = DBHdatarm$Large)
abline(0, 1)
plot(log(DBHdatarm$DBH.year.6), y.hat, col = DBHdatarm$InGap)
abline(0, 1)
```






# visualisations
```{r}
# plot constrains according to mixed2.0
plot(DBHdatarm$DBH.year.0, DBHdatarm$DBH.year.6, col = DBHdatarm$InGap, xlim = c(20,80), xlab = "DBH at Year 0", ylab = "DBH at Year 6", main = "Constraints in the final model")
abline(v = 52.76, col = "blue") # constraint for T2InGap > T2 not InGap
abline(v = 57.87, col = "green") # constraint for T2InGap > T1
abline(v = 74.42, col = "orange") # constraint for T2InGap > Control
legend("topleft", legend = c("InGap points", "not InGap points", "InGap>T2notInGap constrain", "InGap>T1 constrain", "InGap>Control constrain"), fill = c("black", "red", "blue", "green", "orange"), cex = 0.9, bty='n')
```
Interpretation for the constrains: sparse data in the constrain region, so the fitted model cannot explain well. If more data is obtained in the future, results would be more realistic.

```{r}
# visualise predicted interval for the three treatments
treatments = c("Control", "T1: Gap", "T2: Radial")
shades = c("light grey", rgb(255,0,0, 50, maxColorValue = 255), rgb(0,255,0, 50, maxColorValue = 255))
colours = c("black", "red", "green")

plot(1:100, exp(PI$fit), type='n', xlim=c(20,60), ylim = c(15,60), xlab = "DBH: Year 0", ylab = "Estimated DBH: Year 6",
     main='Tree Size Model: Treatments vs Control')
for (i in 1:3) {
  X <- data.frame('DBH.year.0'=1:100, 'InGap'=FALSE, 'Plot'=6, 'Treatment'=treatments[i])
  PI <- predictInterval(merMod = mixed2.0, newdata = X, level = 0.95, n.sims = 1000, stat = "median", type="linear.prediction", include.resid.var = TRUE)
  
  polygon(c(1:100, 100:1), c(exp(PI$upr), rev(exp(PI$lwr))), col = shades[i], border = shades[i])
  lines(1:100, exp(PI$fit), type = "l", col = colours[i])
}
legend("topleft", legend = c("Control", "T1: Gap", "T2: Radial (Not in gap)", '95% CI'), 
       lty=c(1,1,1,NA),
       col= c('black','red','green',NA), bty='n')
```
```{r}
# hypothesis test for T1 = Control
C <- c(0,0,1,0,0,0)
linearHypothesis(mixed2.0, C) # Pr(>Chisq) = 0.0001243, reject H0, so T1 != Control

# hypothesis test for T2 = Control
C <- c(0,0,0,1,0,0)
linearHypothesis(mixed2.0, C) # Pr(>Chisq) = 8.334e-07, reject H0, so T2 != Control

# hypothesis test for T1 = T2
C <- c(0,0,1,-1,0,0)
linearHypothesis(mixed2.0, C) # Pr(>Chisq) = 0.1879, cannot reject H0 at 5% level or 1% level, so keep H0, T1 = T2
```
This plot shows fitted line and 95% predict interval for not-InGap trees in the three treatments.
The red line(T1) and the green line(T2) are both higher and steeper than the black line(Control), but lines and shaded intervals of T1 and T2 almost completely overlap.

Results of the hypothesis tests prove the pattern in the plot that T1 or T2 is significantly different from Control, but there's not enough evidence to say that T1 is different from T2 for not-InGap trees.

Interpretation for T1, T2 > Control:
This responses to hypothesis1.1 (Thinning will increase tree growth in both treatments relative to the control.)

Interpretation for T1 = T2 for not-InGap trees:
1. What do these two groups refer to? 
    -> T1 & not-InGap = all T1 trees = trees around the T1gap area + trees not around the T1gap area
    -> T2 & not-InGap = trees except for the ones in the middle of the T2radial areas = trees around the T2radial area + trees not around the T2radial area.
No matter it's a T1gap area or a T2radial area, it's always an open space, SO basically, trees in the two groups could be considered as having the same "treatment" situations. SO T1=T2 (i.e. having almost completely overlapping shades and lines) is reasonable.
2. In the perspective of the estimated parameter values for these two terms:
    -> TreatmentT1: Gap is 0.028683, TreatmentT2: Rad is 0.039231
Calculating back the log(), T1 contributes exp(0.028683)=1.029 gradient, T2 contributes exp(0.039231)=1.04 gradient. 1.029 is very close to 1.04. Also, intercepts for the lines are very very close to each other as well, SO the red and green lines and their shades are almost completely overlapping is still reasonable.

Although T1=T2, for hypothesis 1.2 (Thinning will have an increased effect on the rate of tree growth in T2 as compared to T1.) Keep te note: For trees starting with the same DBH.year.0 value, being in T1 will have a 2.9% increase in DBH.year.6 value compared to DBH.year.6 in Control; and being in T2 will have a 4% increase in DBH.year.6 value compared to DBH.year.6 in Control. And 2.9% is clearly different from 4%.
    
```{r}
# visualise predicted interval for InGap VS Control
X_ingap <- data.frame('DBH.year.0'=30:100, 'InGap'=TRUE, 'Plot'=6, 'Treatment'='T2: Radial')
PI_2 <- predictInterval(merMod = mixed2.0, newdata = X_ingap, level = 0.95, n.sims = 1000, stat = "median", type="linear.prediction", include.resid.var = TRUE)

X <- data.frame('DBH.year.0'=1:100, 'InGap'=FALSE, 'Plot'=6, 'Treatment'='Control')
PI <- predictInterval(merMod = mixed2.0, newdata = X, level = 0.95, n.sims = 1000, stat = "median", type="linear.prediction", include.resid.var = TRUE)

plot(1:100, exp(PI$fit), type='n', xlim=c(20,60), ylim = c(15,60), xlab = "DBH: Year 0", ylab = "Estimated DBH: Year 6",
     main="Tree Size Model: 'In Gap' vs Control")

polygon(c(1:100, 100:1), c(exp(PI$upr), rev(exp(PI$lwr))), col = "light grey", border = "light grey")
lines(1:100, exp(PI$fit), type = "l")
polygon(c(30:100, 100:30), c(exp(PI_2$upr), rev(exp(PI_2$lwr))), col = rgb(255,0,0, 50, maxColorValue = 255), border = rgb(255,0,0, 50, maxColorValue = 255))
lines(30:100, exp(PI_2$fit), type = "l", col = "red")
legend("topleft", legend = c('Control',"T2: Radial (In gap)", '95% CI'),
       lty=c(1,1,NA),
       col= c('black','red','green',NA), bty='n')
```
```{r}
# hypothesis test for InGap = Control
C <- c(0,0,0,1,1,log(55))
linearHypothesis(mixed2.0, C) # Pr(>Chisq) = 0.006618, reject H0 at 1%, for DBH.year.0 <=55, InGap != Control
C <- c(0,0,0,1,1,log(57))
linearHypothesis(mixed2.0, C) # Pr(>Chisq) = 0.0244, reject H0 at 5%
C <- c(0,0,0,1,1,log(59))
linearHypothesis(mixed2.0, C) # Pr(>Chisq) = 0.06454, cannot reject H0 at 5%
```
The plot shows fitted line and 95% predict interval for InGap VS Control. The red line(InGap) is higher than the black line(Control), but shaded intervals have some overlap.

Results of the hypothesis tests gives threshold.



# other plots
```{r}
# add tree growth for visualisation in response to hypothesis2.1&2.2
DBHdatarm$Tree.growth = DBHdatarm$DBH.year.6-DBHdatarm$DBH.year.0

plot(DBHdatarm$DBH.year.0, DBHdatarm$Tree.growth, col = DBHdatarm$Large, ylim = c(-3,8.5), xlab = "DBH at Year 0", ylab = "Tree growth = DBH.year.6 - DBH.year.0", main = "Comparing tree growth with tree size")
legend("topleft", legend = c("Large", "not Large"), fill = c("red", "black"), bty='n')
```
From the plot, large trees will only increase variance in tree growth.


